---
title: "Create suitability scatter plots and binarized maps (chapter 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2024-03-06"
output: html_document
---
# Overview

EDIT

This vignette will compare the invaded region model for SLF, which was created in `050_run_easternUSA_MaxEnt_models.Rmd`, with the global version of this model, which was created in `070_run_global_historical_MaxEnt_model.Rmd`. First, I will compare these models using flexible AUC. Second, I will also project these models to N. America and compare the proportion of suitable area. Third, I will calculate suitability values for all SLF points and IVR regions. Finally, I will perform an invasion stage analysis following the methods of Gallien et.al, 2012. 

In step 1, I tested the best fit background data using SLF occurrences and climate data only from the SLF invaded range (the eastern half of the United States). 

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here() # here() starts at the root folder of this package.
library(devtools)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# spatial data handling
library(raster) 
library(terra)
library(scrubr)

# plot aesthetics
library(scales)
#install.packages("ggbreak")
library(ggbreak)
library(patchwork)

```

```{r set wd}

mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models")

```

Load in summary files for thresholds.

```{r load in summary files}
# summary file to extract thresholds from

# global
summary_global <- read_csv(file = file.path(mypath, "slf_global_v2", "global_summary_all_iterations.csv"))


# regional_native
summary_regional_invaded <- read_csv(file = file.path(mypath, "slf_regional_invaded_v6", "regional_invaded_summary.csv"))


# regional_native
summary_regional_native <- read_csv(file = file.path(mypath, "slf_regional_native_v2", "regional_native_summary.csv"))

```

Load in dataset containing locations of globally important viticultural regions.

```{r load in wineries dataset}

load(file.path(here(), "data", "wineries.rda"))

IVR_locations <- wineries %>%
  # tidy
  dplyr::select(-c(Latitude, Longitude)) %>%
  dplyr::filter(
    !is.na(x),
    !is.na(y)
    ) %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) 

# remove unnecessary obj
rm(wineries)

```

```{r axis breaks}

breaks <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)

```

# 1981-2010 predictions

I will also de-duplicate the records because there were "many-to-many" joining matches downstream. De-duplicating removed about 17 records. I also remove the species column, which is automatically created by the predict function upstream.

```{r load in xy suitability dfs}

# global
IVR_suit_global_hist <- read_csv(
  file = file.path(mypath, "slf_global_v2", "global_wineries_1981-2010_xy_pred_suit_clamped_cloglog_mean.csv")
  ) %>%
  dplyr::rename("IVR_suit_global_hist" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  # tidy data
  dplyr::filter(!is.na(IVR_suit_global_hist))


# regional_invaded
IVR_suit_regional_invaded_hist <- read_csv(
  file = file.path(mypath, "slf_regional_invaded_v6", "regional_invaded_wineries_1981-2010_xy_pred_suit_clamped_cloglog.csv")
  ) %>%
  dplyr::rename("IVR_suit_regional_invaded_hist" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  # tidy data
  dplyr::filter(!is.na(IVR_suit_regional_invaded_hist))


# regional_native
IVR_suit_regional_native_hist <- read_csv(
  file = file.path(mypath, "slf_regional_native_v2", "regional_native_wineries_1981-2010_xy_pred_suit_clamped_cloglog.csv")
  ) %>%
  dplyr::rename("IVR_suit_regional_native_hist" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  # tidy data
  dplyr::filter(!is.na(IVR_suit_regional_native_hist))

```

I will also re-scale these suitability values based on the MTSS threshold.

```{r re-scale based on MTSS thresh for visualization}


# global
IVR_suit_global_hist_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_global_hist,
  thresh = "MTSS",
  summary.file = summary_global,
  rescale.name = "IVR_suit_global_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_global_hist_rescaled_thresholds <- IVR_suit_global_hist_rescaled[[2]]
IVR_suit_global_hist_rescaled <- IVR_suit_global_hist_rescaled[[1]]



# regional_invaded
IVR_suit_regional_invaded_hist_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_regional_invaded_hist,
  thresh = "MTSS",
  summary.file = summary_regional_invaded,
  rescale.name = "IVR_suit_regional_invaded_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_regional_invaded_hist_rescaled_thresholds <- IVR_suit_regional_invaded_hist_rescaled[[2]]
IVR_suit_regional_invaded_hist_rescaled <- IVR_suit_regional_invaded_hist_rescaled[[1]]




# regional_native
IVR_suit_regional_native_hist_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_regional_native_hist,
  thresh = "MTSS",
  summary.file = summary_regional_native,
  rescale.name = "IVR_suit_regional_native_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_regional_native_hist_rescaled_thresholds <- IVR_suit_regional_native_hist_rescaled[[2]]
IVR_suit_regional_native_hist_rescaled <- IVR_suit_regional_native_hist_rescaled[[1]]


```

## regional_invaded vs global

```{r slf_global_invaded plot style}

scatter_style_global_invaded <- list(
   # aesthetics
   labs(title = "Current suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_invaded' models | 1981-2010",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ),
   xlab("'global' model suitability"),
   theme_bw(),
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[30, 2]), linetype = "dotted", alpha = 0.5),
   # 10 percentile threshold
   geom_vline(xintercept = as.numeric(summary_global[34, ncol(summary_global)]), linetype = "twodash", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[34, 2]), linetype = "twodash", alpha = 0.5),
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed", alpha = 0.5)
)

```

```{r join datasets}

# join datasets for plotting
# default output
IVR_invaded_global_joined <- full_join(IVR_suit_regional_invaded_hist, IVR_suit_global_hist, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)
# rescaled output
IVR_invaded_global_joined_rescaled <- full_join(IVR_suit_regional_invaded_hist_rescaled, IVR_suit_global_hist_rescaled, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)

```

### 0-1 plot

First, I will plot and save the xy suitability on a 0-1 scale for the axes.

```{r plot SLF suitability values}

# plot
(IVR_invaded_global_01scale_plot <- ggplot() +
   # default plot style
   scatter_style_global_invaded +
   # midpoint line
   geom_vline(xintercept = 0.5) + 
   geom_hline(yintercept = 0.5) +
   # data
   geom_point(data = IVR_invaded_global_joined, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_invaded_hist), 
               color = "purple3", size = 1
               ) +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'invaded_regional' model suitability", limits = c(0, 1), breaks = breaks) 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_01scale_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_01scale_global_invaded_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## regional_native vs global

```{r IVR_global_native plot style}

scatter_style_global_native <- list(
  # aesthetics
   labs(title = "Current suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_native' models | 1981-2010",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ),
   xlab("'global' model suitability"),
   theme_bw(),
  # MTSS threshold
  geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", alpha = 0.5),
  geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed", alpha = 0.5)
  
)

```

```{r join datasets}

# join datasets for plotting
# default output
IVR_native_global_joined <- full_join(IVR_suit_regional_native_hist, IVR_suit_global_hist, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)
# rescaled output
IVR_native_global_joined_rescaled <- full_join(IVR_suit_regional_native_hist_rescaled, IVR_suit_global_hist_rescaled, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)

```

### 0-1 plot

First, I will plot and save the xy suitability on a 0-1 scale for the axes.

```{r plot SLF suitability values}

# plot
(IVR_native_global_01scale_plot <- ggplot() +
   # default plot style
   scatter_style_global_native +
   # midpoint line
   geom_vline(xintercept = 0.5) + 
   geom_hline(yintercept = 0.5) +
   # data
   geom_point(data = IVR_native_global_joined, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist), 
               color = "purple3", size = 1
               ) +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'regional_native' model suitability", limits = c(0, 1), breaks = breaks) 
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_01scale_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_01scale_global_native_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

#### rescaled by regional model

Now, I will re-scale the plot to have the MTSS threshold as the center line (currently 0.5), indicating suitability.

```{r plot SLF suitability values}

(IVR_native_global_MTSS_plot_y <- ggplot() +
   # default plot style
   scatter_style_global_native +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_joined, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist), 
               color = "purple3", size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_MTSS_plot_y, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_global_native_regional_1981-2010.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### rescaled by global model

```{r plot SLF suitability values}

(IVR_native_global_MTSS_plot_x <- ggplot()  +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # default plot style
   scatter_style_global_native +
   # data
   geom_point(data = IVR_native_global_joined, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist), 
               color = "purple3", size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_native' model suitability")
)

```

```{r save scatterplot}

if (FALSE) {

  ggsave(
    IVR_native_global_MTSS_plot_x, 
    filename = file.path(here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_global_native_regional_1981-2010.jpg"),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_MTSS_plot_x_edited <- IVR_native_global_MTSS_plot_x +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_MTSS_plot_y_edited <- IVR_native_global_MTSS_plot_y +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_MTSS_plot_combined <- IVR_native_global_MTSS_plot_x_edited +
    IVR_native_global_MTSS_plot_y_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Current suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 1981-2010",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
    )
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_MTSS_plot_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_global_native_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

### by country

I will plot the same points by country.

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_native_global_countries <- IVR_native_global_joined %>%
  inner_join(., IVR_locations, by = c("x", "y")) %>%
  dplyr::select(x:Country)

```

#### rescaled by global model

```{r plot SLF suitability values}

(IVR_native_global_countries_plot_x <- ggplot() +
   # default plot style
   scatter_style_global_native +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_countries, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_native' model suitability")
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_countries_plot_x, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_byCountry_global_native_regional_1981-2010.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### re-scaled by regional model

```{r plot SLF suitability values}

(IVR_native_global_countries_plot_y <- ggplot() +
    # default plot style
   scatter_style_global_native +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_countries, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_countries_plot_y, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_byCountry_global_native_regional_1981-2010.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_countries_plot_x_edited <- IVR_native_global_countries_plot_x +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    legend.position = "none"
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_countries_plot_y_edited <- IVR_native_global_countries_plot_y +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    axis.title.y = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_countries_plot_combined <- IVR_native_global_countries_plot_x_edited +
    IVR_native_global_countries_plot_y_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Current suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 1981-2010",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
      ) +
    plot_layout(guides = "collect") +
    theme(legend.position = "bottom")
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_countries_plot_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_byCountry_global_native_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

### points labeled by range

#### rescaled by global model

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_native_global_range <- IVR_native_global_countries %>%
  dplyr::filter(Country %in% c("China", "United States")) %>%
  rename("range" = "Country")

# replace values
IVR_native_global_range$range[IVR_native_global_range$range == "China"] <- "native"
IVR_native_global_range$range[IVR_native_global_range$range == "United States"] <- "invaded"

```

```{r plot SLF suitability values}

(IVR_native_global_range_plot_x <- ggplot() +
   # default plot style
   scatter_style_global_native +
   # make lines more prominent
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(
     data = IVR_native_global_range, 
     aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = range), 
     size = 1
     ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_range_plot_x, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_ranges_global_native_regional_1981-2010.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )

}

```

#### re-scaled by regional model

```{r plot SLF suitability values}

(IVR_native_global_range_plot_y <- ggplot()  +
    # default plot style
   scatter_style_global_native +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_range, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = range), 
               size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability")
)

```

```{r save scatterplot}

if(FALSE) {

  ggsave(
    IVR_native_global_range_plot_y, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_ranges_global_native_regional_1981-2010.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_range_plot_x_edited <- IVR_native_global_range_plot_x +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    legend.position = "none"
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_range_plot_y_edited <- IVR_native_global_range_plot_y +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    axis.title.y = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_range_plot_combined <- IVR_native_global_range_plot_x_edited +
    IVR_native_global_range_plot_y_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Current suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 1981-2010",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
      ) +
    plot_layout(guides = "collect") +
    theme(legend.position = "bottom")
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_range_plot_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_ranges_global_native_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

```{r remove objects for next analysis}

rm(scatter_style_global_invaded)
rm(scatter_style_global_native)
rm(list = ls(pattern = "_plot"))

```


# 2041-2070 predictions- GFDL SSP370

I will also not re-set the working directory

I do not need to re-load the summary files for each model or the wineries dataset

I will also de-duplicate the records because there were "many-to-many" joining matches downstream. De-duplicating removed about 17 records. I also remove the species column, which is automatically created by the predict function upstream.

```{r load in xy suitability dfs}

# global
IVR_suit_global_GFDL_ssp370 <- read_csv(
  file = file.path(mypath, "slf_global_v2", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_clamped_cloglog_mean.csv")
  ) %>%
  dplyr::rename("IVR_suit_global_GFDL_ssp370" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  # tidy data
  dplyr::filter(!is.na(IVR_suit_global_GFDL_ssp370))


# regional_invaded
IVR_suit_regional_invaded_GFDL_ssp370 <- read_csv(
  file = file.path(mypath, "slf_regional_invaded_v4", "regional_invaded_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_clamped_cloglog.csv")
  ) %>%
  dplyr::rename("IVR_suit_regional_invaded_GFDL_ssp370" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  # tidy data
  dplyr::filter(!is.na(IVR_suit_regional_invaded_GFDL_ssp370))


# regional_native
IVR_suit_regional_native_GFDL_ssp370 <- read_csv(
  file = file.path(mypath, "slf_regional_native_v1", "regional_native_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_clamped_cloglog.csv")
  ) %>%
  dplyr::rename("IVR_suit_regional_native_GFDL_ssp370" = "cloglog_suitability") %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.99) %>%
  dplyr::filter(!is.na(IVR_suit_regional_native_GFDL_ssp370))

```

I will also re-scale these suitability values based on the MTSS threshold.

```{r re-scale based on MTSS thresh for visualization}


# global
IVR_suit_global_GFDL_ssp370_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_global_GFDL_ssp370,
  thresh = "MTSS",
  summary.file = summary_global,
  rescale.name = "IVR_suit_global_GFDL_ssp370",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_global_GFDL_ssp370_rescaled_thresholds <- IVR_suit_global_GFDL_ssp370_rescaled[[2]]
IVR_suit_global_GFDL_ssp370_rescaled <- IVR_suit_global_GFDL_ssp370_rescaled[[1]]



# regional_invaded
IVR_suit_regional_invaded_GFDL_ssp370_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_regional_invaded_GFDL_ssp370,
  thresh = "MTSS",
  summary.file = summary_regional_invaded,
  rescale.name = "IVR_suit_regional_invaded_GFDL_ssp370",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_regional_invaded_GFDL_ssp370_rescaled_thresholds <- IVR_suit_regional_invaded_GFDL_ssp370_rescaled[[2]]
IVR_suit_regional_invaded_GFDL_ssp370_rescaled <- IVR_suit_regional_invaded_GFDL_ssp370_rescaled[[1]]




# regional_native
IVR_suit_regional_native_GFDL_ssp370_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = IVR_suit_regional_native_GFDL_ssp370,
  thresh = "MTSS",
  summary.file = summary_regional_native,
  rescale.name = "IVR_suit_regional_native_GFDL_ssp370",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
IVR_suit_regional_native_GFDL_ssp370_rescaled_thresholds <- IVR_suit_regional_native_GFDL_ssp370_rescaled[[2]]
IVR_suit_regional_native_GFDL_ssp370_rescaled <- IVR_suit_regional_native_GFDL_ssp370_rescaled[[1]]


```

## regional_invaded vs global- GFDL ssp370

```{r slf_global_invaded plot style}

scatter_style_global_invaded_GFDL_ssp370 <- list(
   # aesthetics
   labs(title = "Future suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_invaded' models | 2041-2070 | GFDL ssp3-70",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ),
   xlab("'global' model suitability"),
   theme_bw(),
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[30, 2]), linetype = "dotted", alpha = 0.5),
   # 10 percentile threshold
   geom_vline(xintercept = as.numeric(summary_global[34, ncol(summary_global)]), linetype = "twodash", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[34, 2]), linetype = "twodash", alpha = 0.5),
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", alpha = 0.5),
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed", alpha = 0.5)
)

```

```{r join datasets}

# join datasets for plotting
# default output
IVR_invaded_global_joined_GFDL_ssp370 <- full_join(IVR_suit_regional_invaded_GFDL_ssp370, IVR_suit_global_GFDL_ssp370, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)
# rescaled output
IVR_invaded_global_joined_GFDL_ssp370_rescaled <- full_join(IVR_suit_regional_invaded_GFDL_ssp370_rescaled, IVR_suit_global_GFDL_ssp370_rescaled, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)

```

### 0-1 plot

First, I will plot and save the xy suitability on a 0-1 scale for the axes.

```{r plot SLF suitability values}

# plot
(IVR_invaded_global_01scale_plot_GFDL_ssp370 <- ggplot() +
      # default plot style
   scatter_style_global_invaded_GFDL_ssp370 +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'invaded_regional' model suitability", limits = c(0, 1), breaks = breaks) +
   # midpoint line
   geom_vline(xintercept = 0.5) + 
   geom_hline(yintercept = 0.5) +
   # data
   geom_point(data = IVR_invaded_global_joined_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_invaded_GFDL_ssp370), 
               color = "orchid3", size = 1
               ) 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_01scale_plot_GFDL_ssp370, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_01scale_global_invaded_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```



## regional_native vs global- GFDL ssp370

```{r IVR_global_native plot style}

scatter_style_global_native_GFDL_ssp370 <- list(
  # aesthetics
   labs(title = "Future suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_native' models | 2041-2070 | GFDL ssp3-70",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ),
   xlab("'global' model suitability"),
   theme_bw(),
  # MTP threshold
  geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5),
  geom_hline(yintercept = as.numeric(summary_regional_native[30, 2]), linetype = "dotted", alpha = 0.5),
  # 10 percentile threshold
  geom_vline(xintercept = as.numeric(summary_global[34, ncol(summary_global)]), linetype = "twodash", alpha = 0.5),
  geom_hline(yintercept = as.numeric(summary_regional_native[34, 2]), linetype = "twodash", alpha = 0.5),
  # MTSS threshold
  geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", alpha = 0.5),
  geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed", alpha = 0.5)
  
)

```

```{r join datasets}

# join datasets for plotting
IVR_native_global_joined_GFDL_ssp370 <- full_join(IVR_suit_regional_native_GFDL_ssp370, IVR_suit_global_GFDL_ssp370, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)
# rescaled output
IVR_native_global_joined_GFDL_ssp370_rescaled <- full_join(IVR_suit_regional_native_GFDL_ssp370_rescaled, IVR_suit_global_GFDL_ssp370_rescaled, by = c("Species", "x", "y")) %>%
  dplyr::select(-Species)

```

### 0-1 plot

First, I will plot and save the xy suitability on a 0-1 scale for the axes.

```{r plot SLF suitability values}

# plot
(IVR_native_global_01scale_plot_GFDL_ssp370 <- ggplot()  +
   # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # midpoint line
   geom_vline(xintercept = 0.5) + 
   geom_hline(yintercept = 0.5) +
   # data
   geom_point(data = IVR_native_global_joined_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370), 
               color = "orchid3", size = 1
               ) +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'regional_native' model suitability", limits = c(0, 1), breaks = breaks)
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_01scale_plot_GFDL_ssp370, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_01scale_global_native_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

#### rescaled by regional model

Now, I will re-scale the plot to have the MTSS threshold as the center line (currently 0.5), indicating suitability.

```{r plot SLF suitability values}

(IVR_native_global_MTSS_plot_y_GFDL_ssp370 <- ggplot() +
   # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_joined_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370), 
               color = "orchid3", size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
   
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_MTSS_plot_y_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_global_native_regional_2041-2070_GFDL_ssp370.jpg"
                         ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### rescaled by global model

```{r plot SLF suitability values}

(IVR_native_global_MTSS_plot_x_GFDL_ssp370 <- ggplot() +
   # default plot style
   scatter_style_global_native_GFDL_ssp370 +
    # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_joined_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370), 
               color = "orchid3", size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {

  ggsave(
    IVR_native_global_MTSS_plot_x_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_global_native_regional_2041-2070_GFDL_ssp370.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_MTSS_plot_x_GFDL_ssp370_edited <- IVR_native_global_MTSS_plot_x_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_MTSS_plot_y_GFDL_ssp370_edited <- IVR_native_global_MTSS_plot_y_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_MTSS_plot_GFDL_ssp370_combined <- IVR_native_global_MTSS_plot_x_GFDL_ssp370_edited +
    IVR_native_global_MTSS_plot_y_GFDL_ssp370_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Future suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 2041-2070 | GFDL ssp3-70",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
    )
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_MTSS_plot_GFDL_ssp370_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_global_native_regional_2041-2070_GFDL_ssp370.jpg"
  ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

### by country

I will plot the same points by country.

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_native_global_countries_GFDL_ssp370 <- IVR_native_global_joined_GFDL_ssp370 %>%
  inner_join(., IVR_locations, by = c("x", "y")) %>%
  dplyr::select(x:Country)

```

#### rescaled by global model

```{r plot SLF suitability values}

(IVR_native_global_countries_plot_x_GFDL_ssp370 <- ggplot() +
   # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_countries_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_countries_plot_x_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_byCountry_global_native_regional_2041-2070_GFDL_ssp370.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### re-scaled by regional model

```{r plot SLF suitability values}

(IVR_native_global_countries_plot_y_GFDL_ssp370 <- ggplot() +
      # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_countries_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_countries_plot_y_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_byCountry_global_native_regional_2041-2070_GFDL_ssp370.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_countries_plot_x_GFDL_ssp370_edited <- IVR_native_global_countries_plot_x_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    legend.position = "none"
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_countries_plot_y_GFDL_ssp370_edited <- IVR_native_global_countries_plot_y_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    axis.title.y = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_countries_plot_GFDL_ssp370_combined <- IVR_native_global_countries_plot_x_GFDL_ssp370_edited +
    IVR_native_global_countries_plot_y_GFDL_ssp370_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Future suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 2041-2070 | GFDL ssp3-70",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
      ) +
    plot_layout(guides = "collect") +
    theme(legend.position = "bottom")
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_countries_plot_GFDL_ssp370_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_byCountry_global_native_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

### points labeled by range

#### rescaled by global model

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_native_global_range_GFDL_ssp370 <- IVR_native_global_countries_GFDL_ssp370 %>%
  dplyr::filter(Country %in% c("China", "United States")) %>%
  rename("range" = "Country")

# replace values
IVR_native_global_range_GFDL_ssp370$range[IVR_native_global_range_GFDL_ssp370$range == "China"] <- "native"
IVR_native_global_range_GFDL_ssp370$range[IVR_native_global_range_GFDL_ssp370$range == "United States"] <- "invaded"

```

```{r plot SLF suitability values}

(IVR_native_global_range_plot_x_GFDL_ssp370 <- ggplot() +
   # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # make lines more prominent
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(
     data = IVR_native_global_range_GFDL_ssp370, 
     aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = range), 
     size = 1
     ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if (FALSE) {
  
  ggsave(
    IVR_native_global_range_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_ranges_global_native_regional_2041-2070_GFDL_ssp370.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )

}

```

#### re-scaled by regional model

```{r plot SLF suitability values}

(IVR_native_global_range_plot_y_GFDL_ssp370 <- ggplot() +
      # default plot style
   scatter_style_global_native_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_native_global_range_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = range), 
               size = 1
               ) +
   # scaling for axes
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   ylab("'regional_native' model suitability") 
)

```

```{r save scatterplot}

if(FALSE) {

  ggsave(
    IVR_native_global_range_plot_y_GFDL_ssp370, 
    filename = file.path(
      here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_y_scaled_ranges_global_native_regional_2041-2070_GFDL_ssp370.jpg"
      ),
    height = 8, 
    width = 10,
    device = "jpeg",
    dpi = "retina"
    )
  
}

```

#### both axes

```{r patchwork both axis plots}

# edit x
IVR_native_global_range_plot_x_GFDL_ssp370_edited <- IVR_native_global_range_plot_x_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    legend.position = "none"
    ) +
  guides(x =  guide_axis(angle = 45))

# edit y
IVR_native_global_range_plot_y_GFDL_ssp370_edited <- IVR_native_global_range_plot_y_GFDL_ssp370 +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    # extra edits
    axis.title.y = element_blank()
    ) +
  guides(x =  guide_axis(angle = 45))

# combined plot
(IVR_native_global_range_plot_GFDL_ssp370_combined <- IVR_native_global_range_plot_x_GFDL_ssp370_edited +
    IVR_native_global_range_plot_y_GFDL_ssp370_edited + 
    plot_layout(axis_titles = "collect") +
    plot_annotation(
      title = "Future suitability for SLF establishment in globally important viticultural areas",
      subtitle = "'global' vs 'regional_native' models | 2041-2070 | GFDL ssp3-70",
      caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
      ) +
    plot_layout(guides = "collect") +
    theme(legend.position = "bottom")
)


```

```{r save scatterplot}

ggsave(
  IVR_native_global_range_plot_GFDL_ssp370_combined, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_xy_scaled_ranges_global_native_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

```{r remove objects for next analysis}

rm(scatter_style_global_invaded_GFDL_ssp370)
rm(scatter_style_global_native_GFDL_ssp370)
rm(list = ls(pattern = "_plot"))

```

Finally, I will combine some of these plots to illustrate the shift in suitability for climate change.

# Combine Plots

## regional_invaded vs global- 0-1 scaled plot

```{r join datasets}

# join datasets for plotting
IVR_invaded_global_joined_change <- full_join(IVR_invaded_global_joined, IVR_invaded_global_joined_GFDL_ssp370, by = c("x", "y")) 

```

```{r plot SLF suitability values}

# plot
(IVR_invaded_global_01scale_change_plot <- ggplot(data = IVR_invaded_global_joined_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[30, 2]), linetype = "dotted", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist,
       xend = IVR_suit_global_GFDL_ssp370,
       y = IVR_suit_regional_invaded_hist,
       yend = IVR_suit_regional_invaded_GFDL_ssp370
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist, y = IVR_suit_regional_invaded_hist, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_invaded_GFDL_ssp370, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'invaded_regional' model suitability", limits = c(0, 1), breaks = breaks) +
   # scales style
   labs(title = "Climate change effects on predicted suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_invaded' models",
         caption = "thresholds: dashed = MTSS, dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 1, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "bottom") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_01scale_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_global_invaded_regional.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## regional_invaded vs global- MTSS rescaled plot

```{r join datasets}

# join datasets for plotting
IVR_invaded_global_joined_rescaled_change <- full_join(IVR_invaded_global_joined_rescaled, IVR_invaded_global_joined_GFDL_ssp370_rescaled, by = c("x", "y")) 

```

```{r plot SLF suitability values}

# plot
(IVR_invaded_global_MTSS_rescaled_change_plot <- ggplot(data = IVR_invaded_global_joined_rescaled_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(IVR_suit_global_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(IVR_suit_regional_invaded_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(IVR_suit_global_hist_rescaled_thresholds[5, 2]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(IVR_suit_regional_invaded_hist_rescaled_thresholds[5, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist_rescaled,
       xend = IVR_suit_global_GFDL_ssp370_rescaled,
       y = IVR_suit_regional_invaded_hist_rescaled,
       yend = IVR_suit_regional_invaded_GFDL_ssp370_rescaled
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist_rescaled, y = IVR_suit_regional_invaded_hist_rescaled, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370_rescaled, y = IVR_suit_regional_invaded_GFDL_ssp370_rescaled, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axes scaling
   scale_x_continuous(name = "risk according to 'global' model", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "risk according to 'regional_invaded' model", limits = c(0, 1), breaks = breaks) +
   # scales style
   labs(title = "Risk of SLF establishment in globally important viticultural regions under climate change",
         subtitle = "'global' vs 'regional_invaded' models | suitability transformed based on MTSS threshold",
         caption = "thresholds: dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 1, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "bottom") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_MTSS_rescaled_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_global_invaded_regional_rescaled.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## regional_native vs global- 0-1 scaled plot

```{r join datasets}

# join datasets for plotting
IVR_native_global_joined_change <- full_join(IVR_native_global_joined, IVR_native_global_joined_GFDL_ssp370, by = c("x", "y")) 

```

```{r plot SLF suitability values}

# plot
(IVR_native_global_01scale_change_plot <- ggplot(data = IVR_native_global_joined_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_native[30, 2]), linetype = "dotted", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist,
       xend = IVR_suit_global_GFDL_ssp370,
       y = IVR_suit_regional_native_hist,
       yend = IVR_suit_regional_native_GFDL_ssp370
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "'regional_native' model suitability", limits = c(0, 1), breaks = breaks) +
   # scales style
   labs(title = "Climate change effects on suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_native' models",
         caption = "thresholds: dashed = MTSS, dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 1, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "bottom") 
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_01scale_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_global_native_regional.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

```{r join datasets}

# join datasets for plotting
IVR_native_global_joined_change <- full_join(IVR_native_global_joined, IVR_native_global_joined_GFDL_ssp370, by = c("x", "y")) 

```

# regional_native vs global- MTSS rescaled plot

```{r join datasets}

# join datasets for plotting
IVR_native_global_joined_rescaled_change <- full_join(IVR_native_global_joined_rescaled, IVR_native_global_joined_GFDL_ssp370_rescaled, by = c("x", "y")) 

```

```{r plot SLF suitability values}

# plot
(IVR_native_global_MTSS_rescale_change_plot <- ggplot(data = IVR_native_global_joined_rescaled_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(IVR_suit_global_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(IVR_suit_regional_native_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(IVR_suit_global_hist_rescaled_thresholds[5, 2]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(IVR_suit_regional_native_hist_rescaled_thresholds[5, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist_rescaled,
       xend = IVR_suit_global_GFDL_ssp370_rescaled,
       y = IVR_suit_regional_native_hist_rescaled,
       yend = IVR_suit_regional_native_GFDL_ssp370_rescaled
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist_rescaled, y = IVR_suit_regional_native_hist_rescaled, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370_rescaled, y = IVR_suit_regional_native_GFDL_ssp370_rescaled, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axes scaling
   scale_x_continuous(name = "risk according to 'global' model", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name = "risk according to 'regional_native' model", limits = c(0, 1), breaks = breaks) +
   # scales style
   labs(title = "Risk of SLF establishment in globally important viticultural regions under climate change",
         subtitle = "'global' vs 'regional_native' models | suitability transformed based on MTSS threshold",
         caption = "thresholds: dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 1, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "bottom") 
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_MTSS_rescale_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_global_native_regional_rescaled.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## regional_native vs global- x-scaled plot

```{r plot SLF suitability values}

# plot
(IVR_native_global_x_scaled_change_plot <- ggplot(data = IVR_native_global_joined_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_native[30, 2]), linetype = "dotted", alpha = 0.5) +
   # 10 percentile threshold
   geom_vline(xintercept = as.numeric(summary_global[34, ncol(summary_global)]), linetype = "twodash", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_native[34, 2]), linetype = "twodash", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist,
       xend = IVR_suit_global_GFDL_ssp370,
       y = IVR_suit_regional_native_hist,
       yend = IVR_suit_regional_native_GFDL_ssp370
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axis rescaling
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) +
   xlab("'global' model suitability") + 
   ylab("'invaded_native' model suitability") +
   # scales style
   labs(title = "Climate change effects on suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_native' models",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 1, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "bottom") 
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_x_scaled_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_x_scaled_global_native_regional.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## regional_native vs global- y-scaled plot

```{r plot SLF suitability values}

# plot
(IVR_native_global_y_scaled_change_plot <- ggplot(data = IVR_native_global_joined_change) +
   # threshold lines
   # MTP threshold
   geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_native[30, 2]), linetype = "dotted", alpha = 0.5) +
   # 10 percentile threshold
   geom_vline(xintercept = as.numeric(summary_global[34, ncol(summary_global)]), linetype = "twodash", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(summary_regional_native[34, 2]), linetype = "twodash", alpha = 0.5) +
   # MTSS threshold
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_native[42, 2]), linetype = "dashed") +
   # arrows indicating change
   geom_segment(
     aes(
       x = IVR_suit_global_hist,
       xend = IVR_suit_global_GFDL_ssp370,
       y = IVR_suit_regional_native_hist,
       yend = IVR_suit_regional_native_GFDL_ssp370
     ), 
     arrow = arrow(angle = 8, type = "closed"), linewidth = 0.25, color = "azure4", alpha = 0.7
   ) +
   # historical data
   geom_point(
     aes(x = IVR_suit_global_hist, y = IVR_suit_regional_native_hist, color = "1981-2010"), 
     size = 1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_native_GFDL_ssp370, color = "2041-2070\nGFDL ssp370"), 
     size = 1
     ) +
   # axis rescaling
   scale_y_cut(breaks = as.numeric(summary_regional_native[42, 2] * 2), which = 1, scales = 0.25, space = 0.25) +
   xlab("'global' model suitability") + 
   ylab("'invaded_native' model suitability") +
   # scales style
   labs(title = "Climate change effects on suitability for SLF establishment in globally important viticultural areas",
         subtitle = "'global' vs 'regional_native' models",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ) +
   # aesthetics
   scale_color_manual(name = "Time period", values = c("purple3", "orchid1")) +
   guides(color = guide_legend(nrow = 2, override.aes = list(size = 2))) +
   theme_bw() +
   theme(legend.position = "right") 
)

```

```{r save scatterplot}

ggsave(
  IVR_native_global_y_scaled_change_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_CC_shift_y_scaled_global_native_regional.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```



# Appendix

## global vs regional_invaded models: 1981-2010

### rescaled

Now, I will re-scale the plot to have the MTSS threshold as the center line (currently 0.5), indicating suitability.

NOT INFORMATIVE

```{r plot SLF suitability values}

(IVR_invaded_global_MTSS_plot <- ggplot() +
   # default plot style
   scatter_style_global_invaded +
   # data
   geom_point(data = IVR_invaded_global_joined, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_invaded_hist), 
               color = "purple3", size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_invaded' model suitability") +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_MTSS_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_global_invaded_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

### by continent

I will plot the same points by country.

 PLOT NOT HELPFUL

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_invaded_global_countries <- IVR_invaded_global_joined %>%
  inner_join(., IVR_locations, by = c("x", "y")) %>%
  dplyr::select(x:Country)

```

```{r plot SLF suitability values}

(IVR_invaded_global_countries_plot <- ggplot()  +
   # default plot style
   scatter_style_global_invaded +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_invaded_global_countries, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_invaded_hist, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_invaded' model suitability") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_countries_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_byContinent_global_invaded_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

### native vs invaded range

NOT HELPFUL

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_invaded_global_range <- IVR_invaded_global_countries %>%
  dplyr::filter(Country %in% c("China", "United States")) %>%
  rename("range" = "Country")

# replace values
IVR_invaded_global_range$range[IVR_invaded_global_range$range == "China"] <- "native"
IVR_invaded_global_range$range[IVR_invaded_global_range$range == "United States"] <- "invaded"

```

```{r plot SLF suitability values}

(IVR_invaded_global_range_plot <- ggplot() +
   # default plot style
   scatter_style_global_invaded +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_invaded_global_range, 
               aes(x = IVR_suit_global_hist, y = IVR_suit_regional_invaded_hist, color = range), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_invaded' model suitability") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_range_plot, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_ranges_global_invaded_regional_1981-2010.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

## global vs regional_invaded models: 2041-2070, GFDL ssp370

### rescaled

Now, I will re-scale the plot to have the MTSS threshold as the center line (currently 0.5), indicating suitability.

NOT INFORMATIVE

```{r plot SLF suitability values}

(IVR_invaded_global_MTSS_plot_GFDL_ssp370 <- ggplot() +
      # default plot style
   scatter_style_global_invaded_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_invaded_global_joined_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_invaded_GFDL_ssp370), 
               color = "orchid3", size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25) + 
   ylab("'regional_invaded' model suitability") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_MTSS_plot_GFDL_ssp370, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_global_invaded_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

### by continent

I will plot the same points by country.

 PLOT NOT HELPFUL

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_invaded_global_countries_GFDL_ssp370 <- IVR_invaded_global_joined_GFDL_ssp370 %>%
  inner_join(., IVR_locations, by = c("x", "y")) %>%
  dplyr::select(x:Country)

```

```{r plot SLF suitability values}

(IVR_invaded_global_countries_plot_GFDL_ssp370 <- ggplot() +
   # default plot style
   scatter_style_global_invaded_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_invaded_global_countries_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_invaded_GFDL_ssp370, color = Continent), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_invaded' model suitability") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_countries_plot_GFDL_ssp370, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_byContinent_global_invaded_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```

### native vs invaded range

NOT HELPFUL

```{r find state names for lat/lon data}

# join geocoded data with joined suitability dataset
IVR_invaded_global_range_GFDL_ssp370 <- IVR_invaded_global_countries_GFDL_ssp370 %>%
  dplyr::filter(Country %in% c("China", "United States")) %>%
  rename("range" = "Country")

# replace values
IVR_invaded_global_range_GFDL_ssp370$range[IVR_invaded_global_range_GFDL_ssp370$range == "China"] <- "native"
IVR_invaded_global_range_GFDL_ssp370$range[IVR_invaded_global_range_GFDL_ssp370$range == "United States"] <- "invaded"

```

```{r plot SLF suitability values}

(IVR_invaded_global_range_plot_GFDL_ssp370 <- ggplot() +
      # default plot style
   scatter_style_global_invaded_GFDL_ssp370 +
   # threshold lines
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed") +
   geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed") +
   # data
   geom_point(data = IVR_invaded_global_range_GFDL_ssp370, 
               aes(x = IVR_suit_global_GFDL_ssp370, y = IVR_suit_regional_invaded_GFDL_ssp370, color = range), 
               size = 1
               ) +
   # scaling for axes
   scale_x_cut(breaks = as.numeric(summary_global[42, ncol(summary_global)] * 2), which = 2, scales = 0.25, space = 0.25) +
   ylab("'regional_invaded' model suitability") 
)

```

```{r save scatterplot}

ggsave(
  IVR_invaded_global_range_plot_GFDL_ssp370, 
  filename = file.path(
    here(), "vignette-outputs", "figures", "xy_suitability_IVR_MTSS_x_scaled_ranges_global_invaded_regional_2041-2070_GFDL_ssp370.jpg"
    ),
  height = 8, 
  width = 10,
  device = "jpeg",
  dpi = "retina"
  )

```



# References

Gallien, L., R. Douzet, S. Pratte, N. E. Zimmermann, and W. Thuiller. 2012. Invasive species distribution models  how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography 21:11261136.

S Xu#, M Chen#, T Feng, L Zhan, L Zhou, G Yu*. Use ggbreak to effectively utilize plotting space to deal with large datasets and outliers. Frontiers in Genetics. 2021, 12:774846. doi: 10.3389/fgene.2021.774846
