---
title: "jumps_leaflet_mapping_v2"
author: "Sam Owens"
date: '2022-03-27'
contact: 'sam.owens@temple.edu'
output: html_document
---

This .rmd builds on the previous version of this file, where I mapped a single point for EDA purposes. I tried to wrap that mapped point into a function to map all points in a data frame. This improved and simplified build of the function map_jumps will allow leaflet maps of SLF jumps to be stored in the same data frame as the original points. The tinyslf, which is the dataset for lycormap and jump_locations datasets will be used in this vignette, but any data frame with standardized variables may be used. 

# Load necessary packages, datasets to be used

```{r load necessary packages, eval=FALSE}

install.packages(c("tidyverse", "here", "leaflet"))

library(tidyverse)
library(here)
library(leaflet)

library(lycormap)
# this package is only available to iEco lab members at the time this script was written- it cannot be downloaded from the github CRAN. Please contact a senior member of the lab for access.

```

```{r load necessary datasets, echo=FALSE}

tinyslf <- tinyslf %>%
  filter(slf_present == "TRUE")

jumps <- read_csv(
  file.path(here(), "data_root/slfSpread_jumps_locations_SMO_v2.csv")
)

```

# Create custom Icon for Mapping

``` {r create marker object, echo = FALSE}

SLF_marker <- makeAwesomeIcon(
  icon = "bug",
  iconColor = "black",
  markerColor = "orange",
  library = "fa"
)

```

# Simplified version of map_jumps function

``` {r map_jumps function- simplified}

map_jumps_v2 <- function(point_long, point_lat, point_label, extent = 0.01, zoom = 10) {

  # point_long and point_lat represent the longitude and latitude of the point you want to map
  # point_label is the label you gave to that point 
  # the variable "extent" is the width and height on either side of each point that you want mapped, in degrees latitude or longitude
  # zoom allows you to set the zoom level, if you want tot have a different one- even if you choose to use fitBounds() to zoom in on the point, a zoom level must still be established within setView(). fitBounds() would then override the zoom level

  leaflet() %>%
    setView(lng = point_long,
            lat = point_lat,
            zoom = 10
    ) %>%
    fitBounds(lng1 = point_long + extent,
              lng2 = point_long - extent,
              lat1 = point_lat + extent,
              lat2 = point_lat - extent
    ) %>%
    addTiles() %>%
    addAwesomeMarkers(lng = point_long,
                      lat = point_lat,
                      icon = SLF_marker,
                      popup = point_label
    ) %>%
    addScaleBar(position = "bottomleft") %>%
    
    return()
}

```

# Examples of how to use map_jumps_v2

Simply defining each variable as a row in a particular tibble/data frame (in this case a tibble is used) will produce a map, as shown below

```{r Mapping a single point, from jumps data table}

jump_1_mapped <- map_jumps_v2(
  point_long = jumps$longitude_rounded[1],
  point_lat = jumps$latitude_rounded[1],
  point_label = jumps$ref_ID[1],
  extent = 0.01,
  zoom = 10
)

```


```{r Mapping of all coordinate pairs and pairing adding it to the dataset}

jump_locations_with_maps <- jumps %>%
  mutate(mapped_jumps = purrr::pmap(
    list(longitude_rounded, latitude_rounded, ref_ID), 
    map_jumps_v2 
  ))

jump_locations_with_maps$mapped_jumps[[41]]
# this executes the list stored in row 2 of the jump_locations_with_maps tibble

```


