---
title: "maxent_workflow"
author: "Samuel M. Owens"
date: "2023-09-28"
output: html_document
---

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava)

library(ENMeval) # other useful env functions
library(dismo) # package underneath SDMtune
library(SDMtune) # main package used to run SDMs

library(kableExtra) # required for producing model reports
library(plotROC) # plots ROCs

library(sf) # spatial data 
library(raster) # spatial data 
library(terra) # spatial data 

```


```{r ggplot object for map style}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
        ),
  coord_equal() 
)

```

`Dismo` will run Maxent through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

```{r}

checkMaxentInstallation(verbose = TRUE)

```

This chunk sets the java memory allocation (`Xmx`). I will start with 2GB memory. 

```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = "-Xmx2048m")

# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

# Setup for MaxEnt

- I may need to convert my slf points to a spatial points object

- IF I need to extract point-wise climatic data for MaxEnt, use these steps.

1. stack env covariates 
2. use slf pointwise data to extract a value from each layer at that each point

THIS IS A GOOD VALIDATION STEP- can use is.na()

First, I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used for in the `maxent()` function. 

```{r load in files for all maxent models}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")

# environmental and human impact covariates
# pull those of the proper ending
x_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = '\\_easternUSA.asc$', full.names = TRUE)



# slf presences
p_slf_points <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "slf_all_final_coords_v1_2023-08.csv")) %>%
  dplyr::select(-species)

# background point sets
# entire eastern USA
a_entire_easternUSA_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_flexible_area_points.csv"))
# DD model used with 1% adult emergence threshold
a_adult_emergence_1_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emergence-1_flexible_area_points.csv")) 
# 355km buffer around slf points
a_slf_buffered_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_points_buffered_flexible_area_points.csv"))

```

I will load in the environmental covariates and stack them.

```{r stack env covariates}
 
# use in terra package first because stats are easier
x_env_covariates <- terra::rast(x = x_env_covariates_list)

# attributes
nlyr(x_env_covariates)
names(x_env_covariates)
minmax(x_env_covariates)
# ext(x_env_covariates)
# crs(x_env_covariates)

# I will change the name of the variables because they are throwing errors in SDMtune
names(x_env_covariates) <- c("atc_2015", "bio11_1981_2010", "bio12_1981_2010", "bio15_1981_2010", "bio2_1981_2010")

# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)

```

I will also reduce slf presences to just those within the eastern USA.

```{r}

# extent object for eastern USA
ext.obj <- terra::ext(-96.503906, -59.589844, 28.304381, 47.457809)

# conert to vector
p_slf_points_vect <- terra::vect(x = p_slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert back to data frame
p_slf_points <- terra::as.data.frame(p_slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole))

```

# Maxent Model- entire eastern USA

## create input data object

First, I will create an SWD (sample with data) object containing the presences and background points for this model

```{r prepare SWD data object}

entire_easternUSA_SWD <- SDMtune::prepareSWD(species = "Lycorma delicatula",
                                             env = x_env_covariates,
                                             p = p_slf_points,
                                             a = a_entire_easternUSA_background_points,
                                             verbose = TRUE # print helpful messages
                                             )

entire_easternUSA_SWD@coords # coordinates
entire_easternUSA_SWD@pa # presence / absence (background counted as absence)
entire_easternUSA_SWD@data # extracted data from 

```

```{r save SWD file}

SDMtune::swd2csv(swd = entire_easternUSA_SWD, file_name = c(
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_presences_with_data.csv"),
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_absences_with_data.csv")
  ))

```

## Create training / test split

I will split the presences into training and testing, using 80% of the points for training and 20% for testing. I will then use `SDMtune::randomFold()` to split the training data into 5 partitions for cross-validation. This method was loosely adapted from Srivastava et.al, 2021. 

```{r split data for training and testing}

set.seed(4)

entire_easternUSA_trainTest <-  SDMtune::trainValTest(
  x = entire_easternUSA_SWD,
  test = 0.2,
  only_presence = TRUE
)

# separate off training data
entire_easternUSA_train <- entire_easternUSA_trainTest[[1]]
entire_easternUSA_test <- entire_easternUSA_trainTest[[2]]

entire_easternUSA_train@coords # coordinates
entire_easternUSA_train@pa # presence / absence (background counted as absence)
entire_easternUSA_train@data # extracted data from

```

```{r save training and test datasets as csv}

# save training data
SDMtune::swd2csv(entire_easternUSA_train, 
                 file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_model_training.csv"))
# save testing data
SDMtune::swd2csv(entire_easternUSA_test, 
                 file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_model_testing.csv"))

```

```{r load data and split data into k folds}



```



## Train Maxent model

First, I will train a model with 5 cross-validated replicates. The model will only be trained on 80% of the slf presence data, with the other 20% being used downstream for analysis. The default settings will be used for Maxent, apart from the following changes:

* ALL feature types used (l = linear, q = quadratic, p = product, h = hinge, t = threshold)
* replicates = 5
* iterations = 5000. This is the max number of iterations of optimization algorithm to perform before stopping training. Increasing this number allows the algorithm to make more refined predictions.

Other changes to the default will be used, but are not relevant for training the model.

## Maxent Using dismo pkg

Use `if(dismo::maxent()) { }` to test whether maxent will run on data given to it before actually running the models.

*Note* I prefer to do spatial data manipulation in `terra`, but `dismo` is only compatible with the `raster` package.

```{r object of maxent settings}

maxent_settings <- c(
  "linear=true", # all feature types allowed
  "quadratic=true", 
  "product=true",
  "threshold=true",
  "hinge=true",
  "responsecurves=true", # create response curves
  "jackknife=true", # do jackknife analysis
  "replicates=5", # do 5 replicates for run (k-fold)
  "replicatetype=crossvalidate", # k-fold cross validation
  "randomtestpoints=20", # randomly set aside 20% of presences for testing
  "maximumiterations=5000", # max number of iterations of optimization algorithm to perform before stopping training
  "threads=6", # number of physical cores in my desktop
  "nodata=-9999",# no data value
  "writeplotdata=true" 
) 

```

```{r train maxent model}

# load training data
entire_easternUSA_train <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_model_training.csv"))

entire_easternUSA_model <- SDMtune::train(
  method = "Maxent",
  data = entire_easternUSA_train,
  folds = entire_easternUSA_trainFolds, # 5 folds for dataset
  fc = "lqpht", # feature classes set to ALL
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Evaluate maxent model

I will use various metrics to evaluate the model, including 

I will generate a report, including:
* ROC plot
* marginal response curves
* univariate curves
* jackknife plot
* variable importance
* thresholds
* model settings
* training data points
* test data points
* RDS file of model

```{r}

# load testing data
entire_easternUSA_test <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_model_testing.csv"))

```

