---
title: "Initialization of renv package for dependencies"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2024-10-16"
---

In this file, I will list the dependency packages for scarifSDM and initialize the `renv` package to keep my other package versions consistent. End users should only run sections 1 and 2. Section 3 is for troubleshooting common issues. Section 4 should only be used for the first-time initialization of the package.

# Get package dependencies

```{r set global chunk opts, include = FALSE}

library(knitr)

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

```{r library packages}

library(tidyverse)
library(renv)
library(here)
library(devtools)

```

Run this chunk to see the dependencies for the scarifSDM package.

```{r get dependencies list}

packages <- renv::dependencies() %>%
  # select unique package names
  dplyr::summarize(package = sort(unique(Package))) %>%
  as.matrix() %>%
  as.character()

```

# Retrieve correct dependencies for scarifSDM

```{r check status}

# first, check the status of your current package versions vs those in renv
renv::status()

```

```{r restore as necessary}

# you should run this to ensure the correct package versions are used while you are working within this project. This aligns the sys library with the lockfile.
renv::restore()
# only use this version if you want to edit your system library to match the renv library
renv::restore(clean = TRUE)

```


# Troubleshooting

The first thing you should try is updating your R and Rtools versions. This messed up my renv for quite awhile.

```{r troubleshooting}

# hydrate() corrected an issue where renv could not discover any local packages in the system library
renv::hydrate()

# install specific packages (usually needed for packages not on the CRAN)
# you may need to enter your github credentials into the .Renviron or log into github for this to work
# first, create a github PAT on website
usethis::create_github_token()
# enter created PAT (personal access token) into the .Renviron
## be sure to use defaults and allow download and upload of packages
gitcreds::gitcreds_set()
# now install packages not on the CRAN
devtools::install_github("densitymodelling/dsmextra") # dsmextra
devtools::install_github("ieco-lab/lydemapr") # lydemapr
devtools::install_github("jasonleebrown/humboldt") # humboldt
devtools::install_github("ropensci/rnaturalearthhires")# rnaturalearthhires

# renv restore to previous package versions
# only use if the current package version need to be reverted to a previous state
renv::history()
renv::revert(project = "scarifSDM", commit = "")

# repair common issues
renv::repair()

# update renv to latest version in this project (but this will differ from the source download of this project folder)
renv::upgrade()

# run this to rebuild and reinstall all packages in the library- this starts them from a clean slate
renv::rebuild()

```

# one-time renv initialization

I created this section to initialize the renv package for the first time ONLY. DO NOT run this section if you have downloaded this package as an outside user.

```{r one-time initialization}

# initialization of project renv
# only run once
renv::init()

# record newest package versions in lockfile
renv::snapshot()

# update packages
# only run as needed if a package breaks or something
renv::update()
# alternatively, install the latest CRAN versions:
renv::install()

```

# remove renv

As needed, you may need to completely remove renv and restart from a fresh install.

```{r remove renv}

# deactivate renv
renv::deactivate()

# remove renv from project
renv::deactivate(clean = TRUE)

# completely remove renv from the system
root <- renv::paths$root()
unlink(root, recursive = TRUE)
utils::remove.packages("renv")

```

