---
title: "Initialization of renv package for dependencies"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2025-07-04"
---

In this file, I will list the dependency packages for scari and initialize the `renv` package to keep my other package versions consistent. End users should only run sections 1 and 2. Section 3 is for troubleshooting common issues. Section 4 should only be used for the first-time initialization of the package.

## Notes about using this package's code

Here are a few things any user should keep in mind as they run the code in this package.

I have done some custom work to ensure this package's .html files render correctly and that the code is not overly cumbersome to run. Here are a few of the conventions I have used:

- I have added a number of `if()` statements throughout my vignettes, usually to prevent data from being downloaded again unnecessarily. Simply change the `FALSE` within the `if()` statement to `TRUE` to run a chunk. This is to prevent re-downloading.
- I have a chunk near the beginning of every vignette that sets the global chunk options: `knitr::opts_chunk$set(eval = FALSE)`. This prevents the code from running automatically when the vignette is rendered. You can change this to `TRUE` if you want to run the code in that chunk.
- I have added html comment tags (`<!-- -->`) to the code chunks that I do not want to run automatically as the website is rendered. Much of the code that is commented out is redundant, but I have left it in the code for reference so as not to disrupt the true workflow. You can remove these comments if you want to run the code in those chunks. (Basically, the rendered version of these files is usually much shorter than the R code is.).

# Get package dependencies

```{r set global chunk opts, include = FALSE}

library(knitr)

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

```{r library packages}

library(tidyverse)
library(renv)
library(here)
library(devtools)
library(gitcreds)

```

Run this chunk to see the dependencies for the scari package.

```{r get dependencies list, eval = TRUE}

packages <- renv::dependencies() %>%
  # select unique package names
  dplyr::summarize(package = sort(unique(Package))) %>%
  as.matrix() %>%
  as.character()

```

# one-time renv initialization

I created this section to initialize the renv package for the first time ONLY. DO NOT run this section if you have downloaded this package as an outside user.

```{r one-time initialization}

# ensure that developer packages are installed prior to activation
remotes::install_github("ieco-lab/lydemapr", build_vignettes = FALSE)
remotes::install_github("ropensci/rnaturalearthhires")
remotes::install_github("ropensci/rnaturalearthdata")
remotes::install_github("densitymodelling/dsmextra")
# if these continue to throw errors, you might exit the scari Rproject session and install them. Then, re-open the Rproject.

# initialization of project renv
# only run once
renv::init(project = file.path(here::here()))

# check status
renv::status()

# you may need to call this if the dependencies are used and installed, but not recorded
renv::snapshot()

```

If the above installation does not work, you may first need to have git installed, found here: https://git-scm.com/downloads. Then, open the windows command prompt and run the following code to ensure that R can find git. "git --version" should return the version of git installed on your computer. If it does not, you may need to adjust the system environmental variables by adding system environmental variables. Click on system variables > path and add the following paths:

`C:\Program Files\Git\cmd`
`C:\Program Files\Git\bin`

Windows should be able to locate git now.

If R cant find git, run this code in Rstudio:

```{r git installation}

Sys.setenv(PATH = paste("C:/Program Files/Git/cmd", Sys.getenv("PATH"), sep = ";"))  # adjust path as needed

```

# Retrieve correct dependencies for scari

```{r check status}

# first, check the status of your current package versions vs those in renv
renv::status()

```

```{r restore as necessary}

# you should run this to ensure the correct package versions are used while you are working within this project. This aligns the sys library with the lockfile.
renv::restore()
# only use this version if you want to edit your system library to match the renv library
renv::restore(clean = TRUE)

```


# Troubleshooting

The first thing you should try is updating your R and Rtools versions. This messed up my renv for quite awhile.

```{r remove and reinstall packages as necessary}

# these were the packages initially giving errors for "missing description files"
remove.packages(c("lydemapr", "rnaturalearthhires", "rnaturalearthdata", "dsmextra"))
renv::remove(c("lydemapr", "rnaturalearthhires", "rnaturalearthdata", "dsmextra"))
#reinstall
renv::install(c("lydemapr", "rnaturalearthhires", "rnaturalearthdata", "dsmextra"))

renv::status()

```

```{r troubleshooting}

# hydrate() corrected an issue where renv could not discover any local packages in the system library
renv::hydrate()

# install specific packages (usually needed for packages not on the CRAN)
# you may need to enter your github credentials into the .Renviron or log into github for this to work:
usethis::edit_r_environ()
# then, create a github PAT on website. 
usethis::create_github_token()
# enter created PAT (personal access token) into the .Renviron
## be sure to use defaults and allow download and upload of packages
gitcreds::gitcreds_set()
# now install packages not on the CRAN
devtools::install_github("densitymodelling/dsmextra") # dsmextra
devtools::install_github("ieco-lab/lydemapr") # lydemapr
devtools::install_github("ropensci/rnaturalearthhires")# rnaturalearthhires

# renv restore to previous package versions
# only use if the current package version need to be reverted to a previous state
renv::history()
renv::revert(project = "scari", commit = "")

# repair common issues
renv::repair()

# update renv to latest version in this project (but this will differ from the source download of this project folder)
renv::upgrade()

# run this to rebuild and reinstall all packages in the library- this starts them from a clean slate
renv::rebuild()

```

``` {r update packages or renv}
# update packages
# only run as needed if a package breaks or something
renv::update()
# alternatively, install the latest CRAN versions:
renv::install()
# record newest package versions in lockfile
renv::snapshot()

```

# remove renv

As needed, you may need to completely remove renv and restart from a fresh install.

```{r remove renv}

# deactivate renv
renv::deactivate(project = file.path(here::here()))

# remove renv from project
renv::deactivate(project = file.path(here::here()), clean = TRUE)

# completely remove renv from the system
root <- renv::paths$root()
unlink(root, recursive = TRUE)
utils::remove.packages("renv")

```

