---
title: "Run initial MaxEnt models used to choose background area (project step 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-09-18"
output: html_document
---
# Overview

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava) # for running MaxEnt

library(dismo) # package underneath SDMtune
library(SDMtune) # main package used to run SDMs

# dependencies of SDMtune
library(plotROC) # plots ROCs

library(viridis)

# spatial data handling
library(raster) 
library(terra) 

```

```{r ggplot object for map style}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
        ),
  coord_equal() 
)

```

`SDMtune` will run Maxent through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

```{r check maxent installation}

checkMaxentInstallation(verbose = TRUE)

```

This chunk sets the java memory allocation (`Xmx`). I will increase the memory allocation from 512m (the default) to 2GB of memory. 

```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = "-Xmx2048m")

# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

# MaxEnt Model- entire eastern USA, with "access to cities" covariate

## Tidy datasets for modeling

First, I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used for in the `maxent()` function. 

```{r load in files for all maxent models}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")


# environmental and human impact covariates.
# the env covariates used to train the model
x_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = '\\_easternUSA.asc$', full.names = TRUE)

# these layers will take a long time to load so only do it if you need the global versions
if (FALSE) {
  
  # the same covariates, but on a global scale (used to make predictions)
  x_global_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = "\\.asc$", full.names = TRUE) %>%
    grep("atc_2015_global.asc|bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

}

# slf presences
p_slf_points <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "slf_all_final_coords_v1_2023-08.csv")) %>%
  dplyr::select(-species)

# entire eastern USA background point set
a_entire_easternUSA_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_flexible_area_points.csv"))

```

I will load in the environmental covariates and stack them. I will also shorten their names and exclude possible operators from layer names (for example, using the dash symbol was found to interfere with SDMtune's ability to make predictions for tables downstream).

```{r stack env covariates and make naming consistent}

# stack env covariates
x_env_covariates <- terra::rast(x = x_env_covariates_list)
# stack global versions
x_global_env_covariates <- terra::rast(x = x_global_env_covariates_list)

# attributes
nlyr(x_env_covariates)
names(x_env_covariates)
minmax(x_env_covariates)
# ext(x_env_covariates)
# crs(x_env_covariates)


# layer name object
env_layer_names <- c("atc_2015", "bio11_1981_2010", "bio12_1981_2010", "bio15_1981_2010", "bio2_1981_2010")



# I will change the name of the variables because they are throwing errors in SDMtune
names(x_env_covariates) <- env_layer_names
# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)

# rename raster layers to be the same as the input layers
names(x_global_env_covariates)
# I will change the name of the variables because they are throwing errors in SDMtune
names(x_global_env_covariates) <- env_layer_names

```

I will make some quick edits to the SLF presence data as well. These need to be cropped to the extent of the the background area (in this case, the Eastern USA).

```{r}

# extent object for eastern USA
ext.obj <- terra::ext(-96.503906, -59.589844, 28.304381, 47.457809)

# conert to vector
p_slf_points_vect <- terra::vect(x = p_slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert back to data frame
p_slf_points <- terra::as.data.frame(p_slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole))

```

Finally, I will create an output directory folder to hold the model and its output data.

```{r create directory for file}

mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models")

# create directory for model
dir.create(path = file.path(mypath, "slf_easternUSA_entire_step1", "plots"))

```

## create input data object

I need to create a dataset of presences and background points. This dataset will need to contain point-wise values for each of the predictor covariates. `SDMtune` takes an SWD (sample with data) object for this purpose, containing the presences and background points with associated covariate data to be fed into the model.

```{r prepare SWD data object}

if (FALSE) {

  entire_easternUSA_SWD <- SDMtune::prepareSWD(species = "Lycorma delicatula",
                                               env = x_env_covariates,
                                               p = p_slf_points,
                                               a = a_entire_easternUSA_background_points,
                                               verbose = TRUE # print helpful messages
                                               )
  
  entire_easternUSA_SWD@coords # coordinates
  entire_easternUSA_SWD@pa # presence / absence (background counted as absence)
  entire_easternUSA_SWD@data # extracted data from 

}

```

I will also save the output to be used later.

```{r save SWD file}

SDMtune::swd2csv(swd = entire_easternUSA_SWD, file_name = c(
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_presences_with_data.csv"),
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_absences_with_data.csv")
  ))

```

## Create training / test split

I will split the presences into training and testing, using 80% of the points for training and 20% for testing. I will then use `SDMtune::randomFold()` to split the training data into 5 partitions for cross-validation. This method was loosely adapted from Srivastava et.al, 2021. 

```{r split data for training and testing}

set.seed(4)

entire_easternUSA_trainTest <-  SDMtune::trainValTest(
  x = entire_easternUSA_SWD,
  test = 0.2,
  only_presence = TRUE
)

# separate off training data
entire_easternUSA_train <- entire_easternUSA_trainTest[[1]]
entire_easternUSA_test <- entire_easternUSA_trainTest[[2]]

entire_easternUSA_train@coords # coordinates
entire_easternUSA_train@pa # presence / absence (background counted as absence)
entire_easternUSA_train@data # extracted data from

```

```{r split data into k folds}

# create random folds
entire_easternUSA_trainFolds <- SDMtune::randomFolds(
  data = entire_easternUSA_train,
  k = 5, # 5 folds
  only_presence = TRUE,
  seed = 5 
)

```

## Train Maxent model

First, I will train a model with 5 cross-validated replicates. The model will only be trained on 80% of the slf presence data, with the other 20% being used downstream for analysis. The default settings will be used for Maxent, apart from the following changes:

* ALL feature types used (l = linear, q = quadratic, p = product, h = hinge, t = threshold)
* replicates = 5
* iterations = 5000. This is the max number of iterations of optimization algorithm to perform before stopping training. Increasing this number allows the algorithm to make more refined predictions.

Other changes to the default will be used, but are not relevant for training the model.

```{r train maxent model}

entire_easternUSA_model <- SDMtune::train(
  method = "Maxent",
  data = entire_easternUSA_train,
  folds = entire_easternUSA_trainFolds, # 5 folds for dataset
  fc = "lqpht", # feature classes set to ALL
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Evaluate maxent model

This function should produce all summary statistics for this model

```{r compute summary statistics using custom function}

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = entire_easternUSA_model, 
  model.name = "entire_easternUSA_model", 
  mypath = mypath, 
  env.covar.obj = x_env_covariates, 
  train.obj = entire_easternUSA_train, 
  trainFolds.obj = entire_easternUSA_trainFolds, 
  test.obj = entire_easternUSA_test, 
  plot.types = c("train", "test"), 
  threshold.types = c("cloglog", "logistic")
  )

```

### Fixed area AUC

THIS NEEDS an AUC plot with lines for the fixed vs flexible AUC lines on it



# MaxEnt Model- entire eastern USA, without "access to cities" covariate







```{r compute summary statistics using custom function}

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = , 
  model.name = , 
  mypath = mypath, 
  env.covar.obj = , 
  train.obj = , 
  trainFolds.obj = , 
  test.obj = , 
  plot.types = c("train", "test"), 
  threshold.types = c("cloglog", "logistic")
  )

```



# MaxEnt Model- DD 1% adult emergence


```{r}
# DD model used with 1% adult emergence threshold
a_adult_emergence_1_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emergence-1_flexible_area_points.csv")) 

```






```{r compute summary statistics using custom function}

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = , 
  model.name = , 
  mypath = mypath, 
  env.covar.obj = , 
  train.obj = , 
  trainFolds.obj = , 
  test.obj = , 
  plot.types = c("train", "test"), 
  threshold.types = c("cloglog", "logistic")
  )

```




# MaxEnt Model- 355km buffer around presences

```{r create output directory for model}

# 355km buffer around slf points
a_slf_buffered_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_points_buffered_flexible_area_points.csv"))

```






```{r compute summary statistics using custom function}

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = , 
  model.name = , 
  mypath = mypath, 
  env.covar.obj = , 
  train.obj = , 
  trainFolds.obj = , 
  test.obj = , 
  plot.types = c("train", "test"), 
  threshold.types = c("cloglog", "logistic")
  )

```



# References

Feng, X. 2022, April 24. shandongfx/nimbios_enm. GitHub. Accessed on 2023-9-18.

Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.4). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2023-9-18.

Srivastava, V., A. D. Roe, M. A. Keena, R. C. Hamelin, and V. C. Griess. 2021. Oh the places they’ll go: improving species distribution modelling for invasive forest pests in an uncertain world. Biological Invasions 23:297–349.

Vignali, S., A. Barras, V. Braunisch, and C. B.-U. of Bern. 2023, July 3. SDMtune: Species Distribution Model Selection.


