---
title: "Run initial MaxEnt models used to choose background area (project step 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-09-18"
output: html_document
---
# Overview

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava)

library(dismo) # runs maxent
library(ENMeval) # 

library(sf) # spatial data 
library(raster) # spatial data 
library(terra) # spatial data 

```

`Dismo` will run Maxent through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

This chunk sets the java memory allocation (`Xmx`). I will start with 2GB memory. 

```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = "-Xmx2048m")

# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

# Setup for MaxEnt

- I may need to convert my slf points to a spatial points object

- IF I need to extract point-wise climatic data for MaxEnt, use these steps.

1. stack env covariates 
2. use slf pointwise data to extract a value from each layer at that each point

THIS IS A GOOD VALIDATION STEP- can use is.na()

First, I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used for in the `maxent()` function. 

```{r load in files for all maxent models}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")

# environmental and human impact covariates
# pull those of the proper ending
x_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = '\\_easternUSA.asc$', full.names = TRUE)



# slf presences
p_slf_points <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "slf_all_final_coords_v1_2023-08.csv"))

# background point sets
# entire eastern USA
a_entire_easternUSA_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_flexible_area_points.csv"))
# DD model used with 1% adult emergence threshold
a_adult_emergence_1_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emergence-1_flexible_area_points.csv"))
# 355km buffer around slf points
a_slf_buffered_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_points_buffered_flexible_area_points.csv"))

```

I will load in the environmental covariates and stack them. For the slf presences, I will divide them into training and testing folds using `dismo::kfold()`. I will be running these models with 5 replicates, so I will divide the presence data into 5 groups (80% training, 20% testing). The 20% group that is used for testing is cycled per model replicate.

```{r stack env covariates}

x_env_covariates <- terra::rast(x = x_env_covariates_list)

# attributes
nlyr(x_env_covariates)
names(x_env_covariates)
# ext(x_env_covariates)
# crs(x_env_covariates)

```

```{r divide presences via kfold groups}

fold <- kfold(p_slf_points, k = 5) 

occtest <- occ[fold == 1, ] 
occtrain <- occ[fold != 1, ]

```


# Trial of Maxent

Use `if(dismo::maxent()) { }` to test whether maxent will run on data given to it before actually running the models.

*Note* I prefer to do spatial data manipulation in `terra`, but `dismo` is only compatible with the `raster` package.

```{r trial of maxent (entire eastern USA)}

if(dismo::maxent()) {
  
   maxent(x = x_env_covariates, 
          p = slf_points, 
          path = paste0(getwd(),"/output/maxent_outputs4_cross"), 
          args=prepPara(userfeatures="LQ"
                            )
   )
  
  
}

```

# Maxent Model- entire eastern USA

```{r create output directory for model}

mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models")

# create file directory
dir.create(path = file.path(mypath, "slf_easternUSA_entire_step1"))

```

```{r MaxEnt model of entire eastern USA}

   maxent(x = x_env_covariates, 
          p = slf_points, 
          a = a_entire_easternUSA_background_points,
          path = file.path(mypath, "slf_easternUSA_entire_step1"), 
          args = prepPara(userfeatures = "LQ",
                          )
   )

```



# MaxEnt Model- DD 1% adult emergence

```{r create output directory for model}

```



# MaxEnt Model- 355km buffer around presences

```{r create output directory for model}

```





# References

Feng, X. 2022, April 24. shandongfx/nimbios_enm. GitHub. Accessed on 2023-9-18.

Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.4). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2023-9-18.




