---
title: "Run initial MaxEnt models used to choose background area (project step 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-09-18"
output: html_document
---
# Overview

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava) # for running MaxEnt

library(dismo) # package underneath SDMtune
library(SDMtune) # main package used to run SDMs

# dependencies of SDMtune
library(plotROC) # plots ROCs

library(viridis)

# spatial data handling
library(raster) 
library(terra) 

library(pROC)

```

`SDMtune` will run Maxent through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

```{r check maxent installation}

checkMaxentInstallation(verbose = TRUE)

```

This chunk sets the java memory allocation (`Xmx`). I will increase the memory allocation from 512m (the default) to 2GB of memory. 

```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = "-Xmx2048m")

# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

```{r ggplot object for map style}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
        ),
  coord_equal() 
)

```

# MaxEnt Model- entire eastern USA

## Tidy datasets for modeling

First, I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used for in the `maxent()` function. 

```{r load in files for entire easternUSA}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")


# environmental and human impact covariates.
# the env covariates used to train the model
x_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = '\\_easternUSA.asc$', full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# these layers will take a long time to load so only do it if you need the global versions
if (FALSE) {
  
  # the same covariates, but on a global scale (used to make predictions)
  x_global_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = "\\.asc$", full.names = TRUE) %>%
    grep("bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

}

# slf presences
p_slf_points <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "slf_all_final_coords_v1_2023-08.csv")) %>%
  dplyr::select(-species)

# entire eastern USA background point set
a_easternUSA_entire_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_flexible_area_points.csv"))

```

I will load in the environmental covariates and stack them. I will also shorten their names and exclude possible operators from layer names (for example, using the dash symbol was found to interfere with SDMtune's ability to make predictions for tables downstream).

```{r stack env covariates and make naming consistent}

# stack env covariates
x_env_covariates <- terra::rast(x = x_env_covariates_list)

# attributes
nlyr(x_env_covariates)
names(x_env_covariates)
minmax(x_env_covariates)
# ext(x_env_covariates)
# crs(x_env_covariates)


# layer name object. Check order of layers first
env_layer_names <- c("bio11_1981_2010", "bio12_1981_2010", "bio15_1981_2010", "bio2_1981_2010")



# I will change the name of the variables because they are throwing errors in SDMtune
names(x_env_covariates) <- env_layer_names
# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)


# these are the global versions of the layers and will take a long time to run. Only do so if necessary for model evaluation
if (FALSE) {

  # stack global versions
  x_global_env_covariates <- terra::rast(x = x_global_env_covariates_list)
  # rename raster layers to be the same as the input layers
  names(x_global_env_covariates)
  # I will change the name of the variables because they are throwing errors in SDMtune
  names(x_global_env_covariates) <- env_layer_names

}

```

I will make some quick edits to the SLF presence data as well. These need to be cropped to the extent of the the background area (in this case, the Eastern USA).

```{r crop presences to rasters}

# extent object for eastern USA
ext.obj <- terra::ext(-96.503906, -59.589844, 28.304381, 47.457809)

# conert to vector
p_slf_points_vect <- terra::vect(x = p_slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert back to data frame
p_slf_points <- terra::as.data.frame(p_slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole))

```

### create input data object

I need to create a dataset of presences and background points. This dataset will need to contain point-wise values for each of the predictor covariates. `SDMtune` takes an SWD (sample with data) object for this purpose, containing the presences and background points with associated covariate data to be fed into the model.

```{r prepare SWD data object}

if (FALSE) {

  easternUSA_entire_SWD <- SDMtune::prepareSWD(species = "Lycorma delicatula",
                                               env = x_env_covariates,
                                               p = p_slf_points,
                                               a = a_easternUSA_entire_background_points,
                                               verbose = TRUE # print helpful messages
                                               )
  
  easternUSA_entire_SWD@coords # coordinates
  easternUSA_entire_SWD@pa # presence / absence (background counted as absence)
  easternUSA_entire_SWD@data # extracted data from 

}

```

I will also save the output to be used later.

```{r save SWD file}

SDMtune::swd2csv(swd = easternUSA_entire_SWD, file_name = c(
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_presences_with_data.csv"),
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_absences_with_data.csv")
  ))

```

### Create training / test split

I will split the presences into training and testing, using 80% of the points for training and 20% for testing. I will then use `SDMtune::randomFold()` to split the training data into 5 partitions for cross-validation. This method was loosely adapted from Srivastava et.al, 2021. 

```{r split data for training and testing}

set.seed(4)

easternUSA_entire_trainTest <-  SDMtune::trainValTest(
  x = easternUSA_entire_SWD,
  test = 0.2,
  only_presence = TRUE
)

# separate off training data
easternUSA_entire_train <- easternUSA_entire_trainTest[[1]]
easternUSA_entire_test <- easternUSA_entire_trainTest[[2]]

easternUSA_entire_train@coords # coordinates
easternUSA_entire_train@pa # presence / absence (background counted as absence)
easternUSA_entire_train@data # extracted data from

```

```{r split data into k folds}

# create random folds
easternUSA_entire_trainFolds <- SDMtune::randomFolds(
  data = easternUSA_entire_train,
  k = 5, # 5 folds
  only_presence = TRUE,
  seed = 5 
)

```

## Train Maxent model

First, I will train a model with 5 cross-validated replicates. The model will only be trained on 80% of the slf presence data, with the other 20% being used downstream for analysis. The default settings will be used for Maxent, apart from the following changes:

* ALL feature types used (l = linear, q = quadratic, p = product, h = hinge, t = threshold)
* replicates = 5
* iterations = 5000. This is the max number of iterations of optimization algorithm to perform before stopping training. Increasing this number allows the algorithm to make more refined predictions.

Other changes to the default will be used, but are not relevant for training the model.

```{r train maxent model}

easternUSA_entire_model <- SDMtune::train(
  method = "Maxent",
  data = easternUSA_entire_train,
  folds = easternUSA_entire_trainFolds, # 5 folds for dataset
  fc = "lqpht", # feature classes set to ALL
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Evaluate maxent model

### Summary Statistics

This function produces all summary statistics for this model. For the complete and annotated workflow used to create this function, see "051_maxent_workflow.Rmd". 

```{r compute summary statistics using custom function}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_entire_step1_v1")

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = easternUSA_entire_model, 
  model.name = "easternUSA_entire", 
  mypath = mypath, 
  create.dir = TRUE, # create subdirectory
  env.covar.obj = x_env_covariates, # env covariates raster stacked
  train.obj = easternUSA_entire_train, # training data used to create model
  trainFolds.obj = easternUSA_entire_trainFolds,  # k-folds of training data
  test.obj = easternUSA_entire_test, # data you wish to use to test the model
  plot.types = c("train", "test"), # types of jackknife curves to be created
  threshold.types = c("cloglog", "logistic") # types of univariate and marginal response curves to be created
  )

```

### Fixed area AUC

This function produces the fixed area ROC analysis for this model. For the complete and annotated workflow used to create this function, see "052_fixed_ROC_workflow.Rmd". 

```{r compute fixed ROC and produce curve plots}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_entire_step1_v1")

slfSpread::compute_MaxEnt_fixed_ROC(
  model.obj = easternUSA_entire_model,
  model.name = "easternUSA_entire", 
  mypath = mypath,
  fixed.area.points = file.path(here(), "vignette-outputs", "data-tables", "easternUSA_fixed_area_points_v1.csv"),
  env.covar.obj = x_env_covariates,
  test.obj = easternUSA_entire_test
)

```

The next three models will follow roughly the same structure, so comments will be kept to a minimum because the workflow will be the same. 



# MaxEnt Model- DD 1% adult emergence

## Tidy datasets for modeling

This is the same as the setup for the first model, except I will not re-load `x_env_covariates`, because it will remain unchanged. I will also not re-load the SLF presences `p_slf_presences`.

```{r load in background points for maxent model}

# DD model used with 1% adult emergence threshold
a_adult_emergence_1_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emergence-1_flexible_area_points.csv")) 

```

### create input data object

```{r prepare SWD data object}

if (FALSE) {

  easternUSA_adult_emer_1_SWD <- SDMtune::prepareSWD(species = "Lycorma delicatula",
                                               env = x_env_covariates,
                                               p = p_slf_points,
                                               a = a_adult_emergence_1_background_points,
                                               verbose = TRUE # print helpful messages
                                               )
  
  easternUSA_adult_emer_1_SWD@coords # coordinates
  easternUSA_adult_emer_1_SWD@pa # presence / absence (background counted as absence)
  easternUSA_adult_emer_1_SWD@data # extracted data from 

}

```

I will save the output again, but this time I will delete the presences dataset because it is a duplicate.

```{r save SWD file}

SDMtune::swd2csv(swd = easternUSA_adult_emer_1_SWD, file_name = c(
  file.path(here(), "vignette-outputs", "data-tables", "temp.csv"), # this is a duplicate file
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emer_1_absences_with_data.csv")
  ))

# the presences file is a duplicate, so I will delete that
file.remove(file.path(here(), "vignette-outputs", "data-tables", "temp.csv"))

```

### Create training / test split

I will split the presences into training and testing, using 80% of the points for training and 20% for testing. I will then use `SDMtune::randomFold()` to split the training data into 5 partitions for cross-validation. This method was loosely adapted from Srivastava et.al, 2021. 

```{r split data for training and testing}

set.seed(6)

easternUSA_adult_emer_1_trainTest <-  SDMtune::trainValTest(
  x = easternUSA_adult_emer_1_SWD,
  test = 0.2,
  only_presence = TRUE
)

# separate off training data
easternUSA_adult_emer_1_train <- easternUSA_adult_emer_1_trainTest[[1]]
easternUSA_adult_emer_1_test <- easternUSA_adult_emer_1_trainTest[[2]]

easternUSA_adult_emer_1_train@coords # coordinates
easternUSA_adult_emer_1_train@pa # presence / absence (background counted as absence)
easternUSA_adult_emer_1_train@data # extracted data from

```

```{r split data into k folds}

# create random folds
easternUSA_adult_emer_1_trainFolds <- SDMtune::randomFolds(
  data = easternUSA_adult_emer_1_train,
  k = 5, # 5 folds
  only_presence = TRUE,
  seed = 7
)

```

## Train Maxent model

```{r train maxent model}

easternUSA_adult_emer_1_model <- SDMtune::train(
  method = "Maxent",
  data = easternUSA_adult_emer_1_train,
  folds = easternUSA_adult_emer_1_trainFolds, # 5 folds for dataset
  fc = "lqpht", # feature classes set to ALL
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Evaluate maxent model

### Summary Statistics

This function produces all summary statistics for this model. For the complete and annotated workflow used to create this function, see "051_maxent_workflow.Rmd". 

```{r compute summary statistics using custom function}

# output directory, with folder to create at end
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_adult_emer_1_step1")

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = easternUSA_adult_emer_1_model, 
  model.name = "easternUSA_adult_emer_1", 
  mypath = mypath, 
  create.dir = TRUE, # create subdirectory from end of mypath
  env.covar.obj = x_env_covariates, # env covariates raster stacked
  train.obj = easternUSA_adult_emer_1_train, # training data used to create model
  trainFolds.obj = easternUSA_adult_emer_1_trainFolds,  # k-folds of training data
  test.obj = easternUSA_adult_emer_1_test, # data you wish to use to test the model
  plot.types = c("train", "test"), # types of jackknife curves to be created
  threshold.types = c("cloglog", "logistic") # types of univariate and marginal response curves to be created
  )

```

### Fixed area AUC

This function produces the fixed area ROC analysis for this model. For the complete and annotated workflow used to create this function, see "052_fixed_ROC_workflow.Rmd". 

```{r compute fixed ROC and produce curve plots}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_adult_emer_1_step1")

slfSpread::compute_MaxEnt_fixed_ROC(
  model.obj = easternUSA_adult_emer_1_model,
  model.name = "easternUSA_adult_emer_1", 
  mypath = mypath,
  fixed.area.points = file.path(here(), "vignette-outputs", "data-tables", "easternUSA_fixed_area_points_v1.csv"),
  env.covar.obj = x_env_covariates,
  test.obj = easternUSA_adult_emer_1_test
)

```



# MaxEnt Model- 355km buffer around presences

## Tidy datasets for modeling

This is the same as the setup for the first model, except I will not re-load `x_env_covariates`, because it will remain unchanged. I will also not re-load the SLF presences `p_slf_presences`.

```{r load in background points for maxent model}

# 355km buffer around slf points
a_slf_buffered_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_points_buffered_flexible_area_points.csv"))

```

### create input data object

```{r prepare SWD data object}

if (FALSE) {

  easternUSA_adult_emer_1_SWD <- SDMtune::prepareSWD(species = "Lycorma delicatula",
                                               env = x_env_covariates,
                                               p = p_slf_points,
                                               a = a_slf_buffered_background_points,
                                               verbose = TRUE # print helpful messages
                                               )
  
  easternUSA_adult_emer_1_SWD@coords # coordinates
  easternUSA_adult_emer_1_SWD@pa # presence / absence (background counted as absence)
  easternUSA_adult_emer_1_SWD@data # extracted data from 

}

```

I will save the output again, but this time I will delete the presences dataset because it is a duplicate.

```{r save SWD file}

SDMtune::swd2csv(swd = easternUSA_adult_emer_1_SWD, file_name = c(
  file.path(here(), "vignette-outputs", "data-tables", "temp.csv"), # this is a duplicate file
  file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emer_1_absences_with_data.csv")
  ))

# the presences file is a duplicate, so I will delete that
file.remove(file.path(here(), "vignette-outputs", "data-tables", "temp.csv"))

```

### Create training / test split

I will split the presences into training and testing, using 80% of the points for training and 20% for testing. I will then use `SDMtune::randomFold()` to split the training data into 5 partitions for cross-validation. This method was loosely adapted from Srivastava et.al, 2021. 

```{r split data for training and testing}

set.seed(6)

easternUSA_adult_emer_1_trainTest <-  SDMtune::trainValTest(
  x = easternUSA_adult_emer_1_SWD,
  test = 0.2,
  only_presence = TRUE
)

# separate off training data
easternUSA_adult_emer_1_train <- easternUSA_adult_emer_1_trainTest[[1]]
easternUSA_adult_emer_1_test <- easternUSA_adult_emer_1_trainTest[[2]]

easternUSA_adult_emer_1_train@coords # coordinates
easternUSA_adult_emer_1_train@pa # presence / absence (background counted as absence)
easternUSA_adult_emer_1_train@data # extracted data from

```

```{r split data into k folds}

# create random folds
easternUSA_adult_emer_1_trainFolds <- SDMtune::randomFolds(
  data = easternUSA_adult_emer_1_train,
  k = 5, # 5 folds
  only_presence = TRUE,
  seed = 7
)

```

## Train Maxent model

```{r train maxent model}

easternUSA_adult_emer_1_model <- SDMtune::train(
  method = "Maxent",
  data = easternUSA_adult_emer_1_train,
  folds = easternUSA_adult_emer_1_trainFolds, # 5 folds for dataset
  fc = "lqpht", # feature classes set to ALL
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Evaluate maxent model

### Summary Statistics

This function produces all summary statistics for this model. For the complete and annotated workflow used to create this function, see "051_maxent_workflow.Rmd". 

```{r compute summary statistics using custom function}

# output directory, with folder to create at end
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_adult_emer_1_step1")

slfSpread::compute_MaxEnt_summary_statistics(
  model.obj = easternUSA_adult_emer_1_model, 
  model.name = "easternUSA_adult_emer_1", 
  mypath = mypath, 
  create.dir = TRUE, # create subdirectory from end of mypath
  env.covar.obj = x_env_covariates, # env covariates raster stacked
  train.obj = easternUSA_adult_emer_1_train, # training data used to create model
  trainFolds.obj = easternUSA_adult_emer_1_trainFolds,  # k-folds of training data
  test.obj = easternUSA_adult_emer_1_test, # data you wish to use to test the model
  plot.types = c("train", "test"), # types of jackknife curves to be created
  threshold.types = c("cloglog", "logistic") # types of univariate and marginal response curves to be created
  )

```

### Fixed area AUC

This function produces the fixed area ROC analysis for this model. For the complete and annotated workflow used to create this function, see "052_fixed_ROC_workflow.Rmd". 

```{r compute fixed ROC and produce curve plots}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_adult_emer_1_step1")

slfSpread::compute_MaxEnt_fixed_ROC(
  model.obj = easternUSA_adult_emer_1_model,
  model.name = "easternUSA_adult_emer_1", 
  mypath = mypath,
  fixed.area.points = file.path(here(), "vignette-outputs", "data-tables", "easternUSA_fixed_area_points_v1.csv"),
  env.covar.obj = x_env_covariates,
  test.obj = easternUSA_adult_emer_1_test
)

```

# Results- Fixed and Flexible AUC comparison for all 3 models

PLOT should be a boxplot, with the background size on the x-axis (3 categories) and the fixed-area AUC value on the y-axis.
Another plot could be made using the same data, except the flexible-area AUC on the y-axis



# Appendix

### Create distribution map for area of interest- easternUSA_entire_model

I will use the predict function again, but this time to predict the suitability for the range of interest. The output will be a raster of suitability. This will not be done for every model in this step, but will be done for the entire models with and without ATC (to compare which is a better fit).

```{r predict function to create raster}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_entire_step1_v1")

# write .asc file of avg predictions
SDMtune::predict(
  object = easternUSA_entire_model, 
  data = x_env_covariates, # the covariate layers used to train the model
  fun = "mean", 
  type = "cloglog", 
  clamp = FALSE, 
  progress = TRUE, 
  filename = file.path(mypath, "easternUSA_entire_predicted_suitability.asc"), 
    # the function automatically adds the function name on the end
  filetype = "AAIGrid"
) 

```

I will convert the raster to a data frame and use this to plot it and save a jpg.

```{r convert to df and plot}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_entire_step1_v1")

# load in mean predictions
easternUSA_entire_suit <- terra::rast(x = file.path(mypath, "easternUSA_entire_predicted_suitability_mean.asc")) %>%
  # terra::subst(., from = -9999, to = NA) %>% # change -9999 to NA for plotting
  terra::as.data.frame(., xy = TRUE) 
  
# plot
easternUSA_entire_suit_plot <- ggplot() +
  geom_raster(data = easternUSA_entire_suit, 
              aes(x = x, y = y, fill = mean)) +
  labs(title = "Suitability for SLF") +
  viridis::scale_fill_viridis(option = "D") +
  map_style

# save plot output
ggsave(easternUSA_entire_suit_plot, 
           filename = file.path(mypath, "plots", "easternUSA_entire_predicted_suitability_mean.jpg"),
           height = 8, 
           width = 10,
           device = "jpeg",
           dpi = "retina")

```

# References

Feng, X. 2022, April 24. shandongfx/nimbios_enm. GitHub. Accessed on 2023-9-18.

Steven Phillips. (2017). A Brief Tutorial on Maxent. http://biodiversityinformatics.amnh.org/open_source/maxent/.

Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.4). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2023-9-18.

Srivastava, V., A. D. Roe, M. A. Keena, R. C. Hamelin, and V. C. Griess. 2021. Oh the places they’ll go: improving species distribution modelling for invasive forest pests in an uncertain world. Biological Invasions 23:297–349.

VanDerWal, J., Shoo, L. P., Graham, C., & Williams, S. E. (2009). Selecting pseudo-absence data for presence-only distribution modeling: How far should you stray from what you know? Ecological Modelling, 220(4), 589–594. https://doi.org/10.1016/j.ecolmodel.2008.11.010

Vignali, S., A. Barras, V. Braunisch, and C. B.-U. of Bern. 2023, July 3. SDMtune: Species Distribution Model Selection.


