---
title: "Run initial MaxEnt models used to choose background area (project step 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-09-18"
output: html_document
---
# Overview

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava)

library(dismo) # runs maxent
library(ENMeval) # 

library(sf) # spatial data 
library(raster) # spatial data 
library(terra) # spatial data 

```

```{r ggplot object for map style}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
        ),
  coord_equal() 
)

```

`Dismo` will run Maxent through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

This chunk sets the java memory allocation (`Xmx`). I will start with 2GB memory. 

```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = "-Xmx2048m")

# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

# Setup for MaxEnt

- I may need to convert my slf points to a spatial points object

- IF I need to extract point-wise climatic data for MaxEnt, use these steps.

1. stack env covariates 
2. use slf pointwise data to extract a value from each layer at that each point

THIS IS A GOOD VALIDATION STEP- can use is.na()

First, I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used for in the `maxent()` function. 

```{r load in files for all maxent models}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")

# environmental and human impact covariates
# pull those of the proper ending
x_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent"), pattern = '\\_easternUSA.asc$', full.names = TRUE)



# slf presences
p_slf_points <- read.csv(file = file.path(here(), "vignette-outputs", "data-tables", "slf_all_final_coords_v1_2023-08.csv")) %>%
  mutate(row.number = seq(nrow(.))) %>%
  dplyr::select(row.number, x, y) 

# background point sets
# entire eastern USA
a_entire_easternUSA_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_entire_flexible_area_points.csv")) %>%
  mutate(row.number = seq(nrow(.))) 
# DD model used with 1% adult emergence threshold
a_adult_emergence_1_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_adult_emergence-1_flexible_area_points.csv")) %>%
  mutate(row.number = seq(nrow(.))) 
# 355km buffer around slf points
a_slf_buffered_background_points <- read.csv(file.path(here(), "vignette-outputs", "data-tables", "easternUSA_slf_points_buffered_flexible_area_points.csv")) %>%
  mutate(row.number = seq(nrow(.))) 

```

I will load in the environmental covariates and stack them.

```{r stack env covariates}

x_env_covariates <- terra::rast(x = x_env_covariates_list)

# attributes
nlyr(x_env_covariates)
names(x_env_covariates)
minmax(x_env_covariates)
# ext(x_env_covariates)
# crs(x_env_covariates)

```

## Slf Points kfold

For the slf presences, I will divide them into training and testing folds using `dismo::kfold()`. I will be running these models with 5 replicates, so I will divide the presence data into 5 groups (80% training, 20% testing). The 20% group that is used for testing is cycled per model replicate. This is in preparation for MaxEnt to perform k-fold cross validation. 

```{r divide presences into kfold groups}

# set random seed so it is consistent every time
set.seed(4)
# get kfold of groups
p_slf_points_kfold <- kfold(p_slf_points, k = 5) %>%
  as.data.frame() %>%
  mutate(row.number = seq(nrow(p_slf_points)))
# change names of variables
names(p_slf_points_kfold) <- c("kfold_group", "row.number")


p_slf_points_kfold <- full_join(p_slf_points, p_slf_points_kfold, by = "row.number") %>%
  dplyr::select(kfold_group, x, y)

```

```{r plot}

(ggplot() +
   geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = "white", color = "black", lwd = 0.15) +
   geom_point(data = p_slf_points_kfold, aes(x = x, y = y, color = as.factor(kfold_group)), size = 1) +
   map_style +
   guides(color = guide_legend(override.aes = list(size = 2))) +
   coord_equal(xlim = c(-87.747803, -66.840820), ylim = c(34.678394, 42.956423))
)

```

## Background Points kfold

Now, I will perform the same procedure for each of the 3 sets of background points.

```{r divide entire easternUSA absence dataset into kfold groups}

# set random seed so it is consistent every time
set.seed(5)
# get kfold of groups
a_entire_easternUSA_background_points_kfold <- kfold(a_entire_easternUSA_background_points, k = 5) %>%
  as.data.frame() %>%
  mutate(row.number = seq(nrow(a_entire_easternUSA_background_points)))
# change names of variables
names(a_entire_easternUSA_background_points_kfold) <- c("kfold_group", "row.number")


a_entire_easternUSA_background_points_kfold <- full_join(a_entire_easternUSA_background_points, a_entire_easternUSA_background_points_kfold, by = "row.number") %>%
  dplyr::select(kfold_group, x, y)

```

```{r divide DD absence dataset into kfold groups}

# set random seed so it is consistent every time
set.seed(6)
# get kfold of groups
a_adult_emergence_1_background_points_kfold <- kfold(a_adult_emergence_1_background_points, k = 5) %>%
  as.data.frame() %>%
  mutate(row.number = seq(nrow(a_adult_emergence_1_background_points)))
# change names of variables
names(a_adult_emergence_1_background_points_kfold) <- c("kfold_group", "row.number")


a_adult_emergence_1_background_points_kfold <- full_join(a_adult_emergence_1_background_points, a_adult_emergence_1_background_points_kfold, by = "row.number") %>%
  dplyr::select(kfold_group, x, y)

```

```{r divide 355km buffer absence dataset into kfold groups}

# set random seed so it is consistent every time
set.seed(7)
# get kfold of groups
a_slf_buffered_background_points_kfold <- kfold(a_slf_buffered_background_points, k = 5) %>%
  as.data.frame() %>%
  mutate(row.number = seq(nrow(a_slf_buffered_background_points)))
# change names of variables
names(a_slf_buffered_background_points_kfold) <- c("kfold_group", "row.number")


a_slf_buffered_background_points_kfold <- full_join(a_slf_buffered_background_points, a_slf_buffered_background_points_kfold, by = "row.number") %>%
  dplyr::select(kfold_group, x, y)

```

## 



# Trial of Maxent

Use `if(dismo::maxent()) { }` to test whether maxent will run on data given to it before actually running the models.

*Note* I prefer to do spatial data manipulation in `terra`, but `dismo` is only compatible with the `raster` package.

```{r trial of maxent (entire eastern USA)}

if(dismo::maxent()) {
  
   maxent(x = x_env_covariates, 
          p = p_slf_points_kfold, 
          path = file.path(mypath, ), 
          args = list(userfeatures = "LQ"
                      )
   )
  
  
}

```

# Maxent Model- entire eastern USA

```{r create output directory for model}

mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models")

# create file directory
dir.create(path = file.path(mypath, "slf_easternUSA_entire_step1"))

```

```{r MaxEnt model of entire eastern USA}

   maxent(x = x_env_covariates, 
          p = slf_points, 
          a = a_entire_easternUSA_background_points,
          path = file.path(mypath, "slf_easternUSA_entire_step1"), 
          args = list(userfeatures = "LQ",
                      )
   )

```



# MaxEnt Model- DD 1% adult emergence

```{r create output directory for model}

```



# MaxEnt Model- 355km buffer around presences

```{r create output directory for model}

```





# References

Feng, X. 2022, April 24. shandongfx/nimbios_enm. GitHub. Accessed on 2023-9-18.

Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.4). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2023-9-18.




