---
title: "053_create_suitability_maps_workflow"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-10-20"
output: html_document
---

# Overview

This vignette will outline the workflow used to create the mean and thresholded suitability maps. These maps are created for the easternUSA models and main models. 

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)

# modeling
library(SDMtune) # main package used to run SDMs
# spatial data handling
library(terra) 

library(viridis)

```

```{r ggplot object for map style}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue",
                                colour = "lightblue")
        ),
  labs(fill = "Suitability for SLF"),
  coord_equal() 
)

```

# Create mean distribution map 

```{r predict function to create raster}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_buffered_step1")

# write .asc file of avg predictions
SDMtune::predict(
  object = easternUSA_buffered_model, 
  data = x_env_covariates, # the covariate layers used to train the model
  fun = "mean", 
  type = "cloglog", 
  clamp = FALSE, 
  progress = TRUE, 
  filename = file.path(mypath, "easternUSA_buffered_predicted_suitability.asc"), 
    # the function automatically adds the function name on the end
  filetype = "AAIGrid"
) 

```

```{r convert to df and plot}

# load in mean predictions
easternUSA_buffered_suit <- terra::rast(x = file.path(mypath, "easternUSA_buffered_predicted_suitability_mean.asc")) %>%
  terra::as.data.frame(., xy = TRUE) 
  
# plot
easternUSA_buffered_suit_plot <- ggplot() +
  geom_raster(data = easternUSA_buffered_suit, 
              aes(x = x, y = y, fill = mean)) +
  labs(title = "Mean suitability for SLF",
       subtitle = "Model: easternUSA_buffered") +
  viridis::scale_fill_viridis(option = "D") +
  map_style

# save plot output
ggsave(easternUSA_buffered_suit_plot, 
           filename = file.path(mypath, "plots", "easternUSA_buffered_predicted_suitability_mean.jpg"),
           height = 8, 
           width = 10,
           device = "jpeg",
           dpi = "retina")

```

# Create thresholded suitability map



```{r discern MTSS threshold}

# output directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_buffered_step1")

MTSS_value <- read.csv(file = file.path(mypath, "easternUSA_buffered_summary_all_iterations.csv")) %>%
  .[42, 6]

```

## Create binary raster with MTSS threshold

This is for later use.

```{r terra required classification matrices}

# create regional suitability value matrix for terra
buffered_rescale_class <- data.frame(
  from = c(0, MTSS_value),
  to = c(MTSS_value, 1),
  becomes = c(0, 1)
)

```

```{r reclassify maxent outputs}

# path to directory
mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models/slf_easternUSA_buffered_step1")

# load in raster
buffered_mean_raster <- terra::rast(x = file.path(mypath, "easternUSA_buffered_predicted_suitability_mean.asc"))

if(FALSE) {
  
  terra::classify(x = buffered_mean_raster, 
                  rcl = buffered_rescale_class,
                  right = FALSE, # close left side of value
                  filename = file.path(mypath, "easternUSA_buffered_predicted_suitability_mean_thresholded.asc"), # also write to file
                  overwrite = FALSE
                  ) 
  
}
  
```

## Create mask raster for thresholded raster figure

This raster will only be used as an overlay layer in the thresholded suitability map plot. It is a binary raster of the the entire land mass (the easternUSA), with unsuitable areas as a 1 and everything else as an NA. This plot can be put on top of the mean suitability layer and set to a color to "grey out" unsuitable areas on land. 

```{r reclassify maxent outputs}

# create regional suitability value matrix for terra
buffered_rescale_class <- data.frame(
  from = 0,
  to = MTSS_value,
  becomes = 1
)

if(TRUE) {
  
  # reclassify and write regional raster
  buffered_mask_layer <- terra::classify(x = buffered_mean_raster, 
                                             rcl = buffered_rescale_class,
                                             right = FALSE, # close left side of value
                                             others = NA,
                                             filename = file.path(mypath, "easternUSA_buffered_MTSS_mask_layer.asc"), # also write to file
                                             overwrite = FALSE) 
  
}
  
```

## Create thresholded suitability map

```{r overlay plots}
  
# load in mean predictions
buffered_mask_layer_df <- terra::rast(x = file.path(mypath, "easternUSA_buffered_MTSS_mask_layer.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

# plot mean raster first
easternUSA_buffered_threshold_plot <- ggplot() +
  # plot regular raster of values first
  geom_raster(data = easternUSA_buffered_suit, 
              aes(x = x, y = y, fill = mean)) +
    # plot binary threshold on top
  geom_raster(data = buffered_mask_layer_df, 
              aes(x = x, y = y), fill = "azure4") +
  labs(title = "Suitability for SLF, MTSS training threshold",
       subtitle = "Model: 355km buffer") +
  viridis::scale_fill_viridis(option = "D") +
  map_style

# save plot output
ggsave(easternUSA_buffered_threshold_plot, 
           filename = file.path(mypath, "plots", "easternUSA_buffered_predicted_suitability_mean_thresholded.jpg"),
           height = 8, 
           width = 10,
           device = "jpeg",
           dpi = "retina")

```
