---
title: "Run initial MaxEnt models used to choose background area (project step 1)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-09-18"
output: html_document
---

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
library(devtools)
library(rJava)

library(dismo) # runs maxent
library(ENMeval) # 

library(sf) # spatial data 
library(raster) # spatial data 
library(terra) # spatial data 

```










```{r control MaxEnt settings}

# xss sets java stack size
# xmx sets java memory allocation
options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

- I may need to convert my slf points to a spatial points object

- IF I need to extract point-wise climatic data for MaxEnt, use these steps.

1. stack env covariates 
2. use slf pointwise data to extract a value from each layer at that each point

THIS IS A GOOD VALIDATION STEP- can use is.na()


```{r trial of maxent}



```

# References

Feng, X. 2022, April 24. shandongfx/nimbios_enm. GitHub. Accessed on 2023-9-18.

Steven J. Phillips, Miroslav Dudík, Robert E. Schapire. [Internet] Maxent software for modeling species niches and distributions (Version 3.4.4). Available from url: http://biodiversityinformatics.amnh.org/open_source/maxent/. Accessed on 2023-9-18.




