---
title: "Setup for regional and global models (project step 2)"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2023-10-20"
output: html_document
---

# Overview

In the last vignette, I selected the background area type that would be the most suited to modeling SLF. I found that using the entire background area was the best choice. I will now set up for the next two sets of models, a historical global/regional pair and a climate change global/regional pair. 

(EXPAND)

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() starts at the root folder of this package.
# utilities
library(devtools)
library(rJava) # for running MaxEnt
# spatial data handling
library(raster) 
library(terra) 
library(dismo) 
# other
library(viridis)

```


# 1. Crop Bioclim Layers to N. America

## Historical Data

```{r setup for cropping bioclim layers}

# path to directory
  mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")

# lists of target files in the directory
# load in bioclim layers to be cropped- the original .asc files
env.files <- list.files(path = file.path(mypath, "v1_maxent"), pattern = "\\.asc$", full.names = TRUE) %>%
# extract the 4 bioclim layers I will be using in my models. Access to cities will not be used until later models
  grep("atc_2015_global.asc|bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

# output file names
output.files <- list.files(path = file.path(mypath, "v1_maxent"), pattern = "\\.asc$", full.names = FALSE) %>%
  grep("atc|bio2|bio11|bio12|bio15", ., value = TRUE) %>%
  gsub(pattern = "global", replacement = "easternUSA")

# extent object for eastern USA
ext.obj <- terra::ext(-96.503906, -59.589844, 28.304381, 47.457809)

```

```{r loop to crop bioclim layers to eastern N america}

# view list of filetypes for terra, use .ascii
View(terra::gdal(drivers = TRUE))

if(TRUE) {
  
  # loop to crop extent for all files
  for(a in seq_along(env.files)){

    #ensure that the CRS is consistent
    rast.hold <- terra::rast(env.files[a])
    
    # crop new rasters to extent
    rast.hold <- terra::crop(x = rast.hold, y = ext.obj, overwrite = FALSE)
    
    #write out the new resampled rasters!
    terra::writeRaster(x = rast.hold, filename = file.path(mypath, "v1_maxent", output.files[a]), filetype = "AAIGrid", overwrite = FALSE)
    
    # remove object once its done
    rm(rast.hold)
    
  }
  
}

# I am pretty sure this method resets the raster cell numbers, which might be annoying downstream....

```

Lets plot one of the new layers to make sure the cropping worked.

```{r plot main_layer to ensure cropping worked}

bio11 <- terra::rast(x = file.path(mypath, "v1_maxent", "bio11_1981-2010_easternUSA.asc")) 
# convert to df
bio11_df <- terra::as.data.frame(bio11, xy = TRUE)

# plot main layer as example
(ggplot() +
  # change scale of plots to be standard across figures
  geom_raster(data = bio11_df, 
            aes(x = x, y = y, fill = `CHELSA_bio11_1981-2010_V.2.1`)) +
  xlab("longitude") +
  ylab("latitude") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_equal()
)
# the cropping worked

```

The layers have been cropped to the proper extent.

## SSP370 Future Climate data

# 2. Select Flexible Area Background Points



# Appendix



# References


