---
title: "Setup for regional-scale MaxEnt models"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2024-10-07"
---

# Overview

In the previous vignette, I ran the global model, which was the first step to examine suitability for SLF at multiple spatial scales. In this vignette, I will set up to run three regional-scale models, one for each geographical region where SLF is found. I will run a model based on SLF presence data from the native range (China), the invaded region in Asia (Japan and South Korea), and the invaded region in North America. I hypothesize that by creating models at multiple spatial scales, I can have more confidence in suitability predictions made for specific SLF populations and important viticultural regions, and can predict the risk of specific populations spreading now and under climate change. I also hypothesize that the regional-scale models, once ensembled into a single prediction, will make more refined predictions than a global-scale model could.

This vignette will set up for these regional-scale models in a similar fashion to the setup I created for the global model in vignette 040, but with some additional steps. First, I will use geopolitical boundaries or bounding boxes to select a broad area for each regional model's background points. I will choose a much broader area than SLF currently occupies in each area so that I can further reduce this area based on climate. I will then narrow this area for each model using the [Koppen-Geiger climate zones](https://koeppen-geiger.vu-wien.ac.at/). If a presence point within each region intersects with a zone, that zone will be included in the area used to choose background points.

For each regional-scale model, I will:

1.  Select the countries or bounding box that will represent that regional-scale model area
2.  Select the SLF presences that fall into that region. These will be used to train the model.
3.  Intersect these SLF presences with the K-G climate zones
4.  Keep the zones that intersect, combine their area and use this area to crop the bioclim layers we tidied in vignette 020
5.  Use the masked bioclim layers to select the background points
6.  Choose the test/validation presences, which will come from a different region than the training points
7.  Plot the chosen training and test presence per model.

Here is a table summarizing the model structures:

```{r model_info table, echo = FALSE}

model_info <- data.frame(
  "model" = c(
    "Rn",
    "Ri.NAmerica",
    "Ri.Asia"
    ),
  "train_region" = c(
    "east Asia native region",
    "N. America invaded region",
    "east Asia invaded region (Koreas & Japan)"
  ),
  "train_presences" = c(
    269,
    338,
    200
  ),
  "test_region" = c(
    "N. America invaded region",
    "east Asia invaded region (Koreas & Japan)",
    "N. America invaded region"
  ),
  "test_presences" = c(
    54,
    68,
    40
  ),
  "background_points" = c(
    "10,000",
    "10,000",
    "10,000"
  )
)

# convert to kable
model_info_kable <- knitr::kable(x = model_info) |>
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE)

```

# Setup

I will prepare for this vignette by loading the necessary packages to run this script. I will also be creating maps during this analysis, so I will create a list object containing a standardized map style that I will continue to use.

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here::here() # here() starts at the root folder of this package.
library(devtools)

library(dismo) # generate random background points

# spatial data handling
library(raster) 
library(terra)
library(sf)

# spatial data sources
library(rnaturalearth)
# remotes::install_github("ropensci/rnaturalearthhires")
library(rnaturalearthhires)
library(kgc) # koppen climate rasters

# aesthetics
library(webshot)
library(webshot2)
library(kableExtra)
library(patchwork)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

```{r save tables from above, if desired}
  
# model training dataset
# save as .csv
write_csv(model_info, file.path(here::here(), "vignette-outputs", "data-tables", "regional_ensemble_models_info.csv"))

# save as .html
kableExtra::save_kable(
  model_info_kable, 
  file = file.path(here::here(), "vignette-outputs", "figures", "regional_ensemble_models_info.html"),
  self_contained = TRUE,
  bs_theme = "simplex"
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "regional_ensemble_models_info.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "regional_ensemble_models_info.jpg"),
  zoom = 2
)

# model extents
  

```

```{r ggplot object for map style, eval = TRUE}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(panel.background = element_rect(fill = "lightblue2",
                                        colour = "lightblue2")
        ),
  scale_x_continuous(expand = c(0, 0)),
  scale_y_continuous(expand = c(0, 0)),
  coord_sf()
)

```

## Download countries shapefile

I will need a shapefile of world countries for parts of this analysis, which I will retrieve from [naturalearthdata](https://www.naturalearthdata.com/) using their R package called `rnaturalearth`.

```{r rnaturalearth ne_countries download, eval = TRUE}

# check which types of data are available
# these are in the rnaturalearth package
data(df_layers_cultural) 
# I will use states_provinces

# get metadata
ne_metadata <- ne_find_vector_data(
  scale = 10,
  category = "cultural",
  getmeta = TRUE
) %>%
  dplyr::filter(layer == "admin_0_countries")
# URL to open metadata
utils::browseURL(ne_metadata[, 3])

# if the file isnt already downloaded, download it
if (!file.exists(file.path(here::here(), "data-raw", "ne_countries", "ne_10m_admin_0_countries.shp"))) {
  
  countries_sf <- rnaturalearth::ne_download(
    scale = 10, # highest resolution
    type = "admin_0_countries", # states and provinces
    category = "cultural",
    destdir = file.path(here::here(), "data-raw", "ne_countries"),
    load = TRUE, # load into environment
    returnclass = "sf" # shapefile
  )
  
  # else, just import
  } else {
    
    countries_sf <- sf::read_sf(dsn = file.path(here::here(), "data-raw", "ne_countries", "ne_10m_admin_0_countries.shp"))
  
}



```

# 1. Retrieve K-G climate zones

First, I will use the [kgc](https://cran.r-project.org/web/packages/kgc/index.html) package to retrieve a global raster of the Koppen-Geiger climate zones.

I need to set up a path to my directory and the necessary bioclimatic raster layers.

```{r set wd 1, eval = TRUE}

# path to directory
  mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent")

```

```{r layers to be cropped}

# lists of target files in the directory
# load in bioclim layers to be cropped- the 10km .asc files
env.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\.asc$", full.names = TRUE) %>%
  # extract the 4 bioclim layers I will be using in my models. 
  grep("atc_2015_global.asc|bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)


```

Now I will download the Koppen-Geiger climate zones, which I will use to clip the bioclim layers.

```{r download koppen climate data}
 
# get data
kmz_data <- kgc::kmz

# generate coordinates
kmz_lat <- kgc::genCoords(latlon = "lat", full = TRUE)
kmz_lon <- kgc::genCoords(latlon = "lon", full = TRUE)

# join data and coordinates
kmz_data <- cbind(kmz_lat, kmz_lon) %>%
  cbind(., kmz_data) %>%
  as.data.frame() %>%
  # relocate column
  dplyr::select(kmz_lon, everything()) 



# convert to raster
kmz_data_rast <- terra::rast(
  x = kmz_data,
  type = "xyz",
  crs = "EPSG:4326"
  )
# change name of layer
names(kmz_data_rast) <- "Koeppen_climate_zone"

# validity- minmax should range between 0-32
minmax(kmz_data_rast)

```

```{r retrieve names for zones and make factor}

# retrieve names of zones
kmz_names <- kgc::getZone(1:32)
# make facor levels
kmz_names_factor <- factor(
  x = kmz_names,
  levels = unique(kmz_names),
  ordered = TRUE
)

```

I will need to tidy these data by aggregating them to spatial scale I use for the bioclim variables. I will do this using the mode, because this variable is a set of numbers representing a categorical, so I do not want to average or take the median.

```{r tidy and resample new raster}

# layer used to aggregate raster
agg_layer <- terra::rast(file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "atc_2015_global.asc"))
# raster stats
terra::origin(agg_layer)
terra::crs(agg_layer)

# resample
koppen_zones_rast <- terra::resample(
  x = kmz_data_rast,
  y = agg_layer, # aggregate to this origin and resolution
  method = "mode",
  threads = TRUE,
  filename = file.path(mypath, "models", "working_dir", "koeppen_climate_zones.asc"),
  filetype = "AAIGrid", 
  overwrite = FALSE
    )

```

Now lets plot those zones to make sure this worked.

**NOTE** At some point after I updated R, it stopped automatically encoding UTF-8 special characters. So, I had to replace the umalot "o" with its css code.

```{r visualize, eval = TRUE}

koeppen_zones_rast <- terra::rast(x = file.path(mypath, "models", "working_dir", "koeppen_climate_zones.asc"))
# convert to df
koeppen_zones_rast_plot_layer <- terra::as.data.frame(koeppen_zones_rast, xy = TRUE)


# plot main layer as example
KG_climate_map <- ggplot() +
  # change scale of plots to be standard across figures
  geom_raster(data = koeppen_zones_rast_plot_layer, aes(x = x, y = y, fill = as.factor(koeppen_climate_zones))) +
    labs(
      title = "K\U00F6ppen-Geiger Climate Zones", 
      fill = "K-G climate\nzone"
    ) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  viridis::scale_fill_viridis(
    discrete = TRUE,
    aesthetics = "fill",
    option = "H",
    direction = -1,
    labels = kmz_names_factor
    ) +
  guides(fill = guide_legend(nrow = 3)) +
  theme_classic() +
  theme(
    legend.position = "bottom",
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    ) +
  coord_equal()

```

```{r save K-G plot}

ggsave(
  KG_climate_map, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "koppen_climate_zones.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = 500
)

```

I will also write this raster to a shapefile for easy selection of climate zones.

```{r write to shapefile}

# next, convert raster to polygon
KG_zones_invaded_poly <- terra::as.polygons(
  x = koeppen_zones_rast,
  aggregate = TRUE, # combine cells with the same value into one area
  values = TRUE, # include cell values as attributes
  crs = "EPSG:4326"
) 

gdal(drivers = TRUE)

# write to shapefile
terra::writeVector(
  x = KG_zones_invaded_poly,
  filename = file.path(mypath, "models", "working_dir", "shapefiles", "koeppen_climate_zones.shp"),
  filetype = "ESRI Shapefile",
  overwrite = FALSE
)

```

```{r re-import datasets}

koeppen_zones_rast <- terra::rast(x = file.path(mypath, "models", "working_dir", "koeppen_climate_zones.asc")) 

slf_points <- read_csv(file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2024-08-05.csv")) %>%
  dplyr::select(-species) 

```

# 2. Trim bioclim layers

## 2.1 invaded region in North America (Ri.NAmerica model)

First, I will overlay SLF presence points and use these to select which climate regions to keep. I will use these to create a reference layer. Then, I will use this layer to mask the layers that have already been cropped to N America.

```{r filter presences from invaded range}

# extent object for North America
ext.obj <- terra::ext(-140.976563, -51.064453, 15.182421, 60.586967)

# convert to vector
slf_points_vect <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert back to data frame
slf_points_invaded <- terra::as.data.frame(slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole)) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = "EPSG:4326")

# will not need this object again
rm(slf_points_vect)

```

Now we will intersect the points with the shapefile. I need to crop the shapefile to the the same extent as the presences and then intersect them to keep the climate zones that contain presences. The resulting shapefile will be used to crop the bioclim data and choose background points for that specific model.

### Intersect points and KG zones

```{r intersect slf points with shapefile- Ri.NAmerica}

# load in shapefile again
KG_zones_invaded_poly <- sf::read_sf(dsn = file.path(mypath, "models", "working_dir", "shapefiles", "koeppen_climate_zones.shp")) %>%
  dplyr::rename("koeppen_climate_zones" = "koeppen_cl")

# bounding box for cropping
ext.obj.vect <- c(xmin = -140.976563, ymin = 15.182421, xmax = -51.064453, ymax = 60.586967)
# first, crop K-G raster to N America
KG_zones_invaded_poly <- sf::st_crop(KG_zones_invaded_poly, y = ext.obj.vect) %>%
  # also get rid of ocean
  dplyr::filter(koeppen_climate_zones != 32)




# intersect
regional_invaded_withSLF <- sf::st_filter(x = KG_zones_invaded_poly, y = slf_points_invaded)

# write to file
sf::st_write(
  obj = regional_invaded_withSLF, 
  dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_extent.shp"),
  driver = "ESRI Shapefile",
  append = FALSE
  )

```

```{r re-load K-G zones Ri.NAmerica, eval = TRUE, message = FALSE}

regional_invaded_withSLF <- sf::read_sf(dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_extent.shp")) %>%
  dplyr::rename("koeppen_climate_zones" = "kppn_c_")

# get extent
st_bbox(regional_invaded_withSLF) 

```

Now lets plot it for a sanity check.

```{r plot K-G zones in N America, eval = TRUE}

regional_invaded_withSLF_plot <- ggplot() +
  # world map
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = "azure4", color = "black", lwd = 0.25) +
  # slf native range shapefile
  geom_sf(data = regional_invaded_withSLF, aes(fill = as.factor(koeppen_climate_zones)), color = "black", lwd = 0.25) +
  # scales
  labs(
    title = "K\U00F6ppen-Geiger climate zones as background selection area\nfor 'Ri.NAmerica' model",
    caption = "K-G zones selected if containing SLF presences",
    x = "longitude",
    y = "latitude",
    fill = "K-G climate\nzone"
    ) +
  # scales
  scale_fill_brewer(palette = "Paired", labels = kmz_names_factor) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  # theme
  theme_classic() +
  theme(legend.position = "right") +
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))

```

```{r save invaded K-G plot}

ggsave(
  regional_invaded_withSLF_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "SLF_regional_invaded_extent_KG_climate_zones.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
)

```

So when we mask the bioclim layers, we should see something similar to this.

### Mask bioclim layers

I will start the masking process by importing the bioclim layers and reference layer I need. I will also crop this layer to only N America, so I will need a bounding box for that.

```{r setup for cropping bioclim layers- Ri.NAmerica}

# import layers
env.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_NAmerica.asc$", full.names = TRUE) 

# output file names
output.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_NAmerica.asc$", full.names = FALSE) %>%
  gsub(pattern = "NAmerica", replacement = "regional_invaded_KG")

# also load shapefile again
regional_invaded_withSLF <- terra::vect(file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_extent.shp")) 

# bbox for cropping
ext.obj.crop <- terra::ext(-138.66681, 19.12534, -52.58347, 60.83319 , xy = TRUE)

```

I will now mask these rasters using the shapefile created above.

```{r loop to crop bioclim layers to NAmerica invaded range}
  
# loop to crop extent for all files
for(a in seq_along(env.files)){

  # load in first raster
  rast.hold <- terra::rast(env.files[a])
  
  # mask new rasters using shapefile
  rast.hold <- terra::mask(x = rast.hold, mask = regional_invaded_withSLF, overwrite = FALSE)
      # crop to extent as well
  rast.hold <- terra::crop(x = rast.hold, y = ext.obj.crop, overwrite = FALSE)
  
  #write out the new resampled rasters!
  terra::writeRaster(
    x = rast.hold, 
    filename = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", output.files[a]), 
    filetype = "AAIGrid", 
    overwrite = TRUE
    )
  
  # remove object once its done
  rm(rast.hold)
  
}

# I am pretty sure this method resets the raster cell numbers, which might be annoying downstream....

```

Lets plot one of the new layers to make sure the cropping worked.

```{r plot to ensure cropping worked, eval = TRUE}

bio11_df <- terra::rast(x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "bio11_1981-2010_regional_invaded_KG.asc")) %>%
  # convert to df
  terra::as.data.frame(., xy = TRUE)

# plot main layer as example
(ggplot() +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  # change scale of plots to be standard across figures
  geom_raster(data = bio11_df, 
            aes(x = x, y = y, fill = `CHELSA_bio11_1981-2010_V.2.1`)) +
  xlab("longitude") +
  ylab("latitude") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))
)
# the cropping worked

```

```{r remove objects- Ri.NAmerica}

rm(ext.obj)
rm(ext.obj.vect)
rm(ext.obj.crop)

```

Now that the bioclim layers have been trimmed, I can use these rasters downstream to choose the background points for this model. I will repeat this process for the other two regional models.

## 2.2 Invaded region in South Korea and Japan (Ri.Asia model)

I need to find a way to clip the climate zones by native vs invaded region, because the same climate zones are likely present in Japan and Korea as in the native range. However, I want to segment the data by region. I would normally use the country administrative borders, but there are many contested islands that would be left out. So, I will obtain the bounding box for Japan and South Korea and use that to geographically segment the KG climate zones before masking.

### Intersect points and KG zones

```{r import datasets- Ri.Asia}

koeppen_zones_rast <- terra::rast(x = file.path(mypath, "models", "working_dir", "koeppen_climate_zones.asc")) 

slf_points <- read_csv(file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2024-08-05.csv")) %>%
  dplyr::select(-species) 

```

I just want the presences from the invaded east asian range. I will manually filter out some points that accidentally got into both the native and Ri.Asia models.

```{r filter presences from Ri.Asia range}

# extent object for Japan and SK
ext.obj <- terra::ext(122.93816, 153.98561, 24.21210, 45.52041)

# convert to vector
slf_points_vect <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert to sf object
slf_points_invaded_asian <- terra::as.data.frame(slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole)) %>%
  # filter out specific points that were showing up in both native and invaded_regional model
  dplyr::filter(x > 125) %>%
  # convert to sf object
  sf::st_as_sf(coords = c("x", "y"), crs = "EPSG:4326") 

# will not need this object again
rm(slf_points_vect)

```

Now we will intersect the points with the shapefile. I need to crop the shapefile to the the same extent as the presences and then intersect them to keep the climate zones that contain presences. The resulting shapefile will be used to crop the bioclim data and choose background points for that specific model.

```{r intersect slf points with shapefile- Ri.Asia}

# load in shapefile again
KG_zones_invaded_asian_poly <- sf::read_sf(dsn = file.path(mypath, "models", "working_dir", "shapefiles", "koeppen_climate_zones.shp")) %>%
  dplyr::rename("koeppen_climate_zones" = "koeppen_cl")

# bounding box for cropping
ext.obj.vect <- c(xmin = 122.93816, ymin = 24.21210, xmax = 153.98561, ymax = 45.52041)
# first, crop K-G raster to SK and Japan
KG_zones_invaded_asian_poly <- st_crop(KG_zones_invaded_asian_poly, y = ext.obj.vect) %>%
  # also get rid of ocean
  dplyr::filter(koeppen_climate_zones != 32)




# intersect
regional_invaded_asian_withSLF <- sf::st_filter(x = KG_zones_invaded_asian_poly, y = slf_points_invaded_asian)

# write to file
sf::st_write(
  obj = regional_invaded_asian_withSLF, 
  dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_asian_extent.shp"),
  driver = "ESRI Shapefile",
  append = FALSE
  )

```

```{r re-load K-G zones Ri.Asia, eval = TRUE}

regional_invaded_asian_withSLF <- sf::read_sf(dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_asian_extent.shp")) %>%
  dplyr::rename("koeppen_climate_zones" = "kppn_c_")

# get extent
st_bbox(regional_invaded_asian_withSLF) 

```

Now lets plot it for a sanity check.

```{r plot invaded_asian range, eval = TRUE}

regional_invaded_asian_withSLF_plot <- ggplot() +
  # world map
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = "azure4", color = "black", lwd = 0.25) +
  # slf native range shapefile
  geom_sf(data = regional_invaded_asian_withSLF, aes(fill = as.factor(koeppen_climate_zones)), color = "black", lwd = 0.25) +
  # scales
  labs(
    title = "K\U00F6ppen-Geiger climate zones as background selection area\nfor 'Ri.Asia' model",
    caption = "K-G zones selected if containing SLF presences",
    x = "longitude",
    y = "latitude",
    fill = "K-G climate\nzone"
    ) +
  # scales
  scale_fill_brewer(palette = "Paired", labels = kmz_names_factor) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  # theme
  theme_classic() +
  theme(legend.position = "right") +
  coord_sf(xlim = c(122.93816, 149.74986), ylim = c(30.49986, 46.58424))
  
```

```{r save Ri.Asia K-G plot}

ggsave(
  regional_invaded_asian_withSLF_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "SLF_regional_invaded_asian_extent_KG_climate_zones.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
)

```

### Mask bioclim layers

```{r load in data}

# bioclim layers
env.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) 

# set naming for rasters
output.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = FALSE) %>%
  gsub(pattern = "global", replacement = "regional_invaded_asian_KG")



# load in shapefile as spatVector 
regional_invaded_asian_withSLF <- terra::vect(x = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_asian_extent.shp"))

# for cropping
ext.obj.crop <- terra::ext(122.93816, 30.49986, 149.74986, 46.58424, xy = TRUE)

```

```{r mask}
  
# loop to crop extent for all files
for(a in seq_along(env.files)){

  #ensure that the CRS is consistent
  rast.hold <- terra::rast(env.files[a])
  
  # crop new rasters to extent
  rast.hold <- terra::mask(x = rast.hold, mask = regional_invaded_asian_withSLF)
  # crop to extent as well
  rast.hold <- terra::crop(x = rast.hold, y = ext.obj.crop, overwrite = FALSE)
  
  # write out the new resampled rasters!
  terra::writeRaster(
    x = rast.hold, 
    filename = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", output.files[a]), 
    filetype = "AAIGrid", 
    overwrite = FALSE
    )
  
  # remove object once its done
  rm(rast.hold)
  
}

```

```{r plot for check, eval = TRUE}

bio11_invaded_asian <- terra::rast(
  x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "bio11_1981-2010_regional_invaded_asian_KG.asc")
  ) 

# convert to df
bio11_invaded_asian <- terra::as.data.frame(bio11_invaded_asian, xy = TRUE)

# plot main layer as example
(ggplot() +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  # change scale of plots to be standard across figures
  geom_raster(data = bio11_invaded_asian, aes(x = x, y = y, fill = `CHELSA_bio11_1981-2010_V.2.1`)) +
  xlab("longitude") +
  ylab("latitude") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_sf(xlim = c(122.93816, 149.74986), ylim = c(30.49986, 46.58424))
)
# the cropping worked

```

```{r remove objects- Ri.Asia}

rm(ext.obj)
rm(ext.obj.vect)
rm(ext.obj.crop)

```

Looks good!

## 2.3 Native region in Southeast Asia (Rn model)

The native range for China will be defined as anywhere within Southeast asia that contains the same climate zones that SLF are found in within east and southeast Asia. We will use the `ne_countries` shapefile to clip our KG climate zones shapefile. Then, we will use that final shapefile to clip the bioclim rasters.

### Get shapefile for native range

We will use the `ne_countries` shapefile downloaded earlier to  cut out the parts we want. We can treat the shapefile like any other data table and simply retrieve the countries we want. We will keep all countries in southeast Asia, plus countries in south Asia excluding Pakistan and Afghanistan and east Asia excluding the invaded range in Asia (Japan, SK, NK) (Wikipedia, see references). We are using a much larger extent because there is a lot of conjecture about the true native range of SLF, and we want those climate zones to be considered.

```{r select countries and save}

# re-load the shapefile downloaded earlier
asia_sf <- sf::read_sf(dsn = file.path(here::here(), "data-raw", "ne_countries", "ne_10m_admin_0_countries.shp")) 

# southeast asia shapefile
asia_sf <- asia_sf %>%
  # filter all countries in southeast asia
  dplyr::filter(
    NAME %in% c(
      # southeast asia
      "Brunei", "East Timor", "Indonesia", "Laos", "Malaysia", "Myanmar", "Philippines", "Singapore", "Thailand", "Vietnam", 
      # east asia
      "China", "Hong Kong", "Macau", "Mongolia", "Taiwan", 
      # south asia
      "Bangladesh", "Bhutan", "India", "Maldives", "Nepal", "Sri Lanka")
    ) %>%
  # set crs
  sf::st_transform(x = ., dst = "EPSG:4326") 

# save
sf::st_write(
  obj = asia_sf, 
  dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "asia_admin.shp"),
  driver = "ESRI Shapefile"
  )

```

Macau and Hong Kong did not show up in the original dataset, but were likely considered part of a larger country (in the case of Hong Kong, China).

We will use this shapefile to cut out the KG climate zones, then will intersect the result with the SLF points dataset

### Select KG regions

I will import the SLF presences and KG shapefile first.

```{r import datasets- Rn}

# slf presences
slf_points <- read_csv(file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2024-08-05.csv")) %>%
  dplyr::select(-species) 

# load in KG shapefile 
koeppen_zones_rast <- terra::rast(file.path(mypath, "models", "working_dir", "koeppen_climate_zones.asc")) 

# re-load asia admin shapefile
asia_vect <- terra::vect(file.path(here::here(), "vignette-outputs", "shapefiles", "asia_admin.shp"))

```

Next, I will crop the SLF presences and the KG shapefile datasets using the Asian range shapefile.

```{r crop KG sf by asian sf}

# temp mask layer
mask_layer <- vect(asia_sf)
# convert to vector
slf_points_masked <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
# crop by extent area of interest
terra::mask(., mask = mask_layer) %>%
# convert to geom, which gets coordinates of a spatVector
terra::geom() 

# convert back to data frame
slf_points_native <- terra::as.data.frame(slf_points_masked) %>%
  dplyr::select(-c(geom, part, hole)) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = "EPSG:4326")
  
# will not need these objects again
rm(slf_points_masked)
rm(mask_layer)

```

```{r crop KG raster by asian vector}

# first, mask K-G raster by asian countries
koeppen_zones_rast_native <- terra::mask(koeppen_zones_rast, mask = asia_vect) 

# next, convert raster to polygon
KG_zones_native_poly <- terra::as.polygons(
  x = koeppen_zones_rast_native,
  aggregate = TRUE,
  values = TRUE,
  crs = "EPSG:4326"
) 

# write to shapefile
terra::writeVector(
  x = KG_zones_native_poly,
  filename = file.path(mypath, "models", "working_dir", "shapefiles", "asia_admin_KG.shp"),
  filetype = "ESRI Shapefile",
  overwrite = TRUE
)

```

```{r intersect and save}

# load back in as shapefile
KG_zones_native_poly <- sf::read_sf(dsn = file.path(mypath, "models", "working_dir", "shapefiles", "asia_admin_KG.shp")) %>%
  # rename
  dplyr::rename("koeppen_climate_zones" = "koeppen_cl") %>%
  # also get rid of ocean
  dplyr::filter(koeppen_climate_zones != 32) 
  
# intersect
regional_native_withSLF <- sf::st_filter(x = KG_zones_native_poly, y = slf_points_native)


# write to file
sf::st_write(
  obj = regional_native_withSLF, 
  dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_native_extent.shp"),
  driver = "ESRI Shapefile",
  append = FALSE
  )

```

```{r re-load shapefile as needed, eval = TRUE}

regional_native_withSLF <- sf::read_sf(dsn = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_native_extent.shp")) %>%
  dplyr::rename("koeppen_climate_zones" = "kppn_c_")

# get extent
st_bbox(regional_native_withSLF) 

```

Now lets plot it for a sanity check.

```{r plot china, eval = TRUE}

regional_native_withSLF_plot <- ggplot() +
  # world map
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = "azure4", color = "black", lwd = 0.25) +
  # slf native range shapefile
  geom_sf(data = regional_native_withSLF, aes(fill = as.factor(koeppen_climate_zones)), color = "black", lwd = 0.25) +
  # scales
  labs(
    title = "K\U00F6ppen climate zones as background selection area for 'Rn' model",
    caption = "K-G zones selected if containing SLF presences",
    x = "longitude",
    y = "latitude",
    fill = "K-G climate\nzone"
    ) +
  # scales
  scale_fill_brewer(palette = "Paired", labels = kmz_names_factor) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  # theme
  theme_classic() +
  theme(legend.position = "right") +
  coord_sf(xlim = c(72.666527, 140.999860), ylim = c(-9.666806, 53.249861))

```

```{r save Rn K-G plot}

ggsave(
  regional_native_withSLF_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "SLF_native_extent_KG_climate_zones.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
)

```

### Mask bioclim layers

```{r setup for cropping bioclim layers- Rn}

# import layers
env.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) 

# output file names
output.files <- list.files(path = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = FALSE) %>%
  gsub(pattern = "global", replacement = "regional_native_KG")

# also load shapefile again
regional_native_withSLF <- terra::vect(file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_native_extent.shp")) 

ext.obj.crop <- terra::ext(72.666527, -9.666806, 140.999860, 53.249861, xy = TRUE)

```

I will now mask these rasters using the shapefile created above

```{r loop to crop bioclim layers to Rn range}

# loop to crop extent for all files
for(a in seq_along(env.files)){

  #ensure that the CRS is consistent
  rast.hold <- terra::rast(env.files[a])
  
  # mask new rasters using shapefile
  rast.hold <- terra::mask(x = rast.hold, mask = regional_native_withSLF, overwrite = FALSE)
  # crop to extent as well
  rast.hold <- terra::crop(x = rast.hold, y = ext.obj.crop, overwrite = FALSE)
  
  #write out the new resampled rasters!
  terra::writeRaster(
    x = rast.hold, 
    filename = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", output.files[a]), 
    filetype = "AAIGrid", 
    overwrite = FALSE
    )
  
  # remove object once its done
  rm(rast.hold)
  
}
# I am pretty sure this method resets the raster cell numbers, which might be annoying downstream....

```

Lets plot one of the new layers to make sure the cropping worked.

```{r plot main_layer to ensure cropping worked, eval = TRUE}

bio11_native <- terra::rast(x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "bio11_1981-2010_regional_native_KG.asc")) 

# convert to df
bio11_native <- terra::as.data.frame(bio11_native, xy = TRUE)

# plot main layer as example
ggplot() +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  # change scale of plots to be standard across figures
  geom_raster(data = bio11_native, aes(x = x, y = y, fill = `CHELSA_bio11_1981-2010_V.2.1`)) +
  xlab("longitude") +
  ylab("latitude") +
  theme_minimal() +
  theme(legend.position = "none") +
  coord_sf(xlim = c(72.666527, 140.999860), ylim = c(-9.666806, 53.249861))

# the cropping worked

```

```{r remove objects- Rn}

rm(ext.obj.crop)

```

# 3. Choose background points from selection areas

Now, I need to use the rasters I just created to choose the background points, which are used by MaxEnt to sample the climate and compare that sample with known SLF presences. The background points should be randomized because they are a stand-in for true absences, so I will generate them using `dismo::randomPoints()`.

```{r set wd 2, eval = TRUE}

mypath <- file.path(here::here() %>% 
                     dirname(),
                   "maxent/historical_climate_rasters/chelsa2.1_30arcsec/v1_maxent_10km")

```

```{r load plotting base layer, eval = TRUE}

global_bio11_df <- terra::rast(x = file.path(mypath, "bio11_1981-2010_global.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

```


## 3.1 invaded region of North America (Ri.NAmerica model)

```{r load datasets- Ri.NAmerica}

# load in raster to choose points from
regional_invaded_bio11 <- raster::raster(x = file.path(mypath, "bio11_1981-2010_regional_invaded_KG.asc")) 

# load in slf points to intersect
slf_points <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05.rds")) %>%
  dplyr::select(-species) %>%
  as.data.frame()

```

Create 10,000 random background points. These will be corrected for the latitudinal stretching and will not be selected from presence locations, like with the global model.

```{r random points- Ri.NAmerica}

# set seed
set.seed(5)

# generate random points 
regional_invaded_background <- dismo::randomPoints(
  mask = regional_invaded_bio11, 
  n = 10000, # default number used by maxent
  p = slf_points,
  excludep = TRUE, # exclude cells where slf has been found
  lonlatCorrection = TRUE, # weight samples by latitude because cell size is larger closer to equator
  warn = 2 # higher number gives most warnings, including if sample size not reached
  ) %>%
  as.data.frame(.)

# save as csv
write_csv(
  x = regional_invaded_background, 
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_background_points_v7.csv")
  )
# save as rds file
write_rds(
  regional_invaded_background, 
  file = file.path(here::here(), "data", "regional_invaded_background_points_v7.rds")
)


```

Now lets visualize the result.

```{r import plotting datasets and reproject to equal area, eval = TRUE}

# rasters
# bg raster
regional_invaded_bio11_df <- terra::rast(x = file.path(mypath, "bio11_1981-2010_regional_invaded_KG.asc")) %>%
  terra::as.data.frame(., xy = TRUE)
  
# re-read background points 
regional_invaded_background <- read.csv(file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_background_points_v7.csv")) 

```

```{r plot Ri.NAmerica background, eval = TRUE}

regional_invaded_background_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  # base layer
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  # raster of interest
  geom_raster(data = regional_invaded_bio11_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = regional_invaded_background, aes(x = x, y = y), color = "firebrick2", size = 0.25) +
  ggtitle("'Ri.NAmerica' model background points") +
  # same coords used to cut other N America plots
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))

```

We can see that the background points are isolated to the area selected using the K-G climate zones in North America, so our operation was successful.

```{r save Ri.Namerica background plot}

ggsave(
  regional_invaded_background_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_model_background_points_v7.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

Now, I will repeat this process for the next two models.

## 3.2 invaded region of Asia (South Korea and Japan, Ri.Asia model)

```{r load datasets- Ri.Asia}

# load in raster to choose points from
invaded_asian_bio11 <- raster::raster(x = file.path(mypath, "bio11_1981-2010_regional_invaded_asian_KG.asc")) 

# load in slf points to intersect
slf_points <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05.rds")) %>%
  dplyr::select(-species) %>%
  as.data.frame()

```

```{r random points- Ri.Asia}

# set seed
set.seed(7)
  
# generate random points 
regional_invaded_asian_points <- dismo::randomPoints(
  mask = invaded_asian_bio11, # cropped to regional_invaded_asian range
  n = 10000, # default number used by maxent
  p = slf_points,
  excludep = TRUE, # exclude cells where slf has been found
  lonlatCorrection = TRUE, # weight samples by latitude because cell size is larger closer to equator
  warn = 2 # higher number gives most warnings, including if sample size not reached
  ) %>%
  as.data.frame(.)

# save as csv
write_csv(x = regional_invaded_asian_points, file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_background_points_v2.csv"))
# save as rds file
write_rds(regional_invaded_asian_points, file = file.path(here::here(), "data", "regional_invaded_asian_background_points_v2.rds"))


```

Now lets visualize the result.

```{r load in data for plotting and change projection- Ri.Asia, eval = TRUE}

# background raster
invaded_asian_bio11_df <- terra::rast(x = file.path(mypath, "bio11_1981-2010_regional_invaded_asian_KG.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

# bg points
regional_invaded_asian_points <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_background_points_v2.csv")) 

```

```{r plot Ri.Asia bg, eval = TRUE}

regional_invaded_asian_points_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = invaded_asian_bio11_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = regional_invaded_asian_points, aes(x = x, y = y), color = "firebrick2", size = 0.5) +
  ggtitle("'Ri.Asia' model background points") +
  coord_sf() +
  scale_x_continuous(limits = c(122.93816, 149.74986), expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), expand = c(0, 0))

```

```{r save plot- Ri.Asia bg}

ggsave(
  regional_invaded_asian_points_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_asian_model_background_points_v2.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

## 3.3 Native region (Rn model)

```{r load in datasets}

# mask layer I will use
regional_native_bio2 <- raster::raster(x = file.path(mypath, "bio2_1981-2010_regional_native_KG.asc")) # %>%
  #!is.na()

# also get ext
terra::ext(regional_native_bio2)


# load in slf points to intersect
slf_points <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05.rds")) %>%
  dplyr::select(-species) %>%
  as.data.frame()

```

```{r background points for global models}

# set seed so that the random points for this dataset are the same the next time this code is run
set.seed(6)
  
# generate random points 
regional_native_points <- dismo::randomPoints(
  mask = regional_native_bio2, 
  n = 10000, # default number used by maxent
  p = slf_points,
  excludep = TRUE, # exclude cells where slf has been found
  lonlatCorrection = TRUE, # weight samples by latitude because cell size is larger closer to equator
  warn = 2 # higher number gives most warnings, including if sample size not reached
  ) %>%
  as.data.frame(.)

# save as csv
write_csv(x = regional_native_points, 
          file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_background_points_v3.csv")
          )
# save as rds file
write_rds(regional_native_points, 
          file = file.path(here::here(), "data", "regional_native_background_points_v3.rds")
)
  
  
```

```{r load in files for plotting and change projection- Rn, eval = TRUE}

regional_native_bio2_df <- terra::rast(x = file.path(mypath, "bio2_1981-2010_regional_native_KG.asc")) %>%
  terra::as.data.frame(., xy = TRUE)


regional_native_points <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_background_points_v3.csv")) 

```

```{r plot points, eval = TRUE}

regional_native_points_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = regional_native_bio2_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = regional_native_points, aes(x = x, y = y), color = "firebrick2", size = 0.25) +
  ggtitle("'Rn' model background points") +
  coord_sf(xlim = c(72.666527, 140.999860), ylim = c(-9.666806, 53.249861))

```

```{r save plot- Rn bg}

ggsave(
  regional_native_points_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_native_model_background_points_v3.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

# 4. Choose and plot training points

Next, I will use bounding boxes to isolate SLF presences from each major region. This would not normally be sufficient to select the correct points, but each region is isolated from the others in a fashion that a bounding box easily separates the points by region in this case.

```{r set wd 4, eval = TRUE}

mypath <- file.path(here::here() %>% 
                     dirname(),
                   "maxent")

```

```{r read in data}

# background raster for plotting will be same as in last section

# load in slf points to intersect
slf_points <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05.rds")) %>%
  dplyr::select(-species) 

```

## 4.1 Invaded region of North America (Ri.NAmerica model)

I will convert the SLF points dataset to a spatial vector object so that it can be filtered using a bounding box, then transform it back to a dataframe.

```{r filter presences from Ri.NAmerica range}

# extent object for N America
ext.obj <- terra::ext(-96.503906, -59.589844, 23.5, 47.457809)

# convert to vector
slf_points_vect <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
  # crop by extent area of interest
  terra::crop(., y = ext.obj) %>%
  # convert to geom, which gets coordinates of a spatVector
  terra::geom() 

# convert back to data frame
slf_points_invaded <- terra::as.data.frame(slf_points_vect) %>%
  dplyr::select(-c(geom, part, hole))

# will not need this object again
rm(slf_points_vect)

```

```{r save presences- Ri.NAmerica}

write_csv(
  x = slf_points_invaded,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_train_slf_presences_v7.csv")
)

```

Now I will plot these points on the raster to ensure that the bounding box worked.

```{r plot model presences- Ri.NAmerica, eval = TRUE}

# load in raster
regional_invaded_df <- terra::rast(x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "atc_2015_regional_invaded_KG.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

# re-import point data
slf_points_invaded <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_train_slf_presences_v7.csv")) 

# plot
regional_invaded_slf_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = regional_invaded_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_invaded, aes(x = x, y = y), color = "darkorange", size = 0.5) +
  labs(title = "'Ri.NAmerica' model SLF presences") +
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))
  
```

```{r save Ri.NAmerica presence plot}

ggsave(
  regional_invaded_slf_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_train_slf_presences_v7.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

## 4.2 Invaded Regional (South Korea and Japan)

```{r filter presences from invaded_asian range}

# native range extent shapefile used to select slf presences
mask_layer <- terra::vect(x = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_regional_invaded_asian_extent.shp"))

# convert to vector
slf_points_masked <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
# crop by extent area of interest
terra::mask(., mask = mask_layer) %>%
# convert to geom, which gets coordinates of a spatVector
terra::geom() 

# convert back to data frame
slf_points_invaded_asian <- terra::as.data.frame(slf_points_masked) %>%
  dplyr::select(-c(geom, part, hole)) %>%
    dplyr::filter(x > 125)
  
# will not need this object again
rm(slf_points_masked)

```

```{r save presences- Ri.Asia}

write_csv(
  x = slf_points_invaded_asian,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_train_slf_presences_v2.csv")
)

```

```{r plot model presences- Ri.Asia, eval = TRUE}

invaded_asian_df <- terra::rast(
  x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "atc_2015_regional_invaded_asian_KG.asc")
  ) %>%
  terra::as.data.frame(., xy = TRUE)

# re-import point data
slf_points_invaded_asian <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_train_slf_presences_v2.csv")) 

# plot
regional_invaded_asian_slf_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = invaded_asian_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_invaded_asian, aes(x = x, y = y), color = "darkorange", size = 1) +
  labs(title = "'Ri.Asia' model SLF presences") +
  coord_sf() +
  scale_x_continuous(limits = c(122.93816, 149.74986), expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), expand = c(0, 0))

```

```{r save Ri.Asia presence plot}

ggsave(
  regional_invaded_asian_slf_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_asian_train_slf_presences_v2.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

## 4.3 Native Regional

```{r filter presences from native range}

# native range extent shapefile used to select slf presences
mask_layer <- terra::vect(x = file.path(here::here(), "vignette-outputs", "shapefiles", "SLF_native_range_extent.shp"))

# convert to vector
slf_points_masked <- terra::vect(x = slf_points, geom = c("x", "y"), crs = "EPSG:4326") %>%
# crop by extent area of interest
terra::mask(., mask = mask_layer) %>%
# convert to geom, which gets coordinates of a spatVector
terra::geom() 

# convert back to data frame
slf_points_native <- terra::as.data.frame(slf_points_masked) %>%
  dplyr::select(-c(geom, part, hole))
  
# will not need this object again
rm(slf_points_masked)
rm(mask_layer)

```

```{r save presences- Rn}

write_csv(
  x = slf_points_native,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_train_slf_presences_v3.csv")
)

```

```{r plot model presences- Rn, eval = TRUE}

# raster
regional_native_df <- terra::rast(
  x = file.path(mypath, "historical_climate_rasters", "chelsa2.1_30arcsec", "v1_maxent_10km", "atc_2015_regional_native_KG.asc")
  ) %>%
  terra::as.data.frame(., xy = TRUE)

# re-import point data
slf_points_native <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_train_slf_presences_v3.csv"))


# plot
regional_native_slf_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = regional_native_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_native, aes(x = x, y = y), color = "darkorange", size = 1) +
  labs(title = "'Rn' model SLF presences") +
  coord_sf(xlim = c(72.666527, 140.999860), ylim = c(-9.666806, 53.249861))
  
```

```{r save Rn presence plot}

ggsave(
  regional_native_slf_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_native_train_slf_presences_v3.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

# 5. Choose validation points

I will train each regional-scale on all of its regional slf presences and test based on another region's points. Here are the models and their training / validation (test) regions: `r model_info_kable`

The training and test points should be geographically separated, but this creates an issue of the proper ratios being kept between the training and test points if I were to use the entire dataset from training one model to test another model. So, I will randomly sample a subset of each model's training region as the test points for another model.

```{r select test points for Ri.NAmerica model}

# this model is trained on N American presences and tested on invaded asian presences

set.seed(8)

# randomly select subset of rows for testing
slf_points_invaded_test_rows <-  sample(nrow(slf_points_invaded_asian), size = 68, replace = FALSE) # no replacement
# filter
slf_points_invaded_test <- slf_points_invaded_asian[slf_points_invaded_test_rows, ]

# write to csv
write_csv(
  x = slf_points_invaded_test,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_test_slf_presences_v7.csv")
)


```

```{r select test points for Ri.Asia model}

# this model is trained on invaded asian presences and tested on invaded N American presences

set.seed(9)

# randomly select subset of rows for testing
slf_points_invaded_asian_test_rows <-  sample(nrow(slf_points_invaded), size = 40, replace = FALSE) # no replacement
# filter
slf_points_invaded_asian_test <- slf_points_invaded[slf_points_invaded_asian_test_rows, ]

# write to csv
write_csv(
  x = slf_points_invaded_asian_test,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_test_slf_presences_v2.csv")
)


```

```{r select test points for Rn model}

# this model is trained on Asian native presences and tested on invaded N. American presences

set.seed(10)

# randomly select subset of rows for testing
slf_points_native_test_rows <-  sample(nrow(slf_points_invaded), size = 54, replace = FALSE) # no replacement
# filter
slf_points_native_test <- slf_points_invaded[slf_points_native_test_rows, ]

# write to csv
write_csv(
  x = slf_points_native_test,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_test_slf_presences_v3.csv")
)


```

Now we are ready to run all 3 regional models!

# References

1. Calvin, D. D., Rost, J., Keller, J., Crawford, S., Walsh, B., Bosold, M., & Urban, J. (2023). Seasonal activity of spotted lanternfly (Hemiptera: Fulgoridae), in Southeast Pennsylvania. Environmental Entomology, 52(6), 11081125. https://doi.org/10.1093/ee/nvad093

2. Du, Z., Wu, Y., Chen, Z., Cao, L., Ishikawa, T., Kamitani, S., Sota, T., Song, F., Tian, L., Cai, W., & Li, H. (2021). Global phylogeography and invasion history of the spotted lanternfly revealed by mitochondrial phylogenomics. Evolutionary Applications. https://doi.org/10.1111/eva.13170

3. Gallien, L., Douzet, R., Pratte, S., Zimmermann, N. E., & Thuiller, W. (2012). Invasive species distribution models  how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography, 21(11), 11261136. https://doi.org/10.1111/j.1466-8238.2012.00768.x

4. Lewkiewicz, S. M., De Bona, S., Helmus, M. R., & Seibold, B. (2022). Temperature sensitivity of pest reproductive numbers in age-structured PDE models, with a focus on the invasive spotted lanternfly. Journal of Mathematical Biology, 85(3), 29. https://doi.org/10.1007/s00285-022-01800-9

5. Santana Jr, P. A., Kumar, L., Da Silva, R. S., Pereira, J. L., & Picano, M. C. (2019). Assessing the impact of climate change on the worldwide distribution of Dalbulus maidis (DeLong) using MaxEnt. Pest Management Science, 75(10), 27062715. https://doi.org/10.1002/ps.5379

6. East Asia. 2024, March 6. Wikipedia.

7. South Asia. 2024, March 27. Wikipedia.

8. Southeast Asia. 2024, March 27. Wikipedia.

9. Xin, B., Zhang, Y., Wang, X., Cao, L., Hoelmer, K. A., Broadley, H. J., & Gould, J. R. (2020). Exploratory survey of spotted lanternfly (Hemiptera: Fulgoridae) and its natural enemies in China. Environmental Entomology, 50(1), 3645. https://doi.org/10.1093/ee/nvaa137

# Appendix {.hidden}

## Patchwork together figures for publication

```{r reference layer 5}

mypath <- file.path(here::here() %>% 
                     dirname(),
                   "maxent/historical_climate_rasters/chelsa2.1_30arcsec/v1_maxent_10km")


# reference layer for background of test plots
global_bio11_df <- terra::rast(x = file.path(mypath, "bio11_1981-2010_global.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

```

### Ri.NAmerica model

These chunks will splice together the plots from above as summary plots.

```{r load necessary datasets- Ri.NAmerica}

# validation dataset
slf_points_invaded_test <- read.csv(
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_test_slf_presences_v7.csv")
  )

# raster for the validation plot
invaded_asian_df <- terra::rast(x = file.path(mypath, "atc_2015_regional_invaded_asian_KG.asc")) %>%
  terra::as.data.frame(., xy = TRUE)

```

```{r plot validation region for Ri.NAmerica model}

regional_invaded_test_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = invaded_asian_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_invaded_test, aes(x = x, y = y), color = "darkorange", size = 1) +
  labs(title = "'Ri.NAmerica' model validation (test) presences") +
  coord_sf() +
  scale_x_continuous(limits = c(122.93816, 149.74986), expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), expand = c(0, 0))

```

```{r Ri.NAmerica model elements}

# labels
x_scale_labels <- seq(from = -120, to = -60, by = 20)
y_scale_labels <- seq(from = 20, to = 60, by = 5)

# background selection area
regional_invaded_withSLF_plot <- regional_invaded_withSLF_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(
    tag = "C",
    fill = "zone"
    ) +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


# background points
regional_invaded_background_plot <- regional_invaded_background_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


# training presences
regional_invaded_slf_plot  <- regional_invaded_slf_plot +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


# new labels
x_scale_labels <- seq(from = 125, to = 145, by = 5)
y_scale_labels <- seq(from = 32, to = 46, by = 2)

# test/validation presences
regional_invaded_test_plot <- regional_invaded_test_plot +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(limits = c(122.93816, 149.74986), labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )

# patchwork plot
regional_invaded_patchwork <- (
  (regional_invaded_slf_plot + regional_invaded_background_plot) / 
  (regional_invaded_withSLF_plot + regional_invaded_test_plot) 
  ) +
  plot_annotation(
    title = "'Ri.NAmerica' model elements",
    subtitle = "(A) training presences (B) background points (C) background selection area (D) test presences"
    ) 

```

```{r save plot Ri.NAmerica patchwork}

ggsave(
  regional_invaded_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_model_v7_summary.jpg"),
  height = 10, 
  width = 14,
  device = jpeg,
  dpi = 500
  )

```

### Ri.Asia Model

```{r load necessary datasets- Ri.Asia}

# validation dataset
slf_points_invaded_asian_test <- read.csv(
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_test_slf_presences_v2.csv")
  )

# load in raster for validation plot
regional_invaded <- terra::rast(
  x = file.path(mypath, "atc_2015_regional_invaded_KG.asc")
  ) %>%
  terra::as.data.frame(., xy = TRUE)

```

```{r plot validation region for Ri.Asia model}

regional_invaded_asian_test_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = regional_invaded_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_invaded_asian_test, aes(x = x, y = y), color = "darkorange", size = 1) +
  labs(title = "'Ri.Asia' model test presences") +
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))

```

```{r regional_invaded_asian}

# labels
x_scale_labels <- seq(from = 125, to = 145, by = 5)
y_scale_labels <- seq(from = 32, to = 46, by = 2)

# background selection area
regional_invaded_asian_withSLF_plot <- regional_invaded_asian_withSLF_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(
    tag = "C",
    fill = "zone"
    ) +
  scale_x_continuous(limits = c(122.93816, 149.74986), labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


# background points
regional_invaded_asian_points_plot <- regional_invaded_asian_points_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  scale_x_continuous(limits = c(122.93816, 149.74986), labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


# training presences
regional_invaded_asian_slf_plot  <- regional_invaded_asian_slf_plot +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  scale_x_continuous(limits = c(122.93816, 149.74986), labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(limits = c(30.49986, 46.58424), labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )

# labels
x_scale_labels <- seq(from = -120, to = -60, by = 20)
y_scale_labels <- seq(from = 20, to = 60, by = 5)


# test presences
regional_invaded_asian_test_plot <- regional_invaded_asian_test_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )



regional_invaded_asian_patchwork <- (
  (regional_invaded_asian_slf_plot + regional_invaded_asian_points_plot) / 
  (regional_invaded_asian_withSLF_plot + regional_invaded_asian_test_plot) 
  ) +
  plot_annotation(
    title = "'Ri.Asia' model elements",
    subtitle = "(A) training presences (B) background points (C) background selection area (D) test presences"
    )

```

```{r save Ri.Asia patchwork plot}

ggsave(
  regional_invaded_asian_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_invaded_asian_model_v2_summary.jpg"),
  height = 10, 
  width = 12,
  device = jpeg,
  dpi = 500
  )

```

### Rn model

```{r load necessary datasets- Rn}

# validation dataset
slf_points_native_test <- read.csv(
  file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_test_slf_presences_v3.csv")
  ) 

# load in raster for validation plot
regional_invaded <- terra::rast(
  x = file.path(mypath, "atc_2015_regional_invaded_KG.asc")
  ) %>%
  terra::as.data.frame(., xy = TRUE)

```

```{r plot validation region for Rn model}

regional_native_test_plot <- ggplot() +
  map_style +
  # empty world map to force same aspect ratio
  geom_sf(data = countries_sf, aes(geometry = geometry), fill = NA, color = NA) +
  geom_raster(data = global_bio11_df, aes(x = x, y = y), fill = "azure4") +
  geom_raster(data = regional_invaded_df, aes(x = x, y = y), fill = "azure1") +
  geom_point(data = slf_points_native_test, aes(x = x, y = y), color = "darkorange", size = 1) +
  labs(title = "'Rn' model test presences") +
  coord_sf(xlim = c(-138.66681, -52.58347), ylim = c(19.12534, 60.83319))

```

Patchwork

```{r regional_native}

x_scale_labels <- seq(from = 80, to = 140, by = 10)
y_scale_labels <- seq(from = 0, to = 50, by = 10)

# background selection area
regional_native_withSLF_plot <- regional_native_withSLF_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    plot.tag.location = "panel",
    plot.tag.position = c(0.1, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
    ) +
  labs(
    tag = "C",
    fill = "zone"
    ) +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) 


# background points
regional_native_points_plot <- regional_native_points_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    plot.tag.location = "panel",
    plot.tag.position = c(0.1, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
    ) +
  labs(tag = "B") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) 

  
# training presences
regional_native_slf_plot  <- regional_native_slf_plot +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    plot.tag.location = "panel",
    plot.tag.position = c(0.1, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
    ) +
  labs(tag = "A") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) 



# labels
x_scale_labels <- seq(from = -120, to = -60, by = 20)
y_scale_labels <- seq(from = 20, to = 60, by = 5)

# testing presences
regional_native_test_plot <- regional_native_test_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(labels = x_scale_labels, breaks = x_scale_labels, expand = c(0, 0)) +
  scale_y_continuous(labels = y_scale_labels, breaks = y_scale_labels, expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


regional_native_patchwork <- (
  (regional_native_slf_plot | regional_native_points_plot) / 
  (regional_native_withSLF_plot | regional_native_test_plot)
   ) +
  plot_annotation(
    title = "'Rn' model elements",
    subtitle = "(A) training presences (B) background points (C) background selection area (D) test presences"
    ) +
  plot_layout(widths = unit(20, "cm")) 

```

```{r save Rn patchwork plot}

ggsave(
  regional_native_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "regional_native_model_v3_summary.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```
