---
title: "Run an Extrapolation Detection (NT2) analysis and calculate MIC"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University\n[Center for Biodiversity](http://www.biodiversitycenter.org/)\n[iEco Lab](https://www.iecolab.org/)\nsam.owens@temple.edu]"
date: "2024-10-18"
---

In this vignette, I will run an ExDet analysis for each regional-scale model and calculate the most important covariate based on each model. The ExDet will be used for two main purposes:

1.  This metric will allow us to quantify the level of extrapolation that is happening with our models as they are being projected to new locations and time periods. An ExDet value below 0 or above 1 indicates extrapolation. Below 0 indicates a type 1 extrapolation within a single covariates, while a value above 1 indicates a type 2 extrapolation between multiple covariates.
2.  We will weight each regional model using the ExDet measurement in preparation for an ensemble model. This way, our predictions are weighted by our level of confidence and the level of climate analogy.

I will specifically use the background points that were used to train each model to perform the ExDet analysis. By doing this, I am effectively testing whether the covariates in other locations are significantly different from our model. This analysis will be performed for both the historical and climate change versions of the rasters.

In addition to the ExDet analysis, I will be performing an MIC (Most Influential Covariate) analysis. This analysis reveals which covariate is causing the extrapolation in the ExDet, if present.

Both of these analyses will be done via the `compute_extrapolation()` function within the `dsmextra` package.

# Setup

```{r load necesssary packages, message = FALSE}

library(tidyverse)

library(here) 
# here() is set at the root folder of this package

# spatial data handling
library(terra)
library(raster)
# remotes::install_github("densitymodelling/dsmextra")
#library(dsmextra) # ExDet and MIC
## this package is used but we will be retiring it soon as part of an update, # out for now

# aes
library(viridis)
library(ggnewscale)
library(patchwork)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

This is a styling object for the exdet plots.

```{r map style exdet, eval = TRUE}

map_style_exdet <- list(
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "azure3",
                                        colour = "azure3")
  ),
  scale_x_continuous(expand = c(0, 0)),
  scale_y_continuous(expand = c(0, 0)),
  labs(
    x = "longitude",
    y = "latitude"
  ),
  coord_equal()
)

# color scheme taken from: https://colorbrewer2.org/?type=diverging&scheme=RdBu&n=11

```

First, I will outline this helper function to format all figure numbers to no more than 2 decimal places.

```{r helper function for plot aesthetics, eval = TRUE}

format_decimal <- function(x) as.numeric(sprintf("%.*f", 2, x))

```

This next helper function will make mapping of our exdet results more consistent, as the scale is rather complex. This function includes the above helper.

```{r exdet map function, eval = TRUE}

# raster should be the raster resulting from the exdet analysis
# raster.df should be the dataframe version of this raster.

map_exdet <- function(raster, raster.df) {
  
  ggplot() +
    # default plot style
    map_style_exdet +
    # plot regular raster of values first
    geom_raster(data = raster.df, aes(x = x, y = y, fill = ExDet)) +
    # fill scale
    scale_fill_gradientn(
      name = "ExDet Score",
      # the color scale used- must be 1 color per break point, with at least 1 extra color between (1, 3, 5 and 7 are the break point colors- see breaks)
      colors = c("#67001f", "#f4a582", "white", "white", "white", "#92c5de", "#053061"),
      # the scale values themselves
      values = scales::rescale(x = c(
        min(values(raster), na.rm = TRUE), 
        0, 1, 
        max(values(raster), na.rm = TRUE)
        )),
      # the breaks indicating a change in color
      breaks = c(
        format_decimal(min(values(raster), na.rm = TRUE)), 
        0, 1, 
        format_decimal(max(values(raster), na.rm = TRUE))
        ),
      # the limits of the color scale
      limits = c(
        format_decimal(min(values(raster), na.rm = TRUE)), 
        format_decimal(max(values(raster), na.rm = TRUE))
        ),
      guide = guide_colorbar(frame.colour = "black", ticks.colour = "black", barwidth = 20, order = 1)
      ) 
  
}

  
```

Finally, the style for the MIC maps.

```{r map style MIC, eval = TRUE}

map_style_MIC <- list(
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "azure3",
                                        colour = "azure3")
  ),
  scale_x_continuous(expand = c(0, 0)),
  scale_y_continuous(expand = c(0, 0)),
  labs(
    x = "longitude",
    y = "latitude"
  ),
  coord_equal()
)

# color scheme taken from: https://colorbrewer2.org/?type=diverging&scheme=RdBu&n=11

```

## Import datasets

First, I will import and tidy the global bioclim layers for 1981-2010 and 2041-2070.

```{r load in historical covariates}

# path to directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")

# the env covariates used to train the model global
x_global_hist_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent_10km"), pattern = "\\.asc$", full.names = TRUE) %>%
    grep("bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

```

```{r load in CMIP6 covariates}

# path to directory
  mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/future_climate_rasters/chelsa2.1_30arcsec")

# the env covariates for performing xy predictions for global slf and IVR points

# SSP126
x_global_126_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp126_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# SSP370
x_global_370_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp370_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# SSP585
x_global_585_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp585_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

```

I will unify the variable naming and import the rasters.

```{r naming object, eval = TRUE}

# layer name object. Check order of layers first
env_layer_names <- c("bio11", "bio12", "bio15", "bio2")

```

```{r stack historical covariates and make naming consistent}

# stack env covariates
x_global_hist_env_covariates <- terra::rast(x = x_global_hist_env_covariates_list)

# attributes
nlyr(x_global_hist_env_covariates)
names(x_global_hist_env_covariates)
minmax(x_global_hist_env_covariates)
# ext(x_global_hist_env_covariates)
# crs(x_global_hist_env_covariates)


# I will change the name of the variables because they are throwing errors in SDMtune
names(x_global_hist_env_covariates) <- env_layer_names
# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)

```

```{r stack CMIP6 covariates, message = FALSE}

# SSP126
x_global_126_env_covariates <- terra::rast(x = x_global_126_env_covariates_list)

# attributes
nlyr(x_global_126_env_covariates)
names(x_global_126_env_covariates)
minmax(x_global_126_env_covariates)
# ext(x_global_126_env_covariates)
# crs(x_global_126_env_covariates)

names(x_global_126_env_covariates) <- env_layer_names


# SSP370
x_global_370_env_covariates <- terra::rast(x = x_global_370_env_covariates_list)

# attributes
nlyr(x_global_370_env_covariates)
names(x_global_370_env_covariates)
minmax(x_global_370_env_covariates)
# ext(x_global_370_env_covariates)
# crs(x_global_370_env_covariates)

names(x_global_370_env_covariates) <- env_layer_names


# SSP585
x_global_585_env_covariates <- terra::rast(x = x_global_585_env_covariates_list)

# attributes
nlyr(x_global_585_env_covariates)
names(x_global_585_env_covariates)
minmax(x_global_585_env_covariates)
# ext(x_global_585_env_covariates)
# crs(x_global_585_env_covariates)

names(x_global_585_env_covariates) <- env_layer_names

```

```{r remove lists}

rm(x_global_hist_env_covariates_list)
rm(x_global_126_env_covariates_list)
rm(x_global_370_env_covariates_list)
rm(x_global_585_env_covariates_list)

```

I also need the background point sets that were used to train each regional model. These copies need to include the values of each raster at those particular locations.

```{r background point sets, eval = TRUE}

# regional_native
regional_native_background <- read_csv(
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_background_points_with_data_v3.csv")
  ) %>%
  as.data.frame()


# regional_invaded
regional_invaded_background <- read_csv(
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_background_points_with_data_v7.csv")
  ) %>%
  as.data.frame()


# regional_invaded_asian
regional_invaded_asian_background <- read_csv(
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_background_points_with_data_v2.csv")
  ) %>%
  as.data.frame()


# I will also perform this operation for the global model

# global
global_background <- read_csv(
  file.path(here::here(), "vignette-outputs", "data-tables", "global_background_points_with_data_v3.csv")
  ) %>%
  as.data.frame()

```

I need to import the rasters of the area from which the background points were chosen so these can be displayed on each ExDet and MIC plot

```{r import background area rasters, eval = TRUE}

# path to directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")


# native
regional_native_background_area <- terra::rast(x = file.path(mypath, "v1_maxent_10km", "bio11_1981-2010_regional_native_KG.asc"))
# invaded
regional_invaded_background_area <- terra::rast(x = file.path(mypath, "v1_maxent_10km", "bio11_1981-2010_regional_invaded_KG.asc"))
# invaded_asian
regional_invaded_asian_background_area <- terra::rast(x = file.path(mypath, "v1_maxent_10km", "bio11_1981-2010_regional_invaded_asian_KG.asc"))


# convert to df
regional_native_background_area_df <- terra::as.data.frame(regional_native_background_area, xy = TRUE)
regional_invaded_background_area_df <- terra::as.data.frame(regional_invaded_background_area, xy = TRUE)
regional_invaded_asian_background_area_df <- terra::as.data.frame(regional_invaded_asian_background_area, xy = TRUE)

```

# 1. Extract variables for climate data

First, I will load the background points dataset used to train the model and that will be used to train the ExDet analysis. I have extracted the climate dataset from 1981-2010 for these points, but not for our projected time period of 2041-2070 and our 3 ssp scenarios. I will extract the values for the background points from each model and SSP scenario.

```{r extract CMIP6 values for background points- SSP126}

# regional_native model
# extract values from 2060 projected raster
regional_native_background_126 <- regional_native_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_126_env_covariates, y = ., method = "simple", ID = FALSE)




# regional_invaded model
regional_invaded_background_126 <- regional_invaded_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_126_env_covariates, y = ., method = "simple", ID = FALSE)





# regional_invaded_asian model
regional_invaded_asian_background_126 <- regional_invaded_asian_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_126_env_covariates, y = ., method = "simple", ID = FALSE)



# global
global_background_126 <- global_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_126_env_covariates, y = ., method = "simple", ID = FALSE)


```

```{r extract CMIP6 values for background points- SSP370}

# regional_native model
# extract values from 2060 projected raster
regional_native_background_370 <- regional_native_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_370_env_covariates, y = ., method = "simple", ID = FALSE)




# regional_invaded model
regional_invaded_background_370 <- regional_invaded_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_370_env_covariates, y = ., method = "simple", ID = FALSE)





# regional_invaded_asian model
regional_invaded_asian_background_370 <- regional_invaded_asian_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_370_env_covariates, y = ., method = "simple", ID = FALSE)



# global
global_background_370 <- global_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_370_env_covariates, y = ., method = "simple", ID = FALSE)


```

```{r extract CMIP6 values for background points- SSP585}

# regional_native model
# extract values from 2060 projected raster
regional_native_background_585 <- regional_native_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_585_env_covariates, y = ., method = "simple", ID = FALSE)




# regional_invaded model
regional_invaded_background_585 <- regional_invaded_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_585_env_covariates, y = ., method = "simple", ID = FALSE)





# regional_invaded_asian model
regional_invaded_asian_background_585 <- regional_invaded_asian_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_585_env_covariates, y = ., method = "simple", ID = FALSE)



# global
global_background_585 <- global_background %>%
  dplyr::select("X", "Y") %>%
  # extract new values
  terra::extract(x_global_585_env_covariates, y = ., method = "simple", ID = FALSE)


```

The `compute_extrapolation()` function only takes data columns (no descriptors), so I will isolate those from each background point set that was loaded in.

```{r select only data columns}

# native
regional_native_background <- dplyr::select(regional_native_background, 4:7)
# invaded
regional_invaded_background <- dplyr::select(regional_invaded_background, 4:7)
# regional_invaded_asian
regional_invaded_asian_background <- dplyr::select(regional_invaded_asian_background, 4:7)
# global
global_background <- dplyr::select(global_background, 4:7)

```

Now I will convert the global-scale climate rasters to df. These will be used for projection, to compare the exdet and MIC analyses with. The background point sets are used to train the analyses, and the global rasters are used for projection.

```{r convert projection rasters to df}

# historical
global_projection_hist_df <- terra::as.data.frame(x_global_hist_env_covariates, xy = TRUE)

# CMIP6
global_projection_126_df <- terra::as.data.frame(x_global_126_env_covariates, xy = TRUE)
global_projection_370_df <- terra::as.data.frame(x_global_370_env_covariates, xy = TRUE)
global_projection_585_df <- terra::as.data.frame(x_global_585_env_covariates, xy = TRUE)

```

# 2. Run ExDet / MIC Function

First, I will run the function that calculates the ExDet and MIC indices. I will use the background points datasets to train each analysis and will project to the entire global raster. This will be done for each of the historical and climate change time periods. This function outputs rasters and data frames that can be used for plotting.

To save length in this vignette, I will only display the sections creating the ExDet and MIC plots based on the historical data, as these sections are essentially copy-pasted.

```{r set wd, eval = TRUE}

# output directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent")

```

## 2.1 Historical Climate

```{r regional_native model}

exdet_MIC_regional_native_hist <- dsmextra::compute_extrapolation(
  samples = regional_native_background, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_hist_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded model}

exdet_MIC_regional_invaded_hist <- dsmextra::compute_extrapolation(
  samples = regional_invaded_background, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_hist_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded_asian model}

exdet_MIC_regional_invaded_asian_hist <- dsmextra::compute_extrapolation(
  samples = regional_invaded_asian_background, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_hist_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

also do global model

```{r global model}

exdet_MIC_global_hist <- dsmextra::compute_extrapolation(
  samples = global_background, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_hist_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

## 2.2 CMIP6 2041-2070 projected climate scenarios {.hidden}

### SSP126

```{r regional_native model- SSP126}

exdet_MIC_regional_native_126 <- dsmextra::compute_extrapolation(
  samples = regional_native_background_126, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_126_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded model- SSP126}

exdet_MIC_regional_invaded_126 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_background_126, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_126_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded_asian model- SSP126}

exdet_MIC_regional_invaded_asian_126 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_asian_background_126, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_126_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

also do global model

```{r global model- SSP126}

exdet_MIC_global_126 <- dsmextra::compute_extrapolation(
  samples = global_background_126, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_126_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

### SSP370

```{r regional_native model- SSP370}

exdet_MIC_regional_native_370 <- dsmextra::compute_extrapolation(
  samples = regional_native_background_370, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_370_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded model- SSP370}

exdet_MIC_regional_invaded_370 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_background_370, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_370_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded_asian model- SSP370}

exdet_MIC_regional_invaded_asian_370 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_asian_background_370, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_370_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

also do global model

```{r global model- SSP370}

exdet_MIC_global_370 <- dsmextra::compute_extrapolation(
  samples = global_background_370, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_370_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

### SSP585

```{r regional_native model- SSP585}

exdet_MIC_regional_native_585 <- dsmextra::compute_extrapolation(
  samples = regional_native_background_585, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_585_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded model- SSP585}

exdet_MIC_regional_invaded_585 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_background_585, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_585_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

```{r regional_invaded_asian model- SSP585}

exdet_MIC_regional_invaded_asian_585 <- dsmextra::compute_extrapolation(
  samples = regional_invaded_asian_background_585, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_585_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

also do global model

```{r global model- SSP585}

exdet_MIC_global_585 <- dsmextra::compute_extrapolation(
  samples = global_background_585, # training set of background points
  covariate.names = env_layer_names, # names of covariates
  prediction.grid = global_projection_585_df, # projection area for extrapolation test
  coordinate.system = "EPSG:4326", # CRS
  verbose = TRUE
)

```

# 3. regional_native- Save/plot ExDet/MIC rasters

Now I need to pull the rasters from the above functions and save them to my working directory folder. Then, I will create plots of each raster and save those as well.

## 3.1 regional_native hist

First, I will save the summary files from the analysis to each model folder.

### Save outputs

```{r save summary files- Rn hist}

# summary of data
write_csv(
  x = exdet_MIC_regional_native_hist[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_mic_analysis_results.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_hist[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_analysis_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_hist[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_univariate_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_hist[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_combinatorial_summary.csv")
  )

```

Next, I will save the output rasters.

```{r save rasters- Rn hist}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_native_hist[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "ExDet_slf_regional_native_v3_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_native_hist[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "MIC_slf_regional_native_v3_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

```

### Plot ExDet

Finally, I will create plots of each of the ExDet and MIC analyses.

```{r create exdet plot- Rn hist exdet, eval = TRUE}

# re-import raster
exdet_regional_native_hist <- terra::rast(x = file.path(mypath, "models", "slf_regional_native_v3", "ExDet_slf_regional_native_v3_background_projGlobe_1981-2010.tif"))
# convert to df
exdet_regional_native_hist_df <- terra::as.data.frame(exdet_regional_native_hist, xy = TRUE)
# mapping helper
exdet_regional_native_hist_plot <- map_exdet(
  raster = exdet_regional_native_hist, 
  raster.df = exdet_regional_native_hist_df
  ) 



# add new scale for bg area
exdet_regional_native_hist_plot <- exdet_regional_native_hist_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  # labs
  labs(
    title = "ExDet analysis | bio2, bio11, bio12 and bio15 | 1981-2010",
    subtitle = "Global climate divergence from 'Rn' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Rn hist exdet}


ggsave(
  exdet_regional_native_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_native_v3_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Rn hist mic, eval = TRUE}

# re-import
mic_regional_native_hist_df <- terra::rast(x = file.path(mypath, "models", "slf_regional_native_v3", "MIC_slf_regional_native_v3_background_projGlobe_1981-2010.tif")) %>%
  terra::as.data.frame(., xy = TRUE)

# plot mean raster first
mic_regional_native_hist_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = mic_regional_native_hist_df, aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  # aes
  labs(
    title = "Most Influential Covariate (MIC) analysis | 1981-2010",
    subtitle = "Covariate causing most extrapolation from 'Rn (native)' model training set"
    ) 

```

```{r save plot- Rn hist mic}


ggsave(
  mic_regional_native_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_native_v3_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

I will repeat this workflow for each of the models and their CMIP6 version climate data, so I will not display the workflow for the other models and ssp scenarios.

In the next vignette, I will utilize the ExDet analysis in my ensemble techniques for the regional-scale models. This ensembling technique should give me a unified prediction with which to compare the global-scale model and make predictions.

## 3.2 regional_native SSP126 {.hidden}

### Save outputs

```{r save summary files- Rn ssp126}

# summary of data
write_csv(
  x = exdet_MIC_regional_native_126[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_mic_analysis_results_2041-2070_ssp126_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_126[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_analysis_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_126[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_univariate_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_126[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp126_GFDL.csv")
  )

```

```{r save rasters- Rn ssp126}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_native_126[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_native_126[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_native_126 <- exdet_MIC_regional_native_126[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Rn ssp126 exdet}

# plot mean raster first
exdet_regional_native_126_plot <- map_exdet(
  raster = exdet_regional_native_126,
  raster.df = exdet_MIC_regional_native_126[["data"]][["all"]]
)

# add new scale for bg area
exdet_regional_native_126_plot <- exdet_regional_native_126_plot +
   # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Rn (native)' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Rn ssp126 exdet}


ggsave(
  exdet_regional_native_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Rn ssp126 mic}

# plot mean raster first
mic_regional_native_126_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_native_126[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  # aes
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Rn (native)' model training set"
    ) 

```

```{r save plot- Rn ssp126 mic}


ggsave(
  mic_regional_native_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 3.3 regional_native SSP370 {.hidden}

### Save outputs

```{r save summary files- Rn ssp370}

# summary of data
write_csv(
  x = exdet_MIC_regional_native_370[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_mic_analysis_results_2041-2070_ssp370_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_370[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_analysis_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_370[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_univariate_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_370[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp370_GFDL.csv")
  )

```

```{r save rasters- Rn ssp370}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_native_370[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_native_370[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_native_370 <- exdet_MIC_regional_native_370[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Rn ssp370 exdet}

# plot mean raster first
exdet_regional_native_370_plot <- map_exdet(
  raster = exdet_regional_native_370,
  raster.df = exdet_MIC_regional_native_370[["data"]][["all"]]
)
 
# new fill scale for bg area
exdet_regional_native_370_plot <- exdet_regional_native_370_plot + 
   # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Rn (native)' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Rn ssp370 exdet}


ggsave(
  exdet_regional_native_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Rn ssp370 mic}

# plot mean raster first
mic_regional_native_370_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_native_370[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  # aes
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Rn (native)' model training set"
    ) 

```

```{r save plot- Rn ssp370 mic}


ggsave(
  mic_regional_native_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 3.4 regional_native SSP585 {.hidden}

### Save outputs

```{r save summary files- Rn ssp585}

# summary of data
write_csv(
  x = exdet_MIC_regional_native_585[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_mic_analysis_results_2041-2070_ssp585_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_585[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "exdet_analysis_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_585[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_univariate_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_native_585[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_native_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp585_GFDL.csv")
  )

```

```{r save rasters- Rn ssp585}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_native_585[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_native_585[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_native_v3", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_native_585 <- exdet_MIC_regional_native_585[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Rn ssp585 exdet}

# plot mean raster first
exdet_regional_native_585_plot <- map_exdet(
  raster = exdet_regional_native_585,
  raster.df = exdet_MIC_regional_native_585[["data"]][["all"]]
)

# background area df scale
exdet_regional_native_585_plot <- exdet_regional_native_585_plot +
   # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Rn' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Rn ssp585 exdet}


ggsave(
  exdet_regional_native_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_native_v3_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot - Rn ssp585 mic}

# plot mean raster first
mic_regional_native_585_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_native_585[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_native_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  # aes
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Rn (native)' model training set"
    ) 

```

```{r save plot- Rn ssp585 mic}


ggsave(
  mic_regional_native_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_native_v3_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

# 4.regional_invaded- Save/plot ExDet/MIC rasters {.hidden}

## 4.1 regional_invaded hist

### Save outputs

```{r save summary files- Ri.NAmerica hist}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_hist[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_mic_analysis_results.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_hist[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_analysis_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_hist[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_univariate_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_hist[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_combinatorial_summary.csv")
  )

```

```{r save rasters- Ri.NAmerica hist}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_hist[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "ExDet_slf_regional_invaded_v7_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_hist[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "MIC_slf_regional_invaded_v7_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_hist <- exdet_MIC_regional_invaded_hist[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.NAmerica hist exdet}

# plot mean raster first
exdet_regional_invaded_hist_plot <- map_exdet(
  raster = exdet_regional_invaded_hist,
  raster.df = exdet_MIC_regional_invaded_hist[["data"]][["all"]]
)


exdet_regional_invaded_hist_plot <- exdet_regional_invaded_hist_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12 and bio15 | 1981-2010",
    subtitle = "Global climate divergence from 'Ri.NAmerica' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Ri.NAmerica hist exdet}


ggsave(
  exdet_regional_invaded_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_v7_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.NAmerica hist mic}

# plot mean raster first
mic_regional_invaded_hist_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_hist[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 1981-2010",
    subtitle = "Covariate causing most extrapolation from 'Ri.NAmerica' model model training set"
    ) 

```

```{r save plot- Ri.NAmerica hist mic}


ggsave(
  mic_regional_invaded_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_v7_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 4.2 regional_invaded SSP126

### Save outputs

```{r save summary files- Ri.NAmerica ssp126}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_126[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_mic_analysis_results_2041-2070_ssp126_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_126[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_analysis_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_126[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_univariate_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_126[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_combinatorial_summary_2041-2070_ssp126_GFDL.csv")
  )

```

```{r save rasters- Ri.NAmerica ssp126}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_126[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_126[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_126 <- exdet_MIC_regional_invaded_126[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.NAmerica ssp126 exdet}

# plot mean raster first
exdet_MIC_regional_invaded_126_plot <- map_exdet(
  raster = exdet_regional_invaded_126,
  raster.df = exdet_MIC_regional_invaded_126[["data"]][["all"]]
)

exdet_MIC_regional_invaded_126_plot <- exdet_MIC_regional_invaded_126_plot +
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp126, GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.NAmerica' model training set",
    fill = "ExDet Analysis"
    ) 


```

```{r save plot- Ri.NAmerica ssp126 exdet}


ggsave(
  exdet_MIC_regional_invaded_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.NAmerica ssp126 mic}

# plot mean raster first
mic_regional_invaded_126_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_126[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.NAmerica' model training set"
    ) 

```

```{r save plot- Ri.NAmerica ssp126 mic}


ggsave(
  mic_regional_invaded_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 4.3 regional_invaded SSP370

### Save outputs

```{r save summary files- Ri.NAmerica ssp370}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_370[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_mic_analysis_results_2041-2070_ssp370_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_370[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_analysis_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_370[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_univariate_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_370[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_combinatorial_summary_2041-2070_ssp370_GFDL.csv")
  )

```

```{r save rasters- Ri.NAmerica ssp370}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_370[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_370[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_370 <- exdet_MIC_regional_invaded_370[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot exdet}

# plot mean raster first
exdet_MIC_regional_invaded_370_plot <- map_exdet(
  raster = exdet_regional_invaded_370,
  raster.df = exdet_MIC_regional_invaded_370[["data"]][["all"]]
)

exdet_MIC_regional_invaded_370_plot <- exdet_MIC_regional_invaded_370_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.NAmerica' model training set",
    fill = "ExDet Analysis"
    ) 


```

```{r save plot- Ri.NAmerica ssp370 exdet}


ggsave(
  exdet_MIC_regional_invaded_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.NAmerica ssp370 mic}

# plot mean raster first
mic_regional_invaded_370_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_370[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.NAmerica' model training set"
    ) 

```

```{r save plot- Ri.NAmerica ssp370 mic}


ggsave(
  mic_regional_invaded_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 4.4 regional_invaded SSP585

### Save outputs

```{r save summary files- Ri.NAmerica ssp585}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_585[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_mic_analysis_results_2041-2070_ssp585_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_585[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_exdet_analysis_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_585[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_univariate_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_585[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_v7", "regional_invaded_mic_analysis_combinatorial_summary_2041-2070_ssp585_GFDL.csv")
  )

```

```{r save rasters- Ri.NAmerica ssp585}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_585[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_585[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_v7", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_585 <- exdet_MIC_regional_invaded_585[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.NAmerica ssp585 exdet}

# plot mean raster first
exdet_MIC_regional_invaded_585_plot <- map_exdet(
  raster = exdet_regional_invaded_585,
  raster.df = exdet_MIC_regional_invaded_585[["data"]][["all"]]
)

exdet_MIC_regional_invaded_585_plot <- exdet_MIC_regional_invaded_585_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.NAmerica' model training set",
    fill = "ExDet Analysis"
    ) 


```

```{r save plot- Ri.NAmerica ssp585 exdet}


ggsave(
  exdet_MIC_regional_invaded_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.NAmerica ssp585}

# plot mean raster first
mic_regional_invaded_585_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_585[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.NAmerica' model training set"
    ) 

```

```{r save plot- Ri.NAmerica ssp585 mic}


ggsave(
  mic_regional_invaded_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_v7_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

# 5. regional_invaded_asian- Save/plot ExDet/MIC rasters {.hidden}

## 5.1 regional_invaded_asian hist

### Save outputs

```{r save summary files- Ri.Asia hist}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_asian_hist[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_mic_analysis_results.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_hist[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_analysis_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_hist[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_univariate_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_hist[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_combinatorial_summary.csv")
  )

```

```{r save rasters- Ri.Asia hist}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_hist[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_hist[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_asian_hist <- exdet_MIC_regional_invaded_asian_hist[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.Asia hist}

# plot mean raster first
exdet_regional_invaded_asian_hist_plot <- map_exdet(
  raster = exdet_regional_invaded_asian_hist,
  raster.df = exdet_MIC_regional_invaded_asian_hist[["data"]][["all"]]
)

exdet_regional_invaded_asian_hist_plot <- exdet_regional_invaded_asian_hist_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12 and bio15 | 1981-2010",
    subtitle = "Global climate divergence from 'Ri.Asia' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Ri.Asia hist exdet}


ggsave(
  exdet_regional_invaded_asian_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.Asia hist}

# plot mean raster first
mic_regional_invaded_asian_hist_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_asian_hist[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 1981-2010",
    subtitle = "Covariate causing most extrapolation from 'Ri.ASia' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- Ri.Asia hist mic}


ggsave(
  mic_regional_invaded_asian_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 5.2 regional_invaded_asian- SSP126

### Save outputs

```{r save summary files- Ri.Asia ssp126}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_asian_126[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_mic_analysis_results_2041-2070_ssp126_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_126[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_analysis_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_126[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_univariate_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_126[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_combinatorial_summary_2041-2070_ssp126_GFDL.csv")
  )

```

```{r save rasters- Ri.Asia ssp126}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_126[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_126[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_asian_126 <- exdet_MIC_regional_invaded_asian_126[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.Asia ssp126}

# plot mean raster first
exdet_MIC_regional_invaded_asian_126_plot <- map_exdet(
  raster = exdet_regional_invaded_asian_126,
  raster.df =  exdet_MIC_regional_invaded_asian_126[["data"]][["all"]]
)

exdet_MIC_regional_invaded_asian_126_plot <- exdet_MIC_regional_invaded_asian_126_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.Asia' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Ri.Asia ssp126 exdet}


ggsave(
  exdet_MIC_regional_invaded_asian_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.Asia ssp126}

# plot mean raster first
mic_regional_invaded_asian_126_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_asian_126[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.Asia' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- Ri.Asia ssp126 mic}


ggsave(
  mic_regional_invaded_asian_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 5.3 regional_invaded_asian- SSP370

### Save outputs

```{r save summary files- Ri.Asia ssp370}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_asian_370[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_mic_analysis_results_2041-2070_ssp370_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_370[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_analysis_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_370[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_univariate_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_370[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_combinatorial_summary_2041-2070_ssp370_GFDL.csv")
  )

```

```{r save rasters- Ri.Asia ssp370}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_370[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_370[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_asian_370 <- exdet_MIC_regional_invaded_asian_370[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.Asia ssp370}

# plot mean raster first
exdet_MIC_regional_invaded_asian_370_plot <- map_exdet(
  raster = exdet_regional_invaded_asian_370,
  raster.df = exdet_MIC_regional_invaded_asian_370[["data"]][["all"]]
)

exdet_MIC_regional_invaded_asian_370_plot <- exdet_MIC_regional_invaded_asian_370_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.Asia' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Ri.Asia ssp370 exdet}


ggsave(
  exdet_MIC_regional_invaded_asian_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.Asia ssp370}

# plot mean raster first
mic_regional_invaded_asian_370_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_asian_370[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.Asia'model training set",
    fill = "MIC"
    ) 

```

```{r save plot- Ri.Asia ssp370 mic}


ggsave(
  mic_regional_invaded_asian_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 5.4 regional_invaded_asian- SSP585

### Save outputs

```{r save summary files- Ri.Asia ssp585}

# summary of data
write_csv(
  x = exdet_MIC_regional_invaded_asian_585[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_mic_analysis_results_2041-2070_ssp585_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_585[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_exdet_analysis_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_585[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_univariate_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_regional_invaded_asian_585[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "regional_invaded_asian_mic_analysis_combinatorial_summary_2041-2070_ssp585_GFDL.csv")
  )

```

```{r save rasters- Ri.Asia ssp585}

# exdet
raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_585[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_regional_invaded_asian_585[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_regional_invaded_asian_v2", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_regional_invaded_asian_585 <- exdet_MIC_regional_invaded_asian_585[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- Ri.Asia ssp585}

# plot mean raster first
exdet_MIC_regional_invaded_asian_585_plot <- map_exdet(
  raster = exdet_regional_invaded_asian_585,
  raster.df = exdet_MIC_regional_invaded_asian_585[["data"]][["all"]]
)

exdet_MIC_regional_invaded_asian_585_plot <- exdet_MIC_regional_invaded_asian_585_plot +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  theme(legend.key = element_rect(color = "black")) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Global climate divergence from 'Ri.Asia' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- Ri.Asia ssp585 exdet}


ggsave(
  exdet_MIC_regional_invaded_asian_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- Ri.Asia ssp585}

# plot mean raster first
mic_regional_invaded_asian_585_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_regional_invaded_asian_585[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  # new scale
  ggnewscale::new_scale_fill() +
  # add layer for bg point area
  geom_raster(data = regional_invaded_asian_background_area_df, aes(x = x, y = y, fill = "Background")) +
  # scale 2
  scale_fill_manual(name = "", values = c("Background" = "azure4"), guide = guide_legend(order = 2)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'Ri.Asia' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- Ri.Asia ssp585 mic}


ggsave(
  mic_regional_invaded_asian_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_regional_invaded_asian_v2_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

# 6. global- Save/plot ExDet/MIC rasters {.hidden}

## 6.1. global hist

First, I will save the summary files from the analysis to each model folder.

### Save outputs

```{r save summary files- global hist}

# summary of data
write_csv(
  x = exdet_MIC_global_hist[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_global_v3", "exdet_mic_analysis_results.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_global_hist[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_global_v3", "exdet_analysis_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_hist[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_univariate_summary.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_hist[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_combinatorial_summary.csv")
  )

```

Next, I will save the output rasters.

```{r save rasters- global hist}

# exdet
raster::writeRaster(
  x = exdet_MIC_global_hist[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "ExDet_slf_global_v3_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_global_hist[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "MIC_slf_global_v3_background_projGlobe_1981-2010.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_global_hist <- exdet_MIC_global_hist[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

Finally, I will create plots of each of the ExDet and MIC analyses.

```{r create exdet plot- global hist}

# plot mean raster first
exdet_global_hist_plot <- map_exdet(
  raster = exdet_global_hist,
  raster.df = exdet_MIC_global_hist[["data"]][["all"]]
) +
  # labels
  labs(
    title = "ExDet analysis | bio2, bio11, bio12 and bio15 | 1981-2010",
    subtitle = "Global climate divergence from 'global' model training set",
    fill = "ExDet Analysis"
    ) 

```

```{r save plot- global hist exdet}


ggsave(
  exdet_global_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_global_v3_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- global hist}

# plot mean raster first
mic_global_hist_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_global_hist[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 1981-2010",
    subtitle = "Covariate causing most extrapolation from 'global' model training set",
    fill = "MIC"
    ) 
```

```{r save plot- global hist mic}


ggsave(
  mic_global_hist_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_global_v3_background_projGlobe_1981-2010.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 6.2 global SSP126

### Save outputs

```{r save summary files- global ssp126}

# summary of data
write_csv(
  x = exdet_MIC_global_126[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_global_v3", "exdet_mic_analysis_results_2041-2070_ssp126_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_global_126[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_global_v3", "exdet_analysis_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_126[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_univariate_summary_2041-2070_ssp126_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_126[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp126_GFDL.csv")
  )

```

```{r save rasters- global ssp126}

# exdet
raster::writeRaster(
  x = exdet_MIC_global_126[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_global_126[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp126_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_global_126 <- exdet_MIC_global_126[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot}

# plot mean raster first
exdet_global_126_plot <- map_exdet(
  raster = exdet_global_126,
  raster.df = exdet_MIC_global_126[["data"]][["all"]]
) +
  # labels
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Global climate divergence from 'global' model training set",
    fill = "ExDet Analysis"
    ) 
  
```

```{r save plot- global ssp126 exdet}


ggsave(
  exdet_global_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- global ssp126}

# plot mean raster first
mic_global_126_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_global_126[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp126 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'global' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- global ssp126 mic}


ggsave(
  mic_global_126_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp126_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 6.3 global SSP370

### Save outputs

```{r save summary files- global ssp370}

# summary of data
write_csv(
  x = exdet_MIC_global_370[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_global_v3", "exdet_mic_analysis_results_2041-2070_ssp370_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_global_370[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_global_v3", "exdet_analysis_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_370[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_univariate_summary_2041-2070_ssp370_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_370[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp370_GFDL.csv")
  )

```

```{r save rasters- global ssp370}

# exdet
raster::writeRaster(
  x = exdet_MIC_global_370[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_global_370[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp370_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_global_370 <- exdet_MIC_global_370[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- global ssp370}

# plot mean raster first
exdet_global_370_plot <- map_exdet(
  raster = exdet_global_370,
  raster.df = exdet_MIC_global_370[["data"]][["all"]]
)
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Global climate divergence from 'global' model training set",
    fill = "ExDet Analysis"
    ) 
  
```

```{r save plot- global ssp370 exdet}


ggsave(
  exdet_global_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- global ssp370}

# plot mean raster first
mic_global_370_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_global_370[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp370 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'global' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- global ssp370 mic}


ggsave(
  mic_global_370_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp370_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

## 6.4 global SSP585

### Save outputs

```{r save summary files- global ssp585}

# summary of data
write_csv(
  x = exdet_MIC_global_585[["data"]][["all"]],
  file = file.path(mypath, "models", "slf_global_v3", "exdet_mic_analysis_results_2041-2070_ssp585_GFDL.csv")
  )

# exdet
write_csv(
  x = as.data.frame(exdet_MIC_global_585[["summary"]][["extrapolation"]]),
  file = file.path(mypath, "models", "slf_global_v3", "exdet_analysis_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_585[["summary"]][["mic"]][[1]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_univariate_summary_2041-2070_ssp585_GFDL.csv")
  )

# mic univariate
write_csv(
  x = as.data.frame(exdet_MIC_global_585[["summary"]][["mic"]][[2]]),
  file = file.path(mypath, "models", "slf_global_v3", "mic_analysis_combinatorial_summary_2041-2070_ssp585_GFDL.csv")
  )

```

```{r save rasters- global ssp585}

# exdet
raster::writeRaster(
  x = exdet_MIC_global_585[["rasters"]][["ExDet"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

raster::writeRaster(
  x = exdet_MIC_global_585[["rasters"]][["mic"]][["all"]], 
  filename = file.path(mypath, "models", "slf_global_v3", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp585_GFDL.tif"),
  filetype = "GTiff",
  overwrite = FALSE
  )

# also isolate exdet raster
exdet_global_585 <- exdet_MIC_global_585[["rasters"]][["ExDet"]][["all"]]

```

### Plot ExDet

```{r create exdet plot- global ssp585}

# plot mean raster first
exdet_global_585_plot <- map_exdet(
  raster = exdet_global_585,
  raster.df = exdet_MIC_global_585[["data"]][["all"]]
) +
  labs(
    title = "ExDet analysis | bio2, bio11, bio12, bio15 | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Global climate divergence from 'global' model training set",
    fill = "ExDet Analysis"
    ) 
  
```

```{r save plot- global ssp585 exdet}


ggsave(
  exdet_global_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "ExDet_slf_global_v3_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

### Plot MIC

```{r create MIC plot- global ssp585}

# plot mean raster first
mic_global_585_plot <- ggplot() +
  # default plot style
  map_style_MIC +
  # plot regular raster of values first
  geom_raster(data = exdet_MIC_global_585[["data"]][["all"]], aes(x = x, y = y, fill = as.factor(mic))) +
  # fill scale 1
  viridis::scale_fill_viridis(name = "MIC", discrete = TRUE, labels = c("analogous", "bio11", "bio12", "bio15", "bio2"), guide = guide_legend(order = 1)) +
  labs(
    title = "Most Influential Covariate (MIC) analysis | 2041-2070 | ssp585 GFDL-ESM4",
    subtitle = "Covariate causing most extrapolation from 'global' model training set",
    fill = "MIC"
    ) 

```

```{r save plot- global ssp585 MIC}


ggsave(
  mic_global_585_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_global_v3_background_projGlobe_2041-2070_ssp585_GFDL.jpg"),
  height = 8, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

# References

1. Bouchet, P. J., Miller, D. L., Roberts, J. J., Mannocci, L., Harris, C. M., & Thomas, L. (2020). dsmextra: Extrapolation assessment tools for density surface models. Methods in Ecology and Evolution, 11(11), 14641469. https://doi.org/10.1111/2041-210X.13469

2. Elith, J., Kearney, M., & Phillips, S. (2010). The art of modelling range-shifting species. Methods in Ecology and Evolution, 1(4), 330342. https://doi.org/10.1111/j.2041-210X.2010.00036.x

3. Mesgaran, M. B., Cousens, R. D., & Webber, B. L. (2014). Here be dragons: A tool for quantifying novelty due to covariate range and correlation change when projecting species distribution models. Diversity and Distributions, 20(10), 11471159. https://doi.org/10.1111/ddi.12209


# Appendix {.hidden}

## Patchwork plots- MIC

```{r patchwork regional_native}

mic_regional_native_hist_plot <- mic_regional_native_hist_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
    )

# CMIP6
mic_regional_native_126_plot <- mic_regional_native_126_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )

mic_regional_native_370_plot <- mic_regional_native_370_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "C") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )

mic_regional_native_585_plot <- mic_regional_native_585_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )




mic_regional_native_patchwork <- (
  mic_regional_native_hist_plot / mic_regional_native_126_plot /
  mic_regional_native_370_plot / mic_regional_native_585_plot
    ) +
  plot_layout(axes = "collect", axis_titles = "collect", guides = "collect") +
  plot_annotation(
    title = "Most Influential Covariate (MIC) for 'Rn (native)' model for (A) Present and 2055 climate change scenarios\n(B) SSP126 (C) SSP370 (D) SSP585",
    subtitle = "Covariate causing most extrapolation from model training set"
    ) &
  theme(legend.position = "bottom")

```

```{r save plot- Rn mic patchwork}

ggsave(
  mic_regional_native_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_native_v3_background_projGlobe_patchwork.jpg"),
  height = 14, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

```{r patchwork regional_invaded}

mic_regional_invaded_hist_plot <- mic_regional_invaded_hist_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
    )

# CMIP6
mic_regional_invaded_126_plot <- mic_regional_invaded_126_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


mic_regional_invaded_370_plot <- mic_regional_invaded_370_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "C") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


mic_regional_invaded_585_plot <- mic_regional_invaded_585_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0))  +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25)
  )


mic_regional_invaded_patchwork <- (
  mic_regional_invaded_hist_plot / mic_regional_invaded_126_plot /
  mic_regional_invaded_370_plot / mic_regional_invaded_585_plot
  ) +
  plot_layout(axes = "collect", axis_titles = "collect", guides = "collect") +
  plot_annotation(
    title = "Most Influential Covariate (MIC) for 'Ri.NAmerica' model for (A) Present and 2055 climate change scenarios\n(B) SSP126 (C) SSP370 (D) SSP585",
    subtitle = "Covariate causing most extrapolation from model training set"
    ) &
  theme(legend.position = "bottom")

```

```{r save plot- Ri.NAMerica mic patchwork}

ggsave(
  mic_regional_invaded_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_invaded_v7_background_projGlobe_patchwork.jpg"),
  height = 14, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```

```{r patchwork regional_invaded_asian}

mic_regional_invaded_asian_hist_plot <- mic_regional_invaded_asian_hist_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25),
    legend.position = "none"
    )

# CMIP6
mic_regional_invaded_asian_126_plot <- mic_regional_invaded_asian_126_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25),
    legend.position = "bottom"
  )


mic_regional_invaded_asian_370_plot <- mic_regional_invaded_asian_370_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "C") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25),
    legend.position = "bottom"
  )



mic_regional_invaded_asian_585_plot <- mic_regional_invaded_asian_585_plot + 
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  theme(
    plot.tag.location = "panel",
    plot.tag.position = c(0.05, 0.1),
    plot.tag = element_text(face = "bold", size = 25),
    legend.position = "bottom"
  )



mic_regional_invaded_asian_patchwork <- (
  mic_regional_invaded_asian_hist_plot / mic_regional_invaded_asian_126_plot /
  mic_regional_invaded_asian_370_plot / mic_regional_invaded_asian_585_plot
  ) +
  plot_layout(axes = "collect", axis_titles = "collect") +
  plot_annotation(
    title = "Most Influential Covariate (MIC) for 'Ri.Asia' model for (A) Present and 2055 climate change scenarios\n(B) SSP126 (C) SSP370 (D) SSP585",
    subtitle = "Covariate causing most extrapolation from model training set"
    ) 

```

```{r save plot- Ri.Asia mic patchwork}

ggsave(
  mic_regional_invaded_asian_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "MIC_slf_invaded_asian_v2_background_projGlobe_patchwork.jpg"),
  height = 14, 
  width = 12,
  device = jpeg,
  dpi = "retina"
  )

```
