---
title: "Plot suitability for SLF in important grape growing regions and climate change effects"
author: "Samuel M. Owens"
contact: "sam.owens@temple.edu"
date: "2024-05-06"
output: html_document
---

# Overview

EDIT

This vignette will compare the invaded region model for SLF, which was created in `050_run_easternUSA_MaxEnt_models.Rmd`, with the global version of this model, which was created in `070_run_global_historical_MaxEnt_model.Rmd`. First, I will compare these models using flexible AUC. Second, I will also project these models to N. America and compare the proportion of suitable area. Third, I will calculate suitability values for all SLF points and IVR regions. Finally, I will perform an invasion stage analysis following the methods of Gallien et.al, 2012.

In step 1, I tested the best fit background data using SLF occurrences and climate data only from the SLF invaded range (the eastern half of the United States).

(THEORY)

# Setup

```{r load necesssary packages, echo = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here::here() # here() starts at the root folder of this package.
library(devtools)

# spatial data handling
library(terra)
library(scrubr)

# plot aesthetics
library(scales)
library(patchwork)
library(grid)

```

```{r set wd}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

I will load in some aesthetic objects, including for breaks.

```{r axis breaks}

breaks <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)

```

```{r axis labels}

labels <- c(0, 2, 4, 6, 8, 10)

```

Load in summary files for the global and regional ensemble models, which contain the thresholds.

```{r load in summary files}
# summary file to extract thresholds from

# global
summary_global <- read_csv(file = file.path(mypath, "slf_global_v2", "global_summary_all_iterations.csv"))

summary_regional_ensemble <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_threshold_values.csv"))


```

I will load in dataset containing locations of globally important viticultural regions. I will also de-duplicate the records because there were "many-to-many" joining matches downstream. De-duplicating removed about 17 records. I also remove the species column, which is automatically created by `SDMtune::predict()` upstream.

```{r load in wineries dataset}

load(file.path(here::here(), "data", "wineries.rda"))

IVR_locations <- wineries %>%
  # tidy
  dplyr::select(-c(Latitude, Longitude)) %>%
  dplyr::filter(
    !is.na(x),
    !is.na(y)
    ) %>%
  # de-duplicate
  scrubr::dedup(how = "one", tolerance = 0.7) # change to 0.92 on laptop PC
  # I chose the value of the tolerance that retrieved the number of records closest to the number of records in the xy_global_1995 (further down, 1064 records)

# remove unnecessary obj
rm(wineries)

```

Finally, I will load in the global and regional ensemble suitability maps. These will be used both for extracting the new xy suitability values and for plotting.

```{r load in suitability rasters}

# global
global_1995 <- terra::rast(x = file.path(mypath, "slf_global_v2", "global_pred_suit_clamped_cloglog_globe_1981-2010_mean.asc"))

global_2055 <- terra::rast(x = file.path(mypath, "slf_global_v2", "global_pred_suit_clamped_cloglog_globe_2041-2070_GFDL_ssp370_mean.asc"))


# regional_ensemble
regional_ensemble_1995 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_regional_weighted_mean_globe_1981-2010.asc")
  )

regional_ensemble_2055 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_regional_weighted_mean_globe_2041-2070_GFDL_ssp370.asc")
  )

```

# 1. Retrieve/Import xy suitability

These scatter plots will be based on the suitability for the IVR points in both the global and regional_ensemble models. I have already calculated the xy suitability for the global model based on these points, using the function `slfSpread::predict_xy_suitability()`. This function will not work for the regional_ensemble because it calls for a model object, which we did not use to predict the ensemble suitability. So, I will use `terra::extract()` to perform this action.

I will load in the global model datasets and create the regional_ensemble datasets. I will also do some tidying of my datasets for the plots I will create.

```{r import global suitability}

xy_global_1995 <- read_csv(
  file = file.path(mypath, "slf_global_v2", "global_wineries_1981-2010_xy_pred_suit_clamped_cloglog_mean.csv")
  ) 

xy_global_2055 <- read_csv(
  file = file.path(mypath, "slf_global_v2", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_clamped_cloglog_mean.csv")
  ) 

```

The xy_suitabilities for the global models also contain a few less records than the IVR regions. To ensure that I am retrieving the suitability values for the exact same vineyards in the regional_ensemble, I will perform a filtering join. The IVR dataset should now contain the same records as the global suitability values do

```{r de-duplicate and filter join}

# de-duplicate just to make sure
xy_global_1995 <- scrubr::dedup(xy_global_1995, how = "one", tolerance = 0.99) # change to 0.93 on laptop PC
xy_global_2055 <- scrubr::dedup(xy_global_2055, how = "one", tolerance = 0.99) 


# edits to IVR dataset
# filter join of IVRs
IVR_locations <- semi_join(IVR_locations, xy_global_1995, by = c("x", "y")) 
# add ID column
IVR_locations <- IVR_locations %>%
  dplyr::mutate(ID = row_number()) %>%
  relocate(ID, x, y)

```

```{r save tidied data}

if (FALSE) {

  readr::write_rds(IVR_locations, file.path(here::here(), "data", "wineries_tidied.rds"))
  
  #read it back in
  IVR_locations <- read_rds(file.path(here::here(), "data", "wineries_tidied.rds"))
  
}

```


Now, I will retrieve the suitability values for the regional_ensemble. Instead of returning the coordinates from the map, I will join the coordinates from the original IVR_locations dataset so that the coordinates are exact for joining with other datasets.

```{r retrieve xy suitability for regional_ensemble}

# 1995
xy_regional_ensemble_1995 <- terra::extract(
  x = regional_ensemble_1995,
  y = dplyr::select(IVR_locations, x, y), # points
  method = "simple",
  xy = FALSE, # dont return coordinates
  ID = TRUE
)


# 2055
xy_regional_ensemble_2055 <- terra::extract(
  x = regional_ensemble_2055,
  y = dplyr::select(IVR_locations, x, y), # points
  method = "simple",
  xy = FALSE, # dont return coordinates 
  ID = TRUE
) 

```

```{r join xy coordinates}

# joining object
IVR_coordinates <- dplyr::select(IVR_locations, ID, x, y)

# perform join
xy_regional_ensemble_1995 <- left_join(xy_regional_ensemble_1995, IVR_coordinates, by = "ID") %>%
  dplyr::select(-ID)

xy_regional_ensemble_2055 <- left_join(xy_regional_ensemble_2055, IVR_coordinates, by = "ID") %>%
  dplyr::select(-ID)

```

Now, I will tidy and save the datasets

```{r tidy datasets}

# global model datasets
xy_global_1995 <- xy_global_1995 %>%
  dplyr::select(-Species) %>%
  # rename the column for future joining
  dplyr::rename("xy_global_1995" = "cloglog_suitability") %>%
  dplyr::filter(!is.na(xy_global_1995))

xy_global_2055 <- xy_global_2055 %>%
  dplyr::select(-Species) %>%
  dplyr::rename("xy_global_2055" = "cloglog_suitability") %>%
  dplyr::filter(!is.na(xy_global_2055))



# regional_ensemble datasets
xy_regional_ensemble_1995 <- xy_regional_ensemble_1995 %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_1995" = "sum") %>%
  dplyr::filter(!is.na(xy_regional_ensemble_1995)) %>%
  dplyr::relocate(x, y)

xy_regional_ensemble_2055 <- xy_regional_ensemble_2055 %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_2055" = "sum") %>%
  dplyr::filter(!is.na(xy_regional_ensemble_2055)) %>%
  dplyr::relocate(x, y)

# we lost 1 record because of the NA removal, but that is OK

```

```{r save predicted suitability datasets}

# 1995
write_csv(
  x = xy_regional_ensemble_1995, 
  file = file.path(mypath, "slf_regional_ensemble_v0", "regional_ensemble_wineries_1981-2010_xy_pred_suit.csv")
  )
# 2055
write_csv(
  x =  xy_regional_ensemble_2055,
  file.path(mypath, "slf_regional_ensemble_v0", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.csv")
  )

# also save to .rds
readr::write_rds(
  xy_regional_ensemble_1995,  
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit.rds")
  )

readr::write_rds(
  xy_regional_ensemble_2055,  
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.rds")
  )


# finally, save global datasets
readr::write_rds(
  xy_global_1995,  
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit.rds")
  )

readr::write_rds(
  xy_global_2055,  
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.rds")
  )

```



# 2. Transform xy suitability

I will plot the suitability values in two different ways- I will plot the raw xy suitability and I will transform the data so that the MTSS threshold is the center of the scatter plot. This way, movement across the minimum suitability threshold is more easily visualized. I will transform all 4 vectors of suitability values, 2 per model, in preparation for plotting.

I created the function `slfSpread::rescale_cloglog_suitability()` to accomplish this task. This function uses a vector of exponential transformations for the specified value of `thresh` to apply an exponential equation to the vector of suitability values. It then applies the equation `y = c1 * c2^x + c3` to the vector, where x is the input suitability values, y is the transformed version of those values, c1 and c3 are the maximum and its inverse, and c2 is the interpolated value of the input `thresh`. The transformed suitability vector is re-scaled so that `thresh` is the median (0.5) on a 0-1 scale and all other values are transformed to fit this scale.

```{r re-load pred suit datasets}

# global
xy_global_1995 <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit.rds"))
xy_global_2055 <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.rds"))

# regional
xy_regional_ensemble_1995 <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit.rds"))
xy_regional_ensemble_2055 <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.rds"))

```

I will use the internal function `slfSpread::rescale_cloglog_suitability()` to re-scale these suitability values. 

```{r transform global xy data}

xy_global_1995_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_global_1995,
  thresh = "MTSS",
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_global,
  rescale.name = "xy_global_1995",
  rescale.thresholds = TRUE
)
# separate data from thresholds
xy_global_1995_rescaled_thresholds <- xy_global_1995_rescaled[[2]]
xy_global_1995_rescaled <- xy_global_1995_rescaled[[1]]



xy_global_2055_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_global_2055,
  thresh = "MTSS",  # the global model only has 1 MTSS thresh
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_global,
  rescale.name = "xy_global_2055",
  rescale.thresholds = TRUE
)

xy_global_2055_rescaled_thresholds <- xy_global_2055_rescaled[[2]]
xy_global_2055_rescaled <- xy_global_2055_rescaled[[1]]

```

```{r transform regional_ensemble xy data}

xy_regional_ensemble_1995_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_1995,
  thresh = "MTSS",
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble,
  rescale.name = "xy_regional_ensemble_1995",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_1995_rescaled_thresholds <- xy_regional_ensemble_1995_rescaled[[2]]
xy_regional_ensemble_1995_rescaled <- xy_regional_ensemble_1995_rescaled[[1]]



xy_regional_ensemble_2055_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_2055,
  thresh = "MTSS.CC", # the way the thresholds are calculated for the regional_ensemble model means that the threshold will be slightly different for climate change
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble,
  rescale.name = "xy_regional_ensemble_2055",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_2055_rescaled_thresholds <- xy_regional_ensemble_2055_rescaled[[2]]
xy_regional_ensemble_2055_rescaled <- xy_regional_ensemble_2055_rescaled[[1]]

```

```{r save re-scaled values}

# global
write_rds(
  xy_global_1995_rescaled, 
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled.rds")
  )
write_rds(
  xy_global_2055_rescaled, 
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled.rds")
  )

# regional
write_rds(
  xy_regional_ensemble_1995_rescaled, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled.rds")
  )
write_rds(
  xy_regional_ensemble_2055_rescaled, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled.rds")
  )

```

```{r save thresholds}

# global
write_rds(
  xy_global_1995_rescaled_thresholds, 
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds")
  )
write_rds(
  xy_global_2055_rescaled_thresholds, 
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled_thresholds.rds")
  )

# regional
write_rds(
  xy_regional_ensemble_1995_rescaled_thresholds, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds")
  )
write_rds(
  xy_regional_ensemble_2055_rescaled_thresholds, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled_thresholds.rds")
  )

```


# 3. Plot untransformed suitability values

We need a baseline for visualizing the trends in these scatter plots, so I will first plot the un-transformed datasets.

```{r join datasets}

# join datasets for plotting
xy_joined <- full_join(xy_global_1995, xy_regional_ensemble_1995, by = c("x", "y")) %>%
  # join CC datasets
  full_join(., xy_global_2055, by = c("x", "y")) %>%
  full_join(., xy_regional_ensemble_2055, by = c("x", "y")) %>%
  # order
  dplyr::relocate(x, y, xy_global_1995, xy_global_2055)

```

```{r plot SLF suitability values, fig.asp = 1}

# figure annotation title
# "suitability for Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_plot <- ggplot(data = xy_joined) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = as.numeric(summary_regional_ensemble[3, 4]), linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # historical data
   geom_point(
     aes(x = xy_global_1995, y = xy_regional_ensemble_1995, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055, y = xy_regional_ensemble_2055, shape = "2041-2070\nGFDL ssp370"), 
     size = 2, stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name =  "'global' model cloglog suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name =  "'regional_ensemble' model cloglog suitability", limits = c(0, 1), breaks = breaks) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot}

ggsave(
  xy_joined_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "xy_suitability_IVR_regions_global_regional_ensemble.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

# 4. plot transformed suitability values

I will manually change the scale of these values to a 1-10 scale so that this plot of risk is not confused for a measure of suitability from the model.

```{r re-load rescaled values}

# global
xy_global_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled.rds"))

xy_global_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled.rds"))

# regional
xy_regional_ensemble_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled.rds"))

xy_regional_ensemble_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled.rds"))

```

```{r re-load thresholds}

# global
xy_global_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_global_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled_thresholds.rds"))

# regional
xy_regional_ensemble_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_regional_ensemble_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_rescaled_thresholds.rds"))

```


```{r join datasets}

# join datasets for plotting
xy_joined_rescaled <- full_join(xy_global_1995_rescaled, xy_regional_ensemble_1995_rescaled, by = c("x", "y")) %>%
  # join CC datasets
  full_join(., xy_global_2055_rescaled, by = c("x", "y")) %>%
  full_join(., xy_regional_ensemble_2055_rescaled, by = c("x", "y")) %>%
  # order
  dplyr::relocate(x, y, xy_global_1995_rescaled, xy_global_2055_rescaled) %>%
  dplyr::select(-c(xy_global_1995, xy_global_2055, xy_regional_ensemble_1995, xy_regional_ensemble_2055))


```

I will need to create a second dataset for the arrow segments indicating change. I will filter out only the segments that cross either threshold and then plot these arrows.

First, I need to isolate the MTSS threshold values.

```{r retrieve threshold values}

# global
global_MTSS <- as.numeric(xy_global_1995_rescaled_thresholds[2, 2])
# regional ensemble
regional_ensemble_MTSS_1995 <- as.numeric(xy_regional_ensemble_1995_rescaled_thresholds[2, 2])
regional_ensemble_MTSS_2055 <- as.numeric(xy_regional_ensemble_1995_rescaled_thresholds[4, 2])

```

Next, I will use case_when() (a vectorized ifelse) to calculate when a point shifts across a suitability threshold due to climate change.

```{r filter segments that cross thresholds}

xy_joined_rescaled_intersects <- xy_joined_rescaled %>%
  mutate(
    crosses_threshold = case_when(
      # conditional for starting and ending points that overlap a the threshold
      # x-axis
      xy_global_1995_rescaled > global_MTSS & xy_global_2055_rescaled < global_MTSS ~ "crosses",
      xy_global_1995_rescaled < global_MTSS & xy_global_2055_rescaled > global_MTSS ~ "crosses",
      # y-axis
      xy_regional_ensemble_1995_rescaled > regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled < regional_ensemble_MTSS_2055 ~ "crosses",
       xy_regional_ensemble_1995_rescaled < regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled > regional_ensemble_MTSS_2055 ~ "crosses",
      # else
      .default = "does not cross"
    )
  )

# filter out the crosses
xy_joined_rescaled_intersects <- dplyr::filter(
  xy_joined_rescaled_intersects,
  crosses_threshold == "crosses"
)

```

Now lets plot the data.

```{r plot SLF suitability values, fig.asp = 1}

# figure annotation title
# "Risk of Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_rescaled_plot <- ggplot(data = xy_joined_rescaled) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = global_MTSS, linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = regional_ensemble_MTSS_1995, linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # arrows indicating change
   geom_segment(
     data = xy_joined_rescaled_intersects,
     aes(
       x = xy_global_1995_rescaled,
       xend = xy_global_2055_rescaled,
       y = xy_regional_ensemble_1995_rescaled,
       yend = xy_regional_ensemble_2055_rescaled
     ), 
     arrow = grid::arrow(angle = 5.5, type = "closed"), alpha = 0.3, linewidth = 0.25, color = "black"
   ) +
   # historical data
   geom_point(
     aes(x = xy_global_1995_rescaled, y = xy_regional_ensemble_1995_rescaled, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055_rescaled, y = xy_regional_ensemble_2055_rescaled, shape = "2041-2070\nGFDL ssp370"), 
     size = 2, stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot}

ggsave(
  xy_joined_rescaled_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "SLF_risk_IVR_regions_global_regional_ensemble.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

# I will also save the ggplot object as an rds because I need to facet it later
write_rds(
  xy_joined_rescaled_plot,
  file = file.path(here::here(), "vignette-outputs", "figures", "figures-rds", "SLF_risk_IVR_regions_global_regional_ensemble.rds")
  )

```

# 5. Create summary table of transformed plot

I will now create a summary table to explain the rescaled plots from step 4. The table will depict the quadrant placement of the point in the quadrant plot, both before and after climate change. From this, I will calculate the total number of movements into and out of each quadrant. I will apply the internal function `slfSpread::calculate_risk_quadrant()`. 

I will create a summary table of the quadrant placement (and thus the level of risk) for each point in the IVR_locations dataset. I will use `calculate_risk_quadrant()` to accomplish this.

```{r summary IVRs}

# create dataset and tidy
IVR_locations_joined <- left_join(IVR_locations, xy_joined_rescaled, by = c("x", "y")) %>%
  relocate(ID, x, y) %>%
  # remove NA
  slice(-1064)
  
# calculate risk quadrants
IVR_locations_risk <- IVR_locations_joined %>%
  mutate(
    risk_1995 = slfSpread::calculate_risk_quadrant(
      suit.x = IVR_locations_joined$xy_global_1995_rescaled,
      suit.y = IVR_locations_joined$xy_regional_ensemble_1995_rescaled,
      thresh.x = global_MTSS, # this threshold remains the same
      thresh.y = regional_ensemble_MTSS_1995
    ),
    risk_2055 = slfSpread::calculate_risk_quadrant(
      suit.x = IVR_locations_joined$xy_global_2055_rescaled,
      suit.y = IVR_locations_joined$xy_regional_ensemble_2055_rescaled,
      thresh.x = global_MTSS,
      thresh.y = regional_ensemble_MTSS_2055
    ),
    risk_shift = str_c(risk_1995, risk_2055, sep = "-")
  )

```

```{r factor ordering of risk}

risk_levels <- c("extreme", "high", "moderate", "low")

```

```{r risk categories table}

IVR_risk_table <- IVR_locations_risk %>%
  # create counts and make into acrostic table
  dplyr::group_by(risk_1995, risk_2055) %>%
  dplyr::summarize(count = n()) %>%
  pivot_wider(names_from = risk_2055, values_from = count) %>%
  # tidy
  dplyr::rename("rows_1995_cols_2055" = "risk_1995") %>%
  relocate("rows_1995_cols_2055", "extreme", "high", "moderate") %>%
  arrange(factor(.$rows_1995_cols_2055, levels = risk_levels)) %>%
  # replace missing categories with 0
  replace(is.na(.), 0) %>%
  ungroup()


# tidy
IVR_risk_table <- IVR_risk_table %>%
  # add totals column
  tibble::add_column("total_present" = rowSums(.[, 2:5])) %>%
  # add row totals
  tibble::add_row(rows_1995_cols_2055 = "total_2055", extreme = colSums(dplyr::select(., 2)), high = colSums(dplyr::select(., 3)), moderate = colSums(dplyr::select(., 4)), low = colSums(dplyr::select(., 5)), total_present = 1063) %>%
  # convert to df
  as.data.frame()

```

```{r export table}

write_csv(IVR_risk_table, file = file.path(here::here(), "vignette-outputs", "data-tables", "IVR_risk_table_global_regional_ensemble.csv"))

```

# 6. global risk shift vs regional risk shift

I will create a table to sum the number of points in three different groups.  My goal is to understand how the regional model adds resolution to our calculation of risk. So, I will look at how

I will sum the number of points that are suitable in the global model only, unsuitable in the global model only, and unsuitable in the global model / suitable in the regional model. I will repeat this operation for both time periods.

```{r create table of sums}

global_regional_risk_shift <- tibble(
  time_period = c(1995, 1995, 1995, 2055, 2055, 2055),
  quadrants = c("quad4_quad2", "quad3_quad1", "quad3", "quad4_quad2", "quad3_quad1", "quad3"),
  risk = c("extreme_moderate", "high_low", "high", "extreme_moderate", "high_low", "high"),
  model_suit = c("global_suit", "global_unsuit", "global_unsuit_regional_suit", "global_suit", "global_unsuit", "global_unsuit_regional_suit"),
  IVR_region_count = c(
    # global suitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled >= global_MTSS),
    # global unsuitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled < global_MTSS),
    # global unsuitable and regional suitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled < global_MTSS & IVR_locations_joined$xy_regional_ensemble_1995_rescaled >= regional_ensemble_MTSS_1995),
    # global suitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled >= global_MTSS),
    # global unsuitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled < global_MTSS),
    # global unsuitable and regional suitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled < global_MTSS & IVR_locations_joined$xy_regional_ensemble_2055_rescaled >= regional_ensemble_MTSS_2055)
  )
)

```

```{r add proportions and make specific calculations}

# total # IVRs
total_IVR <- sum(global_regional_risk_shift[1:2, 5])

global_regional_risk_shift <- mutate(
  global_regional_risk_shift, 
  IVR_region_prop = IVR_region_count / total_IVR
  )


# calculate % of unsuit (quad3 and quad 1) that are are in quad3
quad3_risk_prop <- tibble(
  time_period = c("quad3_1995", "quad3_2055"),
  prop_total_unsuit_in_quad3 = c(
    scales::label_percent(accuracy = 0.01) (abs(as.numeric((global_regional_risk_shift[3, 5]) / global_regional_risk_shift[2, 5]))),
    scales::label_percent(accuracy = 0.01) (abs(as.numeric((global_regional_risk_shift[6, 5]) / global_regional_risk_shift[5, 5])))
  )
)



# number and proportion of IVRs shifting with CC
#global_regional_risk_shift_summary <- tibble(
##  "risk_shift" = c("global_unsuit_shift_1995_2055", "quad3_high_shift_diff_1995_2055"),
#  "IVR_region_change_count" = c(
#    as.numeric(global_regional_risk_shift[2, 5]) - as.numeric(global_regional_risk_shift[5, 5]),
#    as.numeric(global_regional_risk_shift[3, 5]) - as.numeric(global_regional_risk_shift[6, 5])
#  ),
#  "IVR_region_change_prop" = c(
#    as.numeric(global_regional_risk_shift[2, 6]) - as.numeric(global_regional_risk_shift[5, 6]),
#    as.numeric(global_regional_risk_shift[3, 6]) - as.numeric(global_regional_risk_shift[6, 6])
#  )
#) %>%
  # make percentages and change signs
#  mutate(
#    IVR_region_change_count = abs(IVR_region_change_count),
 #   IVR_region_change_prop = scales::label_percent(accuracy = 0.01) (abs(IVR_region_change_prop))
#    )

```

With this analysis, I found that currently, 65.2% of the global IVRs are low risk according to the global model. However, 26% of the total IVRs are specifically in quadrant 3 (high risk). This means that the global model alone would label 693 of the 1063 IVRs as low risk, when in actuality 287 or about 41.4% of these are at high risk for SLF establishment (above the MTSS threshold) when we spatially segment the presence data into an ensemble of regional-scale models. After climate change, 893 of the 1063 IVRs (82%) would be unsuitable if the global model alone were used to describe the risk of SLF. However, 341 or 39.1% of these unsuitable IVRs are still suitable in regional_scale models and thus would be missed by an analysis of risk using only a global-scale model.

This means that our regional-scale ensemble is adding resolution and nuance to our estimation of risk for SLF establishment.

# Appendix

## add percentage change to scatter plot

I will add a percentage change for each quadrant that illustrates the total loss or gain per quadrant over time. I will use the IVR_risk_table created above

*this section's code is broken dont use*

```{r % change table}

IVR_quad_percent_change <- tibble(
  "risk_quadrant" = c("extreme", "high", "moderate", "low"),
  "percent_change" = c(
    # extreme
    ((IVR_risk_table[5, 2] - rowSums(IVR_risk_table[5, 2:5])) / (IVR_risk_table[5, 2] + rowSums(IVR_risk_table[5, 2:5]))),
    # high
    ((IVR_risk_table[5, 3] - rowSums(IVR_risk_table[5, 2:5])) / (IVR_risk_table[5, 3] + rowSums(IVR_risk_table[5, 2:5]))),
    # moderate
    ((IVR_risk_table[5, 4] - rowSums(IVR_risk_table[5, 2:5])) / (IVR_risk_table[5, 4] + rowSums(IVR_risk_table[5, 2:5]))),
    ((IVR_risk_table[5, 5] - rowSums(IVR_risk_table[5, 2:5])) / (IVR_risk_table[5, 5] + rowSums(IVR_risk_table[5, 2:5])))
  )
)

```

Now I will plot these percent changes over the risk quadrants.

```{r plot SLF suitability values and % change, fig.asp = 1}

# figure annotation title
# "Risk of Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_rescaled_plot_percent_change <- ggplot(data = xy_joined_rescaled) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = global_MTSS, linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = regional_ensemble_MTSS_1995, linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # historical data
   geom_point(
     aes(x = xy_global_1995_rescaled, y = xy_regional_ensemble_1995_rescaled, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "azure4", alpha = 0.1
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055_rescaled, y = xy_regional_ensemble_2055_rescaled, shape = "2041-2070\nGFDL ssp370"), 
     size = 2, stroke = 0.5, color = "black", fill = "purple3", alpha = 0.1
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   # quadrant % changes
   # extreme
   geom_text(aes(x = 0.75, y = 0.8, label = scales::percent(as.numeric(IVR_quad_percent_change[1, 2]), accuracy = 0.1)), size = 12, color = "darkred") +
   # high
   geom_text(aes(x = 0.25, y = 0.8, label = scales::percent(as.numeric(IVR_quad_percent_change[2, 2]), accuracy = 0.1)), size = 12, color = "darkorange") +
   # moderate
   geom_text(aes(x = 0.75, y = 0.2, label = scales::percent(as.numeric(IVR_quad_percent_change[3, 2]), accuracy = 0.1)), size = 12, color = "gold") +
   # low
   geom_text(aes(x = 0.25, y = 0.2, label = scales::percent(as.numeric(IVR_quad_percent_change[4, 2]), accuracy = 0.1)), size = 12, color = "azure4") +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   #scale_fill_manual(name = "Time period", values = c("azure4", "gray27")) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot}

ggsave(
  xy_joined_rescaled_plot_percent_change, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "xy_suitability_rescaled_CC_shift_IVR_regions_global_regional_ensemble_percentages.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

## plot unweighted mean versions of suitability plots

This workflow follows steps 1-3, except that it plots based on the unweighted mean of the calculated regional ensemble (rather than an ensemble created based on AUC and ExDet weights).

```{r load in summary files}
# summary file to extract thresholds from

summary_regional_ensemble_unweighted <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_threshold_values_unweighted.csv"))


```

Finally, I will load in the global and regional ensemble suitability maps. These will be used both for extracting the new xy suitability values and for plotting.

```{r load in suitability rasters}

# regional_ensemble
regional_ensemble_1995_unweighted <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_regional_unweighted_mean_globe_1981-2010.asc")
  )

regional_ensemble_2055_unweighted <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v0", "ensemble_regional_unweighted_mean_globe_2041-2070_GFDL_ssp370.asc")
  )

```

### Retrieve/Import xy suitability

These scatter plots will be based on the suitability for the IVR points in both the global and regional_ensemble models. I have already calculated the xy suitability for the global model based on these points, using the function `slfSpread::predict_xy_suitability()`. This function will not work for the regional_ensemble because it calls for a model object, which we did not use to predict the ensemble suitability. So, I will use `terra::extract()` to perform this action.

The xy_suitabilities for the global models also contain a few less records than the IVR regions. To ensure that I am retrieving the suitability values for the exact same vineyards in the regional_ensemble, I will perform a filtering join. The IVR dataset should now contain the same records as the global suitability values do

```{r de-duplicate and filter join}

# de-duplicate just to make sure
xy_global_1995 <- scrubr::dedup(xy_global_1995, how = "one", tolerance = 0.99) 
xy_global_2055 <- scrubr::dedup(xy_global_2055, how = "one", tolerance = 0.99) 


# edits to IVR dataset
# filter join of IVRs
IVR_locations <- semi_join(IVR_locations, xy_global_1995, by = c("x", "y")) 
# add ID column
IVR_locations <- IVR_locations %>%
  dplyr::mutate(ID = row_number()) %>%
  relocate(ID, x, y)

```

Now, I will retrieve the suitability values for the regional_ensemble. Instead of returning the coordinates from the map, I will join the coordinates from the original IVR_locations dataset so that the coordinates are exact for joining with other datasets.

```{r retrieve xy suitability for regional_ensemble}

# 1995
xy_regional_ensemble_1995_unweighted <- terra::extract(
  x = regional_ensemble_1995_unweighted,
  y = dplyr::select(IVR_locations, x, y), # points
  method = "simple",
  xy = FALSE, # dont return coordinates
  ID = TRUE
)


# 2055
xy_regional_ensemble_2055_unweighted <- terra::extract(
  x = regional_ensemble_2055_unweighted,
  y = dplyr::select(IVR_locations, x, y), # points
  method = "simple",
  xy = FALSE, # dont return coordinates 
  ID = TRUE
) 

```

```{r join xy coordinates}

# joining object
IVR_coordinates <- dplyr::select(IVR_locations, ID, x, y)

# perform join
xy_regional_ensemble_1995_unweighted <- left_join(xy_regional_ensemble_1995_unweighted, IVR_coordinates, by = "ID") %>%
  dplyr::select(-ID)

xy_regional_ensemble_2055_unweighted <- left_join(xy_regional_ensemble_2055_unweighted, IVR_coordinates, by = "ID") %>%
  dplyr::select(-ID)

```

Now, I will tidy and save the datasets

```{r tidy datasets}

# regional_ensemble datasets
xy_regional_ensemble_1995_unweighted <- xy_regional_ensemble_1995_unweighted %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_1995_unweighted" = "mean") %>%
  dplyr::filter(!is.na(xy_regional_ensemble_1995_unweighted)) %>%
  dplyr::relocate(x, y)

xy_regional_ensemble_2055_unweighted <- xy_regional_ensemble_2055_unweighted %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_2055_unweighted" = "mean") %>%
  dplyr::filter(!is.na(xy_regional_ensemble_2055_unweighted)) %>%
  dplyr::relocate(x, y)

# we lost 1 record because of the NA removal, but that is OK

```

### Transform xy suitability

I will plot the suitability values in two different ways- I will plot the raw xy suitability and I will transform the data so that the MTSS threshold is the center of the scatter plot. This way, movement across the minimum suitability threshold is more easily visualized. I will transform all 4 vectors of suitability values, 2 per model, in preparation for plotting. 

I created the function `slfSpread::rescale_cloglog_suitability()` to accomplish this task. This function uses a vector of exponential transformations for the specified value of `thresh` to apply an exponential equation to the vector of suitability values. It then applies the equation `y = c1 * c2^x + c3` to the vector, where x is the input suitability values, y is the transformed version of those values, c1 and c3 are the maximum and its inverse, and c2 is the interpolated value of the input `thresh`. The transformed suitability vector is re-scaled so that `thresh` is the median (0.5) on a 0-1 scale and all other values are transformed to fit this scale.

```{r transform regional_ensemble xy data}

xy_regional_ensemble_1995_unweighted_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_1995_unweighted,
  thresh = "MTSS",
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble_unweighted,
  rescale.name = "xy_regional_ensemble_1995_unweighted",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_1995_unweighted_rescaled_thresholds <- xy_regional_ensemble_1995_unweighted_rescaled[[2]]
xy_regional_ensemble_1995_unweighted_rescaled <- xy_regional_ensemble_1995_unweighted_rescaled[[1]]



xy_regional_ensemble_2055_unweighted_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_2055_unweighted,
  thresh = "MTSS", # the way the thresholds are calculated for the regional_ensemble model means that the threshold will be slightly different for climate change
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble_unweighted,
  rescale.name = "xy_regional_ensemble_2055_unweighted",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_2055_unweighted_rescaled_thresholds <- xy_regional_ensemble_2055_unweighted_rescaled[[2]]
xy_regional_ensemble_2055_unweighted_rescaled <- xy_regional_ensemble_2055_unweighted_rescaled[[1]]

```

### plot untransformed suitability

```{r join datasets}

# join datasets for plotting
xy_joined_unweighted <- full_join(xy_global_1995, xy_regional_ensemble_1995_unweighted, by = c("x", "y")) %>%
  # join CC datasets
  full_join(., xy_global_2055, by = c("x", "y")) %>%
  full_join(., xy_regional_ensemble_2055_unweighted, by = c("x", "y")) %>%
  # order
  dplyr::relocate(x, y, xy_global_1995, xy_global_2055)

```

```{r plot SLF suitability values, fig.asp = 1}

# figure annotation title
# "suitability for Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
xy_joined_unweighted_plot <- ggplot(data = xy_joined_unweighted) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = as.numeric(summary_regional_ensemble_unweighted[2, 2]), linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # historical data
   geom_point(
     aes(x = xy_global_1995, y = xy_regional_ensemble_1995_unweighted, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055, y = xy_regional_ensemble_2055_unweighted, shape = "2041-2070\nGFDL ssp370"), 
     size = 2, stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name =  "'global' model cloglog suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name =  "'regional_ensemble' model cloglog suitability", limits = c(0, 1), breaks = breaks) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1) 

```

```{r save scatterplot}

ggsave(
  xy_joined_unweighted_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "xy_suitability_IVR_regions_global_regional_ensemble_unweighted.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

### plot transformed suitability

```{r join datasets}

# join datasets for plotting
xy_joined_unweighted_rescaled <- full_join(xy_global_1995_rescaled, xy_regional_ensemble_1995_unweighted_rescaled, by = c("x", "y")) %>%
  # join CC datasets
  full_join(., xy_global_2055_rescaled, by = c("x", "y")) %>%
  full_join(., xy_regional_ensemble_2055_unweighted_rescaled, by = c("x", "y")) %>%
  # order
  dplyr::relocate(x, y, xy_global_1995_rescaled, xy_global_2055_rescaled) %>%
  dplyr::select(-c(xy_global_1995, xy_global_2055, xy_regional_ensemble_1995_unweighted, xy_regional_ensemble_2055_unweighted))


```

I will need to create a second dataset for the arrow segments indicating change. I will filter out only the segments that cross either threshold and then plot these arrows.

First, I need to isolate the MTSS threshold values.

```{r retrieve threshold values}

# global
global_MTSS <- as.numeric(xy_global_1995_rescaled_thresholds[2, 2])
# regional ensemble
regional_ensemble_MTSS_1995_unweighted <- as.numeric(xy_regional_ensemble_1995_unweighted_rescaled_thresholds[2, 2])
regional_ensemble_MTSS_2055_unweighted <- as.numeric(xy_regional_ensemble_1995_unweighted_rescaled_thresholds[4, 2])

```

```{r plot SLF suitability values, fig.asp = 1}

# figure annotation title
# "Risk of Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_unweighted_rescaled_plot <- ggplot(data = xy_joined_unweighted_rescaled) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = global_MTSS, linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = regional_ensemble_MTSS_1995_unweighted, linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # historical data
   geom_point(
     aes(x = xy_global_1995_rescaled, y = xy_regional_ensemble_1995_unweighted_rescaled, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055_rescaled, y = xy_regional_ensemble_2055_unweighted_rescaled, shape = "2041-2070\nGFDL ssp370"), 
     size = 2,  stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot}

ggsave(
  xy_joined_unweighted_rescaled_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "SLF_risk_IVR_regions_global_regional_ensemble_unweighted.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

# References

Gallien, L., R. Douzet, S. Pratte, N. E. Zimmermann, and W. Thuiller. 2012. Invasive species distribution models – how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography 21:1126–1136.

Smith, T. 2021, August 11. Evaluating Invasion Stage with SDMs - plantarum.ca. <https://plantarum.ca/2021/08/11/invasion-stage/>.
