---
title: "Plot shift in SLF risk to important viticultural regions under climate change"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2024-08-14"
---

# Overview

In the last vignette, I created risk maps and range shift maps as the first step in visualizing and comparing the predictions made by my global and regional ensemble models. I will apply these maps for drawing on general trends of SLF suitable area shift under climate change. Downstream, I also plan to visualize these suitability maps across geographical regions and at a finer scale, such as for viticulturally important countries and provinces. 

In this vignette, I will focus on making more specific predictions for SLF establishment risk scatter plots of the suitability values for the globally important viticultural regions in the global and regional ensemble models. I will use a quadrant scatter plot to visualize the movement of these regions across the minimum threshold suitability value (the MTSS value). By plotting one model scale on each axis, we can visualize the agreement between our modeled scales. 

Gallien et al, 2012 originally defined this method for discerning the "stage of invasion" for invasive populations according to different scales of SDM. I will apply this method and re-interpret it as a method for assessing the risk of establishment. These two models will not completely agree on the suitability for a particular location, so agreement between the modeled scales can give us more confidence in that prediction. Where the modeled scales disagree, we might make different interpretations of the biological mechanism at play. 

```{r example quadrant plot, echo = FALSE}

library(tidyverse)

ggplot() +
   geom_vline(xintercept = 0.5, linetype = "dashed", linewidth = 0.7) +
   geom_hline(yintercept = 0.5, linetype = "dashed", linewidth = 0.7) + 
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), labels = c(0, 2, 4, 6, 8, 10)) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0), labels = c(0, 2, 4, 6, 8, 10)) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   theme_bw() +
   coord_fixed(ratio = 1)
   

```
**Fig. 1** Example quadrant plot for assessing SLF risk across two different scales of SDM.

We rate risk from low to extreme, based on the agreement between the modeled scales. We rate suitability in the regional_ensemble only (top left, high) as higher risk than suitability in the global model only (bottom right, moderate) because we believe that unique suitability in the regional-scale ensemble indicates a risk for adaptation to new climatic niches (Gallien et al, 2012). Where the regional-scale model detects suitability when the global model cannot, SLF might be adapting to a climatic niche outside its global mean niche. 

The MTSS threshold of suitability will be used as the central crosshairs of this plot. We will then plot the same set of points presently and in the future under our climate change scenarios. Movement across the MTSS threshold on either axis due to climate change will indicate that climate change is changing the suitability of these regions for SLF establishment. I will need to transform these suitability values so that the MTSS threshold falls on 0.5. This will allow me to visualize the movement of these regions across the minimum threshold suitability value.

First, I will plot the un-transformed version of the data and then will transform and plot the data again. I will use a table to calculate the number of regions that move across the MTSS threshold and fall into each category before and after climate change. Finally, I will save the predicted suitability values with our `wineries` dataset for downstream analysis. Our `wineries.rda` dataset contains the locations of 1,086 globally important viticultural regions.

# Setup

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
# here() starts at the root folder of this package.
library(devtools)

# spatial data handling
library(terra)
library(CoordinateCleaner)

# plot aesthetics
library(scales)
library(patchwork)
library(grid)
library(kableExtra)
library(webshot)
library(webshot2)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

```{r set wd, eval = TRUE}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

I will load in some aesthetic objects for plotting.

```{r axis breaks and labels, eval = TRUE}

breaks <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)

labels <- c(0, 2, 4, 6, 8, 10)

```

## load rasters

Load in summary files for the global and regional ensemble models, which contain the thresholds.

```{r load in summary files, eval = TRUE}
# summary file to extract thresholds from

# global
summary_global <- read_csv(file = file.path(mypath, "slf_global_v3", "global_summary_all_iterations.csv"))

summary_regional_ensemble <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_threshold_values.csv"))


```

I will load in the global and regional ensemble suitability maps. These will be used both for extracting the new xy suitability values and for plotting.

```{r load in suitability rasters, eval = TRUE}

# global
global_1995 <- terra::rast(x = file.path(mypath, "slf_global_v3", "global_pred_suit_clamped_cloglog_globe_1981-2010_mean.asc"))

global_2055 <- terra::rast(x = file.path(mypath, "slf_global_v3", "global_pred_suit_clamped_cloglog_globe_2041-2070_GFDL_ssp_averaged.asc"))


# regional_ensemble
# historical
regional_ensemble_1995 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_regional_weighted_mean_globe_1981-2010.asc")
  )
# CMIP6
## ssp 126
regional_ensemble_2055_126 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_regional_weighted_mean_globe_2041-2070_GFDL_ssp126.asc")
  )
# ssp 370
regional_ensemble_2055_370 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_regional_weighted_mean_globe_2041-2070_GFDL_ssp370.asc")
  )
# ssp 585
regional_ensemble_2055_585 <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_regional_weighted_mean_globe_2041-2070_GFDL_ssp585.asc")
  )
# ssp mean
regional_ensemble_2055_ssp_mean <- terra::rast(
  x = file.path(mypath, "slf_regional_ensemble_v1", "ensemble_regional_weighted_mean_globe_2041-2070_GFDL_ssp_averaged.asc")
  )

```

## tidy wineries dataset

 I will load in dataset containing locations of globally important viticultural regions. I will also de-duplicate the records because there were "many-to-many" joining matches downstream. De-duplicating removed about 17 records. I also remove the species column, which is automatically created by `SDMtune::predict()` upstream.

```{r load in wineries dataset}

load(file.path(here::here(), "data", "wineries.rda"))

IVR_locations <- wineries %>%
  # tidy
  dplyr::select(-c(Latitude, Longitude)) %>%
  dplyr::filter(
    !is.na(x),
    !is.na(y)
    ) %>%
  dplyr::mutate(record_type = "IVR_location")

# remove duplicate coordinates
IVR_locations <- CoordinateCleaner::clean_coordinates(
  x = IVR_locations,
  lon = "x",
  lat = "y",
  species = "record_type",
  tests = "duplicates",
  value = "clean" # just return same df without duplicates
) %>%
  dplyr::select(-record_type)

# remove unnecessary obj
rm(wineries)

```

##  import and tidy xy suitability

These scatter plots will be based on the suitability for the IVR points in both the global and regional_ensemble models. I have already calculated the xy suitability for the global model based on these points, using the function `slfSpread::predict_xy_suitability()`. This function will not work for the regional_ensemble because it calls for a model object, which we did not use to predict the ensemble suitability. So, I will use `terra::extract()` to perform this action.

I will load in the global model datasets and create the regional_ensemble datasets. I will also do some tidying of my datasets for the plots I will create.

```{r import global model suitability}

# historical
xy_global_1995 <- read_csv(
  file = file.path(mypath, "slf_global_v3", "global_wineries_1981-2010_xy_pred_suit_clamped_cloglog_mean_20000m_buffer.csv")
  ) 

# CMIP6
## ssp 126
xy_global_2055_126 <- read_csv(
  file = file.path(mypath, "slf_global_v3", "global_wineries_2041-2070_GFDL_ssp126_xy_pred_suit_clamped_cloglog_mean_20000m_buffer.csv")
  ) 
## ssp 370
xy_global_2055_370 <- read_csv(
  file = file.path(mypath, "slf_global_v3", "global_wineries_2041-2070_GFDL_ssp370_xy_pred_suit_clamped_cloglog_mean_20000m_buffer.csv")
  ) 
## ssp 585
xy_global_2055_585 <- read_csv(
  file = file.path(mypath, "slf_global_v3", "global_wineries_2041-2070_GFDL_ssp585_xy_pred_suit_clamped_cloglog_mean_20000m_buffer.csv")
  ) 

```

The xy_suitabilities for the global models also contain a few less records than the IVR regions. To ensure that I am retrieving the suitability values for the exact same vineyards in the regional_ensemble, I will perform a filtering join. The IVR dataset should now contain the same records as the global suitability values do.

```{r de-duplicate and tidy wineries dataset}

# de-duplicate just to make sure
xy_global_1995 <- xy_global_1995 %>%
  dplyr::mutate(species = "Lycorma delicatula") %>%
  CoordinateCleaner::clean_coordinates(
    x = .,
    lon = "x",
    lat = "y",
    species = "species",
    tests = "duplicates",
    value = "clean"
  ) %>%
  dplyr::select(x, y, max_cloglog_suitability) %>%
  # rename
  dplyr::rename("max_cloglog_suit_hist" = "max_cloglog_suitability") 


# edits to IVR dataset
# filter join of IVRs
IVR_locations <- semi_join(IVR_locations, xy_global_1995, by = c("x", "y")) 
# add ID column
IVR_locations <- IVR_locations %>%
  dplyr::mutate(ID = row_number()) %>%
  relocate(ID, x, y)

```

```{r save tidied data, eval = TRUE}

if (FALSE) {

  readr::write_rds(IVR_locations, file.path(here::here(), "data", "wineries_tidied.rds"))
  
}
  
#read it back in
IVR_locations <- read_rds(file.path(here::here(), "data", "wineries_tidied.rds"))


```

Now I will clean the remainder of the coordinates

```{r remove duplicates of other datasets}

# ssp 126
xy_global_2055_126 <- xy_global_2055_126 %>%
  CoordinateCleaner::clean_coordinates(
    x = .,
    lon = "x",
    lat = "y",
    species = "Species",
    tests = "duplicates",
    value = "clean"
  ) %>%
  dplyr::select(-Species)

# ssp 370
xy_global_2055_370 <- xy_global_2055_370 %>%
  CoordinateCleaner::clean_coordinates(
    x = .,
    lon = "x",
    lat = "y",
    species = "Species",
    tests = "duplicates",
    value = "clean"
  ) %>%
  dplyr::select(-Species)

# ssp 585
xy_global_2055_585 <- xy_global_2055_585 %>%
  CoordinateCleaner::clean_coordinates(
    x = .,
    lon = "x",
    lat = "y",
    species = "Species",
    tests = "duplicates",
    value = "clean"
  ) %>%
  dplyr::select(-Species)


# also rename columns
xy_global_2055_126 <- xy_global_2055_126 %>%
  dplyr::rename("max_cloglog_suit_ssp126" = "max_cloglog_suitability_20000m_buffer") 

xy_global_2055_370 <- xy_global_2055_370 %>%
  dplyr::rename("max_cloglog_suit_ssp370" = "max_cloglog_suitability_20000m_buffer") 

xy_global_2055_585 <- xy_global_2055_585 %>%
  dplyr::rename("max_cloglog_suit_ssp585" = "max_cloglog_suitability_20000m_buffer") 

```

## take mean of global model ssp scenarios 

I will take the mean of the suitability value within each of these predicted suitability datasets. These suitability values were taken from a 20km buffer around each viticultural area point to account for the potentially large size of these viticultural areas. We will keep the max values from those buffers. I will mean these values across ssp scenarios. 

```{r mean global ssp predictions}

# first join datasets
xy_global_2055_ssp_mean <- xy_global_2055_126 %>%
  left_join(., xy_global_2055_370, join_by(x, y)) %>%
  left_join(., xy_global_2055_585, join_by(x, y)) %>%
  # take mean of columns
  dplyr::mutate(max_suit_ssp_averaged = rowMeans(.[, 3:5])) %>%
  dplyr::select(x, y, max_suit_ssp_averaged)

```

```{r write as csv- global}

write_csv(
  x = xy_global_2055_ssp_mean,
  file = file.path(mypath, "slf_global_v3", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_clamped_cloglog_mean_20000m_buffer.csv")
    )

```

## retrieve suitability values for regional_ensemble

Now, I will retrieve the suitability values for the regional_ensemble using the 20km buffer method described above. I will use `buffer()` and `zonal()` from the `terra` package to first create these 20km buffers around each point, then to take zonal statistics (the max value) from each buffer zone. 

Instead of returning the coordinates from the map, I will join the coordinates from the original IVR_locations dataset so that the coordinates are exact for joining with other datasets. I will first convert this dataset to an sv object and create a buffer region around each point for extraction from each of the raster

```{r convert IVR regions dataset to sv and take buffers}

# first, convert to a vector sv in terra
IVR_locations_sv <- terra::vect(
    x = IVR_locations[, 1:3],
    crs = "EPSG:4326",
    geom = c("x", "y")
  )

# take a buffer region around each IVR location
IVR_locations_buffer <- terra::buffer(
  x = IVR_locations_sv,
  width = 20000 # 20km
)

```

Now, I will retrieve the max suitability value within that buffer for each of the IVRs in each of the predicted rasters.

```{r for loop to retrieve winery xy suitability for regional_ensemble}

# rasters to retrieve suitability from
rasters <- list(regional_ensemble_1995, regional_ensemble_2055_126, regional_ensemble_2055_370, regional_ensemble_2055_585)

# column identifiers
col_names <- c("hist", "ssp126", "ssp370", "ssp585")
# output file names
output_names <- c(
  "regional_ensemble_v1_wineries_1981-2010",
  "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp126",
  "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp370",
  "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp585"
  )


# for loop to retrieve xy suitability
for (a in seq_along(rasters)) {
  
  # import
  column_name <- paste0("max_cloglog_suit_", col_names[a], "_20000m")
  rast_hold <- rasters[[a]]
  
  # function
  xy_hold <- terra::zonal(
    x = rast_hold, # raster of predictions we created
    z = IVR_locations_buffer, # buffer zones
    fun = "max",
    na.rm = TRUE # remove NAs from buffer stats
    # ID column
    ) %>%
    # re-join ID and xy coords
    cbind(IVR_locations[, 1:3]) %>%
    dplyr::relocate(ID, x, y)
  
  # rename suit column
  colnames(xy_hold[, 4]) <- column_name
  
  # write to csv
  write_csv(
    x = xy_hold,
    file = file.path(mypath, "slf_regional_ensemble_v1", paste0(output_names[a], "_xy_pred_suit.csv"))
  )
  
  
  # success message
  cli::cli_alert_success(paste("xy suit for", col_names[a], "written to file"))
  # remove temp objects
  rm(xy_hold)
  rm(rast_hold)
  rm(column_name)
  
}

```

### take mean of ssps

Now I will re-load the data files I just created

```{r re-load xy suit files}

# historical
xy_regional_ensemble_1995 <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v1", "regional_ensemble_v1_wineries_1981-2010_xy_pred_suit.csv"))
# CMIP6
## ssp 126
xy_regional_ensemble_2055_126 <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v1", "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp126_xy_pred_suit.csv"))
## ssp 370
xy_regional_ensemble_2055_370 <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v1", "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp370_xy_pred_suit.csv"))
## ssp 585
xy_regional_ensemble_2055_585 <- read_csv(file = file.path(mypath, "slf_regional_ensemble_v1", "regional_ensemble_v1_wineries_2041-2070_GFDL_ssp585_xy_pred_suit.csv"))

```

```{r mean regional ensemble ssp predictions}

# first join datasets
xy_regional_ensemble_2055_ssp_mean <- xy_regional_ensemble_2055_126 %>%
  left_join(., xy_regional_ensemble_2055_370, join_by(ID, x, y)) %>%
  left_join(., xy_regional_ensemble_2055_370, join_by(ID, x, y)) %>%
  # take mean of columns
  dplyr::mutate(max_suit_ssp_averaged = rowMeans(.[, 4:6])) %>%
  dplyr::select(ID, x, y, max_suit_ssp_averaged)

```

```{r write as csv- regional ensemble}

write_csv(
  x = xy_regional_ensemble_2055_ssp_mean,
  file = file.path(mypath, "slf_global_v3", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.csv")
  )

```

## write to file

Now, I will tidy and save the datasets to .rds for our analysis.

```{r tidy datasets}

# global model datasets
xy_global_1995 <- xy_global_1995 %>%
  # add ID column
  cbind(., IVR_locations[, 1]) %>%
  dplyr::relocate(ID) %>%
  # rename
  dplyr::rename("xy_global_1995" = "max_cloglog_suit_hist")
  
# 
xy_global_2055_ssp_mean <- xy_global_2055_ssp_mean %>%
  # add ID column
  cbind(., IVR_locations[, 1]) %>%
  dplyr::relocate(ID) %>%
  dplyr::rename("xy_global_2055" = "max_suit_ssp_averaged") 



# regional_ensemble datasets
xy_regional_ensemble_1995 <- xy_regional_ensemble_1995 %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_1995" = "max_suit_hist_20000m") 

xy_regional_ensemble_2055_ssp_mean <- xy_regional_ensemble_2055_ssp_mean %>%
  # rename the column for future joining
  dplyr::rename("xy_regional_ensemble_2055" = "max_suit_ssp_averaged") 

```

```{r save main suitability datasets as .rds}

# regional_ensemble
readr::write_rds(
  xy_regional_ensemble_1995,  
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit.rds")
  )

readr::write_rds(
  xy_regional_ensemble_2055_ssp_mean,  
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds")
  )


# save global datasets
readr::write_rds(
  xy_global_1995,  
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit.rds")
  )

readr::write_rds(
  xy_global_2055_ssp_mean,  
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds")
  )

```

# 1. Transform xy suitability

I will plot the suitability values in two different ways- I will plot the raw xy suitability and then will transform the data so that the MTSS threshold is the center of the scatter plot. This way, movement across the minimum suitability threshold is more easily visualized. I will transform all 4 vectors of suitability values, 2 per model, in preparation for plotting.

I created the function `slfSpread::rescale_cloglog_suitability()` to accomplish this task. This function uses a vector of exponential transformations for the specified value of `thresh` to apply an exponential equation to the vector of suitability values. It then applies the equation `y = c1 * c2^x + c3` to the vector, where x is the input suitability values, y is the transformed version of those values, c1 and c3 are the maximum and its inverse, and c2 is the interpolated value of the input `thresh`. The transformed suitability vector is re-scaled so that `thresh` is the median (0.5) on a 0-1 scale and all other values are transformed to fit this scale.

```{r re-laod IVR dataset, eval = TRUE}

IVR_locations <- read_rds(file.path(here::here(), "data", "wineries_tidied.rds"))

```

```{r re-load pred suit datasets, eval = TRUE}

# global
xy_global_1995 <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit.rds"))
xy_global_2055 <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds"))

# regional
xy_regional_ensemble_1995 <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit.rds"))
xy_regional_ensemble_2055 <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds"))

```

I will use the internal function `slfSpread::rescale_cloglog_suitability()` to re-scale these suitability values. 

```{r transform global xy data, eval = TRUE}

xy_global_1995_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_global_1995,
  thresh = "MTSS",
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_global,
  rescale.name = "xy_global_1995",
  rescale.thresholds = TRUE
)
# separate data from thresholds
xy_global_1995_rescaled_thresholds <- xy_global_1995_rescaled[[2]]
xy_global_1995_rescaled <- xy_global_1995_rescaled[[1]]



xy_global_2055_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_global_2055,
  thresh = "MTSS",  # the global model only has 1 MTSS thresh
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_global,
  rescale.name = "xy_global_2055",
  rescale.thresholds = TRUE
)

xy_global_2055_rescaled_thresholds <- xy_global_2055_rescaled[[2]]
xy_global_2055_rescaled <- xy_global_2055_rescaled[[1]]

```

```{r transform regional_ensemble xy data, eval = TRUE}

xy_regional_ensemble_1995_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_1995,
  thresh = "MTSS",
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble,
  rescale.name = "xy_regional_ensemble_1995",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_1995_rescaled_thresholds <- xy_regional_ensemble_1995_rescaled[[2]]
xy_regional_ensemble_1995_rescaled <- xy_regional_ensemble_1995_rescaled[[1]]



xy_regional_ensemble_2055_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = xy_regional_ensemble_2055,
  thresh = "MTSS.CC", # the way the thresholds are calculated for the regional_ensemble model means that the threshold will be slightly different for climate change
  exponential.file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"),
  summary.file = summary_regional_ensemble,
  rescale.name = "xy_regional_ensemble_2055",
  rescale.thresholds = TRUE
)

xy_regional_ensemble_2055_rescaled_thresholds <- xy_regional_ensemble_2055_rescaled[[2]]
xy_regional_ensemble_2055_rescaled <- xy_regional_ensemble_2055_rescaled[[1]]

```

```{r save re-scaled values}

# global
write_rds(
  xy_global_1995_rescaled, 
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled.rds")
  )
write_rds(
  xy_global_2055_rescaled, 
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds")
  )

# regional
write_rds(
  xy_regional_ensemble_1995_rescaled, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled.rds")
  )
write_rds(
  xy_regional_ensemble_2055_rescaled, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds")
  )

```

```{r save thresholds}

# global
write_rds(
  xy_global_1995_rescaled_thresholds, 
  file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds")
  )
write_rds(
  xy_global_2055_rescaled_thresholds, 
  file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds")
  )

# regional
write_rds(
  xy_regional_ensemble_1995_rescaled_thresholds, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds")
  )
write_rds(
  xy_regional_ensemble_2055_rescaled_thresholds, 
  file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds")
  )

```

# 2. Plot untransformed suitability values

We need a baseline for visualizing the trends in these scatter plots, so I will first plot the un-transformed datasets.

```{r join datasets- untransformed, eval = TRUE}

# join datasets for plotting
xy_joined <- full_join(xy_global_1995, xy_regional_ensemble_1995, join_by(ID, x, y)) %>%
  # join CC datasets
  full_join(., xy_global_2055, join_by(ID, x, y)) %>%
  full_join(., xy_regional_ensemble_2055, join_by(ID, x, y)) %>%
  # order
  dplyr::relocate(ID, x, y, xy_global_1995, xy_global_2055)

```

```{r plot SLF suitability values- untransformed, fig.asp = 1, eval = TRUE}

# figure annotation title
# "suitability for Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_plot <- ggplot(data = xy_joined) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = as.numeric(summary_regional_ensemble[3, 4]), linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # historical data
   geom_point(
     aes(x = xy_global_1995, y = xy_regional_ensemble_1995, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055, y = xy_regional_ensemble_2055, shape = "2041-2070\nGFDL-ESM4\nmean ssp126/370/585"), 
     size = 2, stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name =  "'global' model cloglog suitability", limits = c(0, 1), breaks = breaks) + 
   scale_y_continuous(name =  "'regional_ensemble' model cloglog suitability", limits = c(0, 1), breaks = breaks) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot- untransformed}

ggsave(
  xy_joined_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "IVR_xy_suitability_IVR_regions_global_regional_ensemble.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

# 3. plot transformed suitability values

I will manually change the scale of these values to a 1-10 scale so that this plot of risk is not confused for a measure of suitability from the model.

```{r re-load rescaled values, eval = TRUE}

# global
xy_global_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled.rds"))

xy_global_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds"))

# regional
xy_regional_ensemble_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled.rds"))

xy_regional_ensemble_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds"))

```

```{r re-load thresholds, eval = TRUE}

# global
xy_global_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_global_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds"))

# regional
xy_regional_ensemble_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_regional_ensemble_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds"))

```


```{r join datasets- rescaled, eval = TRUE}

# join datasets for plotting
xy_joined_rescaled <- full_join(xy_global_1995_rescaled, xy_regional_ensemble_1995_rescaled, join_by(ID, x, y)) %>%
  # join CC datasets
  full_join(., xy_global_2055_rescaled, join_by(ID, x, y)) %>%
  full_join(., xy_regional_ensemble_2055_rescaled, join_by(ID, x, y)) %>%
  # order
  dplyr::relocate(ID, x, y, xy_global_1995_rescaled, xy_global_2055_rescaled) %>%
  dplyr::select(-c(xy_global_1995, xy_global_2055, xy_regional_ensemble_1995, xy_regional_ensemble_2055))


```

I will need to create a second dataset for the arrow segments indicating change. I will filter out only the segments that cross either threshold and then plot these arrows.

First, I need to isolate the MTSS threshold values.

```{r retrieve threshold values, eval = TRUE}

# global
global_MTSS <- as.numeric(xy_global_1995_rescaled_thresholds[2, 2])
# regional ensemble
regional_ensemble_MTSS_1995 <- as.numeric(xy_regional_ensemble_1995_rescaled_thresholds[2, 2])
regional_ensemble_MTSS_2055 <- as.numeric(xy_regional_ensemble_1995_rescaled_thresholds[4, 2])

```

Next, I will use `case_when()` (a vectorized ifelse) to calculate when a point shifts across a suitability threshold due to climate change.

```{r filter segments that cross thresholds, eval = TRUE}

xy_joined_rescaled_intersects <- xy_joined_rescaled %>%
  dplyr::mutate(
    crosses_threshold = dplyr::case_when(
      # conditional for starting and ending points that overlap a the threshold
      # x-axis
      xy_global_1995_rescaled > global_MTSS & xy_global_2055_rescaled < global_MTSS ~ "crosses",
      xy_global_1995_rescaled < global_MTSS & xy_global_2055_rescaled > global_MTSS ~ "crosses",
      # y-axis
      xy_regional_ensemble_1995_rescaled > regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled < regional_ensemble_MTSS_2055 ~ "crosses",
       xy_regional_ensemble_1995_rescaled < regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled > regional_ensemble_MTSS_2055 ~ "crosses",
      # else
      .default = "does not cross"
    )
  )

# filter out the crosses
xy_joined_rescaled_intersects <- dplyr::filter(
  xy_joined_rescaled_intersects,
  crosses_threshold == "crosses"
)

```

Now lets plot the data.

```{r plot SLF suitability values- rescaled, fig.asp = 1, eval = TRUE}

# figure annotation title
# "Risk of Lycorma delicatula establishment in globally important viticultural areas, projected for climate change"

# plot
(xy_joined_rescaled_plot <- ggplot(data = xy_joined_rescaled) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = global_MTSS, linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = regional_ensemble_MTSS_1995, linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # arrows indicating change
   geom_segment(
     data = xy_joined_rescaled_intersects,
     aes(
       x = xy_global_1995_rescaled,
       xend = xy_global_2055_rescaled,
       y = xy_regional_ensemble_1995_rescaled,
       yend = xy_regional_ensemble_2055_rescaled
     ), 
     arrow = grid::arrow(angle = 5.5, type = "closed"), alpha = 0.3, linewidth = 0.25, color = "black"
   ) +
   # historical data
   geom_point(
     aes(x = xy_global_1995_rescaled, y = xy_regional_ensemble_1995_rescaled, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "orchid1"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055_rescaled, y = xy_regional_ensemble_2055_rescaled, shape = "2041-2070\nGFDL-ESM4\nmean ssp126/370/585"), 
     size = 2, stroke = 0.7, color = "black", fill = "purple3"
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)
)

```

```{r save scatterplot- rescaled}

ggsave(
  xy_joined_rescaled_plot, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "IVR_risk_plot.jpg"
    ),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

# I will also save the ggplot object as an rds because I need to facet it later
write_rds(
  xy_joined_rescaled_plot,
  file = file.path(here::here(), "vignette-outputs", "figures", "figures-rds", "IVR_risk_plot.rds")
  )

```

The plot shows the projected change in suitability of our viticultural regions under climate change. The arrows indicate that a region is crossing a suitability value, from being suitable for SLF establishment to being unsuitable.

# 4. Create summary table of transformed plot

I will now create a summary table to explain the rescaled plots from step 4. The table will depict the quadrant placement of the point in the quadrant plot, both before and after climate change. From this, I will calculate the total number of movements into and out of each quadrant. I will apply the internal function `slfSpread::calculate_risk_quadrant()`. 

I will create a summary table of the quadrant placement (and thus the level of risk) for each point in the IVR_locations dataset. I will use `calculate_risk_quadrant()` to accomplish this.

```{r summary IVRs, eval = TRUE}

# create dataset and tidy
IVR_locations_joined <- left_join(IVR_locations, xy_joined_rescaled, join_by(ID, x, y)) %>%
  dplyr::relocate(ID, x, y) 
  
# calculate risk quadrants
IVR_locations_risk <- IVR_locations_joined %>%
  dplyr::mutate(
    risk_1995 = slfSpread::calculate_risk_quadrant(
      suit.x = IVR_locations_joined$xy_global_1995_rescaled,
      suit.y = IVR_locations_joined$xy_regional_ensemble_1995_rescaled,
      thresh.x = global_MTSS, # this threshold remains the same
      thresh.y = regional_ensemble_MTSS_1995
    ),
    risk_2055 = slfSpread::calculate_risk_quadrant(
      suit.x = IVR_locations_joined$xy_global_2055_rescaled,
      suit.y = IVR_locations_joined$xy_regional_ensemble_2055_rescaled,
      thresh.x = global_MTSS,
      thresh.y = regional_ensemble_MTSS_2055
    ),
    risk_shift = str_c(risk_1995, risk_2055, sep = "-")
  )

```

```{r factor ordering of risk, eval = TRUE}

risk_levels <- c("extreme", "high", "moderate", "low")

```

```{r risk categories table, eval = TRUE}

IVR_risk_table <- IVR_locations_risk %>%
  # create counts and make into acrostic table
  dplyr::group_by(risk_1995, risk_2055) %>%
  dplyr::summarize(count = n()) %>%
  pivot_wider(names_from = risk_2055, values_from = count) %>%
  # tidy
  dplyr::rename("rows_1995_cols_2055" = "risk_1995") %>%
  relocate("rows_1995_cols_2055", "extreme", "high", "moderate") %>%
  arrange(factor(.$rows_1995_cols_2055, levels = risk_levels)) %>%
  # replace missing categories with 0
  replace(is.na(.), 0) %>%
  ungroup()


# tidy
IVR_risk_table <- IVR_risk_table %>%
  # add totals column
  tibble::add_column("total_present" = rowSums(.[, 2:5])) %>%
  # add row totals
  tibble::add_row(rows_1995_cols_2055 = "total_2055", extreme = colSums(dplyr::select(., 2)), high = colSums(dplyr::select(., 3)), moderate = colSums(dplyr::select(., 4)), low = colSums(dplyr::select(., 5)), total_present = 1074) %>%
  # convert to df
  as.data.frame()

```

```{r export table}

write_csv(IVR_risk_table, file = file.path(here::here(), "vignette-outputs", "data-tables", "IVR_risk_table.csv"))

```

We now have a table calculating the number of regions per quadrant, before and after climate change!

# 5. global risk shift vs regional risk shift

I will create a table to sum the number of points in three different groups.  My goal is to understand how the regional model adds resolution to our calculation of risk. I will sum the number of points that are suitable in the global model only, unsuitable in the global model only, and unsuitable in the global model / suitable in the regional model. I will repeat this operation for both time periods.

```{r create table of sums}

global_regional_risk_shift <- tibble(
  time_period = c(1995, 1995, 1995, 2055, 2055, 2055),
  quadrants = c("quad4_quad2", "quad3_quad1", "quad3", "quad4_quad2", "quad3_quad1", "quad3"),
  risk = c("extreme_moderate", "high_low", "high", "extreme_moderate", "high_low", "high"),
  model_suit = c("global_suit", "global_unsuit", "global_unsuit_regional_suit", "global_suit", "global_unsuit", "global_unsuit_regional_suit"),
  IVR_region_count = c(
    # global suitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled >= global_MTSS),
    # global unsuitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled < global_MTSS),
    # global unsuitable and regional suitable 1995
    sum(IVR_locations_joined$xy_global_1995_rescaled < global_MTSS & IVR_locations_joined$xy_regional_ensemble_1995_rescaled >= regional_ensemble_MTSS_1995),
    # global suitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled >= global_MTSS),
    # global unsuitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled < global_MTSS),
    # global unsuitable and regional suitable 2055
    sum(IVR_locations_joined$xy_global_2055_rescaled < global_MTSS & IVR_locations_joined$xy_regional_ensemble_2055_rescaled >= regional_ensemble_MTSS_2055)
  )
)

```

```{r add proportions and make specific calculations}

# total # IVRs
total_IVR <- sum(global_regional_risk_shift[1:2, 5])

global_regional_risk_shift <- mutate(
  global_regional_risk_shift, 
  IVR_region_prop = IVR_region_count / total_IVR
  )


# calculate % of unsuit (quad3 and quad 1) that are are in quad3
quad3_risk_prop <- tibble(
  time_period = c("quad3_1995", "quad3_2055"),
  prop_total_unsuit_in_quad3 = c(
    scales::label_percent(accuracy = 0.01) (abs(as.numeric((global_regional_risk_shift[3, 5]) / global_regional_risk_shift[2, 5]))),
    scales::label_percent(accuracy = 0.01) (abs(as.numeric((global_regional_risk_shift[6, 5]) / global_regional_risk_shift[5, 5])))
  )
)



# number and proportion of IVRs shifting with CC
#global_regional_risk_shift_summary <- tibble(
##  "risk_shift" = c("global_unsuit_shift_1995_2055", "quad3_high_shift_diff_1995_2055"),
#  "IVR_region_change_count" = c(
#    as.numeric(global_regional_risk_shift[2, 5]) - as.numeric(global_regional_risk_shift[5, 5]),
#    as.numeric(global_regional_risk_shift[3, 5]) - as.numeric(global_regional_risk_shift[6, 5])
#  ),
#  "IVR_region_change_prop" = c(
#    as.numeric(global_regional_risk_shift[2, 6]) - as.numeric(global_regional_risk_shift[5, 6]),
#    as.numeric(global_regional_risk_shift[3, 6]) - as.numeric(global_regional_risk_shift[6, 6])
#  )
#) %>%
  # make percentages and change signs
#  mutate(
#    IVR_region_change_count = abs(IVR_region_change_count),
 #   IVR_region_change_prop = scales::label_percent(accuracy = 0.01) (abs(IVR_region_change_prop))
#    )

```

With this analysis, I found that currently, 65.2% of the global IVRs are low risk according to the global model. However, 26% of the total IVRs are specifically in quadrant 3 (high risk). This means that the global model alone would label 693 of the 1074 IVRs as low risk, when in actuality 287 or about 41.4% of these are at high risk for SLF establishment (above the MTSS threshold) when we spatially segment the presence data into an ensemble of regional-scale models. After climate change, 893 of the 1063 IVRs (82%) would be unsuitable if the global model alone were used to describe the risk of SLF. However, 341 or 39.1% of these unsuitable IVRs are still suitable in regional_scale models and thus would be missed by an analysis of risk using only a global-scale model.

This means that our regional-scale ensemble is adding resolution and nuance to our estimation of risk for SLF establishment.

```{r save table from above, if desired}

# add %
global_regional_risk_shift <- mutate(global_regional_risk_shift, IVR_region_prop = scales::label_percent(accuracy = 0.01) (IVR_region_prop))

# make kable
global_regional_risk_shift <- kable(global_regional_risk_shift, "html", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:2, width_min = '4cm') %>%
  kableExtra::add_header_above(., header = c("IVR risk plot quadrant proportions" = 6), bold = TRUE)


# save as .html
kableExtra::save_kable(
  global_regional_risk_shift, 
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_plot_quadrant_props.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_plot_quadrant_props.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_plot_quadrant_props.jpg"),
  zoom = 2
)
  
file.remove(file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_plot_quadrant_props.html"))


```

# 6. Save wineries with pred suitability

I will now join the risk levels dataset I have created with the original `wineries_tidied` dataset and save this. I will also add a count of the risk levels so they can be compared quantitatively. I will code extreme risk = 4, high = 3, moderate = 2, and low = 1. Then, I will subtract the risk level in 1995 from the risk level in 2055 to quantify the change in risk level over time and calculate summary stats.

```{r create dataset, eval = TRUE}

# add risk level counts
wineries_tidied_suit <- IVR_locations_risk %>%
  dplyr::mutate(
    risk_1995_count = dplyr::case_when(
      risk_1995 == "extreme" ~ 4,
      risk_1995 == "high" ~ 3,
      risk_1995 == "moderate" ~ 2,
      risk_1995 == "low" ~ 1
    ),
    risk_2055_count = dplyr::case_when(
      risk_2055 == "extreme" ~ 4,
      risk_2055 == "high" ~ 3,
      risk_2055 == "moderate" ~ 2,
      risk_2055 == "low" ~ 1
    ),
    risk_shift_count = risk_2055_count - risk_1995_count
  )

# rename columns
wineries_tidied_suit <- wineries_tidied_suit %>%
  dplyr::rename(
    "global_model_suit_hist" = "xy_global_1995_rescaled",
    "regional_ensemble_model_suit_hist" = "xy_regional_ensemble_1995_rescaled",
    "global_model_suit_2041-2070" = "xy_global_2055_rescaled",
    "regional_ensemble_model_suit_2041-2070" = "xy_regional_ensemble_2055_rescaled",
    "risk_level_hist" = "risk_1995",
    "risk_level_2041-2070" = "risk_2055",
    "risk_count_hist" = "risk_1995_count",
    "risk_count_2041-2070" = "risk_2055_count"
  )

# rearrange columns
wineries_tidied_suit <- wineries_tidied_suit %>%
  dplyr::relocate(risk_count_hist, .after = risk_level_hist) %>%
  dplyr::relocate(`risk_count_2041-2070`, .after = `risk_level_2041-2070`) %>%
  dplyr::relocate(risk_shift_count, .after = risk_shift) 

```

``` {r save wineries_tidied with risk and suitability values}

# save 
readr::write_csv(
  x = wineries_tidied_suit,
  file = file.path(here::here(), "vignette-outputs", "data-tables", "wineries_tidied_with_suit_risk.csv")
  )

readr::write_rds(
  x = wineries_tidied_suit,
  file = file.path(here::here(), "data", "wineries_tidied_with_suit_risk.rds")
  )

```

We now have risk quadrant plots and tables that we can use to assess the level of SLF establishment risk to viticulture.

# References

1. Gallien, L., Douzet, R., Pratte, S., Zimmermann, N. E., & Thuiller, W. (2012). Invasive species distribution models – how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography, 21(11), 1126–1136. https://doi.org/10.1111/j.1466-8238.2012.00768.x

2. Smith, T. 2021, August 11. Evaluating Invasion Stage with SDMs - plantarum.ca. <https://plantarum.ca/2021/08/11/invasion-stage/>.

