---
title: "Check model goodness of fit via AUC, TSS and confusion matrices"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University\n[Center for Biodiversity](http://www.biodiversitycenter.org/)\n[iEco Lab](https://www.iecolab.org/)\nsam.owens@temple.edu]"
date: "2024-08-16"
---

I will plot the AUC curves and variable contribution graphs by combining all 4 models into one plot. This will allow for a direct comparison of the AUC (goodness of fit) and var contribution between models.

# Setup

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
# here::here() starts at the root folder of this package.
library(devtools)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# html tools
library(kableExtra)
library(webshot)
library(webshot2)

```

These plots depict the activity of the variables put into the MaxEnt models. I will create a combined figure for the 3 models in the regional_ensemble

```{r set wd 1}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

```{r style object}

ensemble_colors <- c(
  "Rn (native)" = "#4daf4a",
  "Ri.NAmerica" =  "#e41a1c",
  "Ri.Asia" = "#377eb8"
)

```

```{r load model objects}

regional_native_model <- read_rds(file = file.path(mypath, "slf_regional_native_v3", "regional_native_model.rds"))

regional_invaded_model <- read_rds(file = file.path(mypath, "slf_regional_invaded_v7", "regional_invaded_model.rds"))

regional_invaded_asian_model <- read_rds(file = file.path(mypath, "slf_regional_invaded_asian_v2", "regional_invaded_asian_model.rds"))

```

```{r load testing objects}

regional_native_test <- read_rds(file = file.path(mypath, "slf_regional_native_v3", "regional_native_test.rds"))

regional_invaded_test <- read_rds(file = file.path(mypath, "slf_regional_invaded_v7", "regional_invaded_test.rds"))

regional_invaded_asian_test <- read_rds(file = file.path(mypath, "slf_regional_invaded_asian_v2", "regional_invaded_asian_test.rds"))

```

# ROC regional ensemble

```{r plot ROC per model}

regional_native_ROC <- SDMtune::plotROC(
  model = regional_native_model,
  test = regional_native_test
) %>%
  ggplot_build()


regional_invaded_ROC <- SDMtune::plotROC(
  model = regional_invaded_model,
  test = regional_invaded_test
) %>%
  ggplot_build()

regional_invaded_asian_ROC <- SDMtune::plotROC(
  model = regional_invaded_asian_model,
  test = regional_invaded_asian_test
) %>%
  ggplot_build()
           

```

```{r ROC plot- regional ensemble}

ROC_ensemble <- ggplot() +
  # native model data
  geom_line(data = regional_native_ROC$data[[1]], aes(x = x, y = y, color = "Rn (native)", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded model data
  geom_line(data = regional_invaded_ROC$data[[1]], aes(x = x, y = y, color = "Ri.NAmerica", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_asian model data
  geom_line(data = regional_invaded_asian_ROC$data[[1]], aes(x = x, y = y, color = "Ri.Asia", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # midpoint line
  geom_abline(slope = 1, linetype = "dashed") +
  # scales
  scale_x_continuous(name = "False positive rate (1 - specificity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  scale_y_continuous(name = "True positive rate (sensitivity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  labs(
    title = "ROC curve (true vs false positive rate) for 'regional_ensemble' models"
    ) +
  theme_bw() +
  # aes
  scale_color_manual(
    name = "model",
    values = ensemble_colors,
    aesthetics = "color"
  ) +
  scale_linetype_manual(
    name = "Data segment",
    values = c("solid", "dotted"),
    labels = c("Test", "Train"),
    guide = guide_legend(reverse = TRUE)
  ) +
  theme(legend.position = "bottom") +
  coord_fixed(ratio = 1)

```

```{r save plot- AUC, eval = FALSE}

ggsave(
  ROC_ensemble, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "ROC_plot_regional_ensemble.jpg"
    ),
  height = 8, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```

# Add global model to AUC plot

## ROC

```{r load plotting objects}

global_model <- read_rds(file = file.path(mypath, "slf_global_v3", "global_model.rds"))
global_test <- read_rds(file = file.path(mypath, "slf_global_v3", "global_test.rds"))

ensemble_colors <- c(
  "Rn (native)" = "#4daf4a",
  "Ri.NAmerica" =  "#e41a1c",
  "Ri.Asia" = "#377eb8",
  "global" = "black"
)


```

```{r add global ROC}

global_ROC <- SDMtune::plotROC(
  model = global_model@models[[5]], #pick one because they did not vary too much
  test = global_test
) %>%
  ggplot_build()

```

```{r ROC plot- global added}

ROC_ensemble_global <- ggplot() +
  # global
  geom_line(data = global_ROC$data[[1]], aes(x = x, y = y, color = "global", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # native model data
  geom_line(data = regional_native_ROC$data[[1]], aes(x = x, y = y, color = "Rn (native)", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded model data
  geom_line(data = regional_invaded_ROC$data[[1]], aes(x = x, y = y, color = "Ri.NAmerica", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_asian model data
  geom_line(data = regional_invaded_asian_ROC$data[[1]], aes(x = x, y = y, color = "Ri.Asia", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # midpoint line
  geom_abline(slope = 1, linetype = "dashed") +
  # scales
  scale_x_continuous(name = "False positive rate (1 - specificity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  scale_y_continuous(name = "True positive rate (sensitivity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  labs(
    title = "ROC curve (true vs false positive rate) for 'regional_ensemble' and 'global' models"
    ) +
  theme_bw() +
  # aes
  scale_color_manual(
    name = "model",
    values = ensemble_colors,
    aesthetics = "color"
  ) +
  scale_linetype_manual(
    name = "Data segment",
    values = c("solid", "dotted"),
    labels = c("Test", "Train"),
    guide = guide_legend(reverse = TRUE)
  ) +
  theme(legend.position = "bottom") +
  coord_fixed(ratio = 1)

ROC_ensemble_global

```

```{r save plot- AUC combined, eval = FALSE}

ggsave(
  ROC_ensemble_global, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "ROC_plot_regional_ensemble_global.jpg"
    ),
  height = 8, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```

I will create a table for some important summary statistics for the models. These include sensitivity, specificity, commission error, and omission error. These are calculated according to the cloglog MTSS threshold per model.

# Sensitivity and Specificity of Models

```{r set wd 2}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

We calculate these according to the cloglog MTSS threshold per model

```{r import confusion matrices}

# global
global_conf_matr <- read_csv(file = file.path(mypath, "slf_global_v3", "global_thresh_confusion_matrix_all_iterations.csv")) %>%
  # only MTSS threshold
  slice(7)


# regional models
# native
regional_native_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_native_v3", "regional_native_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_asia
regional_invaded_asian_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_asian_v2", "regional_invaded_asian_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_NAmerica
regional_invaded_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_v7", "regional_invaded_thresh_confusion_matrix.csv")) %>%
  slice(7)

```

Be sure to change the number of test positives and background negatives for each model if this changes.

```{r calculate sens and spec}

model_sens_spec <- tibble(
  metric = rep(c("sensitivity", "specificity", "commission error", "omission error"), 4),
  description = rep(c("true_positive rate", "true_negative rate", "1 - true_negative rate", "false_negative rate"), 4),
    model = c(
    rep("global", 4),
    rep("regional\nnative", 4),
    rep("regional\ninvaded NAmerica", 4),
    rep("regional\ninvaded Asia", 4)
  ),
  thresh = "MTSS.cloglog",
  thresh_value = c(
    rep(global_conf_matr$threshold_value, 4), 
    rep(regional_native_conf_matr$threshold_value, 4), 
    rep(regional_invaded_conf_matr$threshold_value, 4),
    rep(regional_invaded_asian_conf_matr$threshold_value, 4)
    ),
  test_pos = c(
    rep(155, 4), # global
    rep(54, 4),  # native
    rep(68, 4), # invaded_NAmerica
    rep(40, 4) # invaded_Asia
  ),
  bg_neg = c(
    rep(20000, 4), # global
    rep(10000, 12) # regional
  ),
  value = c(
    # global
    round(global_conf_matr$tp_mean / 155, 4), round(global_conf_matr$tn_mean / 20000, 4), round(1 - (global_conf_matr$tn_mean / 20000), 4), round((global_conf_matr$fn_mean / 20000), 4),
    # native
    round(regional_native_conf_matr$tp / 54, 4), round(regional_native_conf_matr$tn / 10000, 4), round(1 - (regional_native_conf_matr$tn / 10000), 4), round((regional_native_conf_matr$fn / 10000), 4),
    # invaded_NAmerica
    round(regional_invaded_conf_matr$tp / 68, 4), round(regional_invaded_conf_matr$tn / 10000, 4), round(1 - (regional_invaded_conf_matr$tn / 10000), 4), round((regional_invaded_conf_matr$fn / 10000), 4),
    # invaded_asia
    round(regional_invaded_asian_conf_matr$tp / 40, 4), round(regional_invaded_asian_conf_matr$tn / 10000, 4), round(1 - (regional_invaded_asian_conf_matr$tn / 10000), 4), round((regional_invaded_asian_conf_matr$fn / 10000), 4)
  )
)

```

```{r save as .csv, eval = FALSE}

write_csv(model_sens_spec, file.path(here::here(), "vignette-outputs", "data-tables", "model_sensitivity_specificity.csv"))

```

I will also write it to a kable so it can be printed

```{r save tables from above, if desired, eval = FALSE}

# add column formatting
model_sens_spec <- mutate(model_sens_spec, bg_neg = scales::label_comma() (bg_neg))

# convert to kable
model_sens_spec_kable <- knitr::kable(x = model_sens_spec, format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

# save as .html
kableExtra::save_kable(
  model_sens_spec_kable, 
  file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
  self_contained = TRUE,
  bs_theme = "simplex"
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.jpg"),
  zoom = 2
)

# model extents

model_sens_spec_kable

```

# References

1. Jiménez‐Valverde, A. (2012). Insights into the area under the receiver operating characteristic curve (AUC) as a discrimination measure in species distribution modelling. Global Ecology and Biogeography, 21(4), 498–507. https://doi.org/10.1111/j.1466-8238.2011.00683.x

2. Zou, K. H., O’Malley, A. J., & Mauri, L. (2007). Receiver-Operating Characteristic Analysis for Evaluating Diagnostic Tests and Predictive Models. Circulation, 115(5), 654–657. https://doi.org/10.1161/CIRCULATIONAHA.105.594929
