---
title: "Calculate model fit statistics"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2024-08-23"
---

I will create a table for some important summary statistics for the models. These include sensitivity, specificity, commission error, and omission error. These are calculated according to the cloglog MTSS threshold per model.

# Setup

```{r library packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here::here() # here() starts at the root folder of this package.
library(devtools)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# html tools
library(kableExtra)
library(webshot)
library(webshot2)

```

```{r set wd}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

# Sensitivity and Specificity of Models

We calculate these according to the cloglog MTSS threshold per model

```{r import confusion matrices}

# global
global_conf_matr <- read_csv(file = file.path(mypath, "slf_global_v2", "global_thresh_confusion_matrix_all_iterations.csv")) %>%
  # only MTSS threshold
  slice(7)


# regional models
# native
regional_native_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_native_v2", "regional_native_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_asia
regional_invaded_asian_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_asian_v1", "regional_invaded_asian_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_NAmerica
regional_invaded_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_v6", "regional_invaded_thresh_confusion_matrix.csv")) %>%
  slice(7)

```

```{r calculate sens and spec}

model_sens_spec <- tibble(
  metric = rep(c("sensitivity", "specificity", "commission error", "omission error"), 4),
  description = rep(c("true_positive rate", "true_negative rate", "1 - true_negative rate", "false_negative rate"), 4),
  thresh = "MTSS.cloglog",
  thresh_value = c(
    rep(global_conf_matr$threshold_value, 4), 
    rep(regional_native_conf_matr$threshold_value, 4), 
    rep(regional_invaded_conf_matr$threshold_value, 4),
    rep(regional_invaded_asian_conf_matr$threshold_value, 4)
    ),
  model = c(
    rep("global", 4),
    rep("regional\nnative", 4),
    rep("regional\ninvaded NAmerica", 4),
    rep("regional\ninvaded Asia", 4)
  ),
  test_pos = c(
    rep(155, 4), # global
    rep(328, 4),  # native
    rep(241, 4), # invaded_NAmerica
    rep(328, 4) # invaded_Asia
  ),
  bg_neg = c(
    rep(20000, 4), # global
    rep(10000, 12) # regional
  ),
  value = c(
    # global
    round(global_conf_matr$tp_mean / 155, 4), round(global_conf_matr$tn_mean / 20000, 4), round(1 - (global_conf_matr$tn_mean / 20000), 4), round((global_conf_matr$fn_mean / 20000), 4),
    # native
    round(regional_native_conf_matr$tp / 328, 4), round(regional_native_conf_matr$tn / 10000, 4), round(1 - (regional_native_conf_matr$tn / 10000), 4), round((regional_native_conf_matr$fn / 10000), 4),
    # invaded_NAmerica
    round(regional_invaded_conf_matr$tp / 241, 4), round(regional_invaded_conf_matr$tn / 10000, 4), round(1 - (regional_invaded_conf_matr$tn / 10000), 4), round((regional_invaded_conf_matr$fn / 10000), 4),
    # invaded_asia
    round(regional_invaded_asian_conf_matr$tp / 328, 4), round(regional_invaded_asian_conf_matr$tn / 10000, 4), round(1 - (regional_invaded_asian_conf_matr$tn / 10000), 4), round((regional_invaded_asian_conf_matr$fn / 10000), 4)
  )
)

```

```{r save as .csv}

write_csv(model_sens_spec, file.path(here::here(), "vignette-outputs", "data-tables", "model_sensitivity_specificity.csv"))

```

I will also write it to a kable so it can be printed

```{r save tables from above, if desired}

if(TRUE) {
  
  # add column formatting
  model_sens_spec <- mutate(model_sens_spec, bg_neg = scales::label_comma() (bg_neg))
  
  # convert to kable
model_sens_spec_kable <- knitr::kable(x = model_sens_spec, format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
  
  # save as .html
  kableExtra::save_kable(
    model_sens_spec_kable, 
    file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
    self_contained = TRUE,
    bs_theme = "simplex"
    )
  
  # initialize webshot by 
 # webshot::install_phantomjs()
  # convert to pdf
  webshot::webshot(
    url = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
    file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.jpg"),
    zoom = 2
  )
  
  # model extents
    
}

```

# References

1. Jiménez‐Valverde, A. (2012). Insights into the area under the receiver operating characteristic curve (AUC) as a discrimination measure in species distribution modelling. Global Ecology and Biogeography, 21(4), 498–507. https://doi.org/10.1111/j.1466-8238.2011.00683.x

2. Zou, K. H., O’Malley, A. J., & Mauri, L. (2007). Receiver-Operating Characteristic Analysis for Evaluating Diagnostic Tests and Predictive Models. Circulation, 115(5), 654–657. https://doi.org/10.1161/CIRCULATIONAHA.105.594929

