---
title: "Format risk tables for SLF populations and viticultural regions"
output: rmarkdown::html_vignette
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu]"
date: "2024-08-16"
---

# Overview

In this vignette, I will format an assortment of tables and figures for publication:

1. IVR risk tables
2. SLF risk tables
3. SLF risk plots- faceted by country
4. Map of globally important viticultural regions
5. bioclim variables

# Setup

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here::here() # here() starts at the root folder of this package.
library(devtools)

# spatial data handling
library(terra)
library(tidygeocoder) # analysis 7

# table aesthetics
library(scales)
library(grid)
library(formattable)
library(kableExtra)
library(webshot)
library(webshot2)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

# IVR risk table

```{r read in IVR table, eval = TRUE}

IVR_risk_table <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "IVR_risk_table.csv"), row.names = 1) %>%
  dplyr::rename(
    "extreme_2055" = "extreme",
    "high_2055" = "high",
    "moderate_2055" = "moderate",
    "low_2055" = "low"
  )

# change rownames
rownames(IVR_risk_table) <- c("extreme_present", "high_present", "moderate_present", "low_present", "total_2055")


```

I will calculate the percentage of IVRs that increase and decrease risk overall and will report this in the paper.

```{r calculate percentages- IVR risk table, eval = TRUE}

total_IVR <- sum(IVR_risk_table[1:4, 1:4])

IVR_shift_prop_table <- tibble(
  risk_shift = c("no_shift", "risk_increase", "risk_decrease"),
  prop_change = c(
    sum(IVR_risk_table[1, 1], IVR_risk_table[2, 2], IVR_risk_table[3, 3], IVR_risk_table[4, 4]) / total_IVR,
    sum(IVR_risk_table[2:4, 1], IVR_risk_table[3:4, 2], IVR_risk_table[4, 3]) / total_IVR,
    sum(IVR_risk_table[1, 2], IVR_risk_table[1:2, 3], IVR_risk_table[1:3, 4]) / total_IVR
  )
) %>%
  # make % format
  dplyr::mutate(prop_change = scales::label_percent(accuracy = 0.01) (prop_change))


```

For reporting purposes, we see that about 69.7% of the 1063 IVRs experience no change in risk of SLF establishment due to climate change by 2055. Meanwhile, about 1.7% actually experience an increase in risk due to climate change and the remaining 28.6% move down one or more levels of risk by 2055.

```{r save table from above, if desired- IVR prop plot}
  
# make kable
IVR_shift_prop_table <- knitr::kable(IVR_shift_prop_table, "html", escape = FALSE) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:2, width_min = '4cm') %>%
  kableExtra::add_header_above(., header = c("IVR risk table shift proportions" = 2), bold = TRUE)


# save as .html
kableExtra::save_kable(
  IVR_shift_prop_table, 
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table_shift_prop.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table_shift_prop.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table_shift_prop.jpg"),
  zoom = 4
)

# rm html
file.remove(file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table_shift_prop.html"))


```

Now I will format and export the table. I will add colored proportion bars to the columns, which indicate the category distribution of IVRs entering that risk category. In other words, the columns represent the risk category in 2055, so the proportion bars represent the distribution of which current risk category those points move from by 2055. I will also add positive and negative signs to indicate a risk increase or decrease. The diagonal (top L to bottom R) will have no signs because this represents no shift in risk by 2055. All points experiencing a decline in risk sit above the diagonal, and points experiencing an increase in risk sit below the diagonal.

```{r format table- IVR risk table, eval = TRUE}

# convert top half (above diagonal) to negative numbers
IVR_risk_table[1, 2] <- -(IVR_risk_table[1, 2])
IVR_risk_table[1:2, 3] <- -(IVR_risk_table[1:2, 3])
IVR_risk_table[1:3, 4] <- -(IVR_risk_table[1:3, 4])

# add positive sign to bottom half
IVR_risk_table[2:4, 1] <- sprintf("%+.0f", IVR_risk_table[2:4, 1])
IVR_risk_table[3:4, 2] <- sprintf("%+.0f", IVR_risk_table[3:4, 2])
IVR_risk_table[4, 3] <- sprintf("%+.0f", IVR_risk_table[4, 3])

# add color formatting to totals
# extreme risk
IVR_risk_table[1, 5] <- cell_spec(IVR_risk_table[1, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkred")
IVR_risk_table[5, 1] <- cell_spec(IVR_risk_table[5, 1], format = "html", bold = TRUE, escape = FALSE, color = "darkred")
# high risk
IVR_risk_table[2, 5] <- cell_spec(IVR_risk_table[2, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkorange")
IVR_risk_table[5, 2] <- cell_spec(IVR_risk_table[5, 2], format = "html", bold = TRUE, escape = FALSE, color = "darkorange")
# moderate risk
IVR_risk_table[3, 5] <- cell_spec(IVR_risk_table[3, 5], format = "html", bold = TRUE, escape = FALSE, color = "gold")
IVR_risk_table[5, 3] <- cell_spec(IVR_risk_table[5, 3], format = "html", bold = TRUE, escape = FALSE, color = "gold")
# low risk
IVR_risk_table[4, 5] <- cell_spec(IVR_risk_table[4, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkgrey")
IVR_risk_table[5, 4] <- cell_spec(IVR_risk_table[5, 4], format = "html", bold = TRUE, escape = FALSE, color = "darkgrey")

# bold total
IVR_risk_table[5, 5] <- cell_spec(IVR_risk_table[5, 5], format = "html", bold = TRUE, escape = FALSE)

# print table, e.g., in html format
IVR_risk_table <- kable(IVR_risk_table, format = "html", escape = FALSE) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:5, width_min = '4cm') %>%
  # add footnotes
  kableExtra::add_footnote("number signs indicate whether climate change is increasing (+) or decreasing (-) risk", notation = "alphabet") %>%
  # add header
  kableExtra::add_header_above(., header = c("Risk of L delicatula establishment in globally important viticultural regions" = 6), bold = TRUE)

  
```

```{r save table from above, if desired- IVR risk table}
  
# save as .html
kableExtra::save_kable(
  IVR_risk_table, 
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table.jpg"),
  zoom = 4
)

# rm html
file.remove(file.path(here::here(), "vignette-outputs", "figures", "IVR_risk_table.html"))

```

# SLF risk table

I will do the same for the SLF populations.

```{r read in SLF table, eval = TRUE}

slf_risk_table <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "slf_risk_table.csv"), row.names = 1) %>%
  dplyr::rename(
    "extreme_2055" = "extreme",
    "high_2055" = "high",
    "moderate_2055" = "moderate",
    "low_2055" = "low"
  )

# change rownames
rownames(slf_risk_table) <- c("extreme_present", "high_present", "moderate_present", "low_present", "total_2055")

```

```{r calculate percentages- SLF risk table, eval = TRUE}

total_slf <- sum(slf_risk_table[1:4, 1:4])

slf_shift_prop_table <- tibble(
  risk_shift = c("no_shift", "risk_increase", "risk_decrease"),
  prop_change = c(
    sum(slf_risk_table[1, 1], slf_risk_table[2, 2], slf_risk_table[3, 3], slf_risk_table[4, 4]) / total_slf,
    sum(slf_risk_table[2:4, 1], slf_risk_table[3:4, 2], slf_risk_table[4, 3]) / total_slf,
    sum(slf_risk_table[1, 2], slf_risk_table[1:2, 3], slf_risk_table[1:3, 4]) / total_slf
  )
) %>%
  # make % format
  dplyr::mutate(prop_change = scales::label_percent(accuracy = 0.01) (prop_change))

```

For reporting purposes, we see that about 86.2% of the 769 slf populations experience no change in risk of persistence due to climate change by 2055. Meanwhile, about 2.7% actually experience an increase in risk due to climate change and the remaining 11% move down one or more levels of risk by 2055.


```{r save table from above, if desired- SLF prop plot}
  
# make kable
slf_shift_prop_table <- kable(slf_shift_prop_table, "html", escape = FALSE) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:2, width_min = '4cm') %>%
  kableExtra::add_header_above(., header = c("SLF risk table shift proportions" = 2), bold = TRUE)


# save as .html
kableExtra::save_kable(
  slf_shift_prop_table, 
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table_shift_prop.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table_shift_prop.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table_shift_prop.jpg"),
  zoom = 4
)

# rm html
file.remove(file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table_shift_prop.html"))

```

```{r format table- SLF risk table, eval = TRUE}

# convert top half (above diagonal) to negative numbers
slf_risk_table[1, 2] <- -(slf_risk_table[1, 2])
slf_risk_table[1:2, 3] <- -(slf_risk_table[1:2, 3])
slf_risk_table[1:3, 4] <- -(slf_risk_table[1:3, 4])

# add positive sign to bottom half
slf_risk_table[2:4, 1] <- sprintf("%+.0f", slf_risk_table[2:4, 1])
slf_risk_table[3:4, 2] <- sprintf("%+.0f", slf_risk_table[3:4, 2])
slf_risk_table[4, 3] <- sprintf("%+.0f", slf_risk_table[4, 3])

# add color formatting to totals
# extreme risk
slf_risk_table[1, 5] <- cell_spec(slf_risk_table[1, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkred")
slf_risk_table[5, 1] <- cell_spec(slf_risk_table[5, 1], format = "html", bold = TRUE, escape = FALSE, color = "darkred")
# high risk
slf_risk_table[2, 5] <- cell_spec(slf_risk_table[2, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkorange")
slf_risk_table[5, 2] <- cell_spec(slf_risk_table[5, 2], format = "html", bold = TRUE, escape = FALSE, color = "darkorange")
# moderate risk
slf_risk_table[3, 5] <- cell_spec(slf_risk_table[3, 5], format = "html", bold = TRUE, escape = FALSE, color = "gold")
slf_risk_table[5, 3] <- cell_spec(slf_risk_table[5, 3], format = "html", bold = TRUE, escape = FALSE, color = "gold")
# low risk
slf_risk_table[4, 5] <- cell_spec(slf_risk_table[4, 5], format = "html", bold = TRUE, escape = FALSE, color = "darkgrey")
slf_risk_table[5, 4] <- cell_spec(slf_risk_table[5, 4], format = "html", bold = TRUE, escape = FALSE, color = "darkgrey")

# bold total
slf_risk_table[5, 5] <- cell_spec(slf_risk_table[5, 5], format = "html", bold = TRUE, escape = FALSE)

# print table, e.g., in html format
slf_risk_table <- kable(slf_risk_table, "html", escape = FALSE) %>% 
  kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:5, width_min = '4cm') %>%
  # add footnotes
  kableExtra::add_footnote("number signs indicate whether climate change is increasing (+) or decreasing (-) risk", notation = "alphabet") %>%
  # add header
  add_header_above(., header = c("Risk of persistence for known L delicatula populations" = 6), bold = TRUE)

  
```


```{r save table from above, if desired- SLF risk table}
  
# save as .html
kableExtra::save_kable(
  slf_risk_table, 
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table.jpg"),
  zoom = 4
)

# rm html
file.remove(file.path(here::here(), "vignette-outputs", "figures", "slf_risk_table.html"))

```

# SLF quadrant plots- facet by country

I want to segment the SLF presences by country and create risk plots similar to what I have for IVR regions.

```{r load in SLF dataset}

slf_populations <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05.rds"))

```

## Geocode country codes

```{r geocode country}

# geocode
slf_populations_geocoded <- tidygeocoder::reverse_geocode(
  .tbl = slf_populations, 
  lat = y, 
  long = x, 
  full_results = TRUE, # give each element of the address in a column
  method = "arcgis",
  progress_bar = TRUE
  )

# get unique values- should only be 4 countries
unique(slf_populations_geocoded$CountryCode)

# filter out presences with no country code
slf_populations_geocoded_tidy <- slf_populations_geocoded %>%
  dplyr::filter(!CountryCode %in% c("", "NA"), !is.na(CountryCode)) %>%
  dplyr::select(species, x, y, CountryCode) 

slf_populations_geocoded_tidy$CountryCode <- toupper(slf_populations_geocoded_tidy$CountryCode)

nrow(slf_populations_geocoded_tidy)
# we only lost about 4 records

# save
write_csv(
  x = slf_populations_geocoded_tidy, 
  file = file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2024-08-05_withCountries.csv")
  )

# also write to rds
readr::write_rds(
  x = slf_populations_geocoded_tidy, 
  file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05_withCountries.rds")
)

```

```{r re-import geocoded data, eval = TRUE}

slf_populations_geocoded_tidy <- readr::read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2024-08-05_withCountries.rds"))

```

```{r summarize slf records per country, eval = TRUE}

slf_populations_geocoded_table <- slf_populations_geocoded_tidy %>%
  dplyr::group_by(CountryCode) %>%
  dplyr::summarize(slf_record_count = n()) %>%
  dplyr::arrange(desc(slf_record_count)) %>%
  dplyr::mutate(country_name = c("United_States", "China", "South_Korea", "Japan", "North_Korea")) %>%
  dplyr::rename("country_code" = "CountryCode") %>%
  dplyr::relocate(country_code, country_name) %>%
  tibble::add_row(country_code = "Total", country_name = "-", slf_record_count = sum(.$slf_record_count))

```

```{r kable of slf records per country}
  
slf_populations_geocoded_table[6, 1] <- cell_spec(slf_populations_geocoded_table[6, 1], format = "html", bold = TRUE, escape = FALSE)

# make kable
slf_populations_geocoded_kable <- knitr::kable(slf_populations_geocoded_table, "html", escape = FALSE) %>% 
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = FALSE) %>%
  # standardize col width
  kableExtra::column_spec(1:3, width_min = '4cm') %>%
  kableExtra::add_header_above(., header = c("SLF population counts per country" = 3), bold = TRUE) 

# save as .html
kableExtra::save_kable(
  slf_populations_geocoded_kable, 
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_records_per_country.html"),
  self_contained = TRUE
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "slf_records_per_country.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "slf_records_per_country.jpg"),
  zoom = 2
)

# rm html
file.remove(file.path(here::here(), "vignette-outputs", "figures", "slf_records_per_country.html"))

```

## Load in suitability data

```{r re-load rescaled values, eval = TRUE}

# global
xy_global_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "global_slf_all_coords_1981-2010_xy_pred_suit_rescaled.rds"))

xy_global_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "global_slf_all_coords_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds"))

# regional
xy_regional_ensemble_1995_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_slf_all_coords_1981-2010_xy_pred_suit_rescaled.rds"))

xy_regional_ensemble_2055_rescaled <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_slf_all_coords_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled.rds"))

```

```{r re-load thresholds, eval = TRUE}

# global
xy_global_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_slf_all_coords_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_global_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "global_slf_all_coords_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds"))

# regional
xy_regional_ensemble_1995_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_slf_all_coords_1981-2010_xy_pred_suit_rescaled_thresholds.rds"))

xy_regional_ensemble_2055_rescaled_thresholds <- read_rds(file = file.path(here::here(), "data", "regional_ensemble_slf_all_coords_2041-2070_GFDL_ssp_mean_xy_pred_suit_rescaled_thresholds.rds"))

```

```{r join datasets, eval = TRUE}

# join datasets for plotting
xy_joined_rescaled <- full_join(xy_global_1995_rescaled, xy_regional_ensemble_1995_rescaled, join_by(ID, x, y)) %>%
  # join CC datasets
  full_join(., xy_global_2055_rescaled, join_by(ID, x, y)) %>%
  full_join(., xy_regional_ensemble_2055_rescaled, join_by(ID, x, y)) %>%
  # order
  dplyr::relocate(ID, x, y, xy_global_1995_rescaled, xy_global_2055_rescaled) %>%
  dplyr::select(-c(xy_global_1995, xy_global_2055, xy_regional_ensemble_1995, xy_regional_ensemble_2055))


```

Isolate threshold values

```{r retrieve threshold values, eval = TRUE}

# global
global_MTSS <- as.numeric(xy_global_1995_rescaled_thresholds[2, 2])
# regional ensemble
regional_ensemble_MTSS_1995 <- as.numeric(xy_regional_ensemble_1995_rescaled_thresholds[2, 2])
regional_ensemble_MTSS_2055 <- as.numeric(xy_regional_ensemble_2055_rescaled_thresholds[4, 2])

```

## for loop to create plots

Now lets plot the data. I will create 4 plots, 1 per country using a for loop.

```{r plot SLF suitability values, fig.asp = 1, eval = TRUE}

# create empty list for appending
slf_plots_output <- list()

# plotting objects
breaks <- c(0.0, 0.2, 0.4, 0.6, 0.8, 1.0)

labels <- c(0, 2, 4, 6, 8, 10)


# for loop
for(i in unique(slf_populations_geocoded_tidy$CountryCode)) {

  ## create datasets
  
  # filter datasets
  slf_populations_internal <- slf_populations_geocoded_tidy %>%
    dplyr::filter(CountryCode == i)
  
  # filter out only records from country using join
  xy_joined_rescaled_internal <- xy_joined_rescaled %>%
    right_join(., slf_populations_internal, join_by(x, y)) %>%
    dplyr::select(-c(species, CountryCode))
  
  
  ## boundary crossing arrows 
  
  # first get the ones that cross a threshold
  slf_intersects <- xy_joined_rescaled_internal %>%
    dplyr::mutate(
      crosses_threshold = dplyr::case_when(
        # conditional for starting and ending points that overlap a the threshold
        # x-axis
        xy_global_1995_rescaled > global_MTSS & xy_global_2055_rescaled < global_MTSS ~ "crosses",
        xy_global_1995_rescaled < global_MTSS & xy_global_2055_rescaled > global_MTSS ~ "crosses",
        # y-axis
        xy_regional_ensemble_1995_rescaled > regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled < regional_ensemble_MTSS_2055 ~ "crosses",
         xy_regional_ensemble_1995_rescaled < regional_ensemble_MTSS_2055 & xy_regional_ensemble_2055_rescaled > regional_ensemble_MTSS_2055 ~ "crosses",
        # else
        .default = "does not cross"
      )
    )
  
  # filter out the crosses
  slf_intersects <- dplyr::filter(
    slf_intersects,
    crosses_threshold == "crosses"
  )
  
  
  
  
  # plot
   slf_plot <- ggplot(data = xy_joined_rescaled_internal) +
   # threshold lines
   # MTSS thresholds
   geom_vline(xintercept = global_MTSS, linetype = "dashed", linewidth = 0.7) + # global
   geom_hline(yintercept = regional_ensemble_MTSS_1995, linetype = "dashed", linewidth = 0.7) + # regional_ensemble- there are two MTSS thresholds for this model, but the difference is so small that you will never see it on the plot
   # arrows indicating change
   geom_segment(
     data = slf_intersects,
     aes(
       x = xy_global_1995_rescaled,
       xend = xy_global_2055_rescaled,
       y = xy_regional_ensemble_1995_rescaled,
       yend = xy_regional_ensemble_2055_rescaled
     ), 
     arrow = grid::arrow(angle = 5.5, type = "closed"), alpha = 0.3, linewidth = 0.25, color = "black"
   ) +
   # historical data
   geom_point(
     aes(x = xy_global_1995_rescaled, y = xy_regional_ensemble_1995_rescaled, shape = "Present"), 
     size = 2, stroke = 0.7, color = "black", fill = "white"
     ) +
   # GFDL ssp370 data
   geom_point(
     aes(x = xy_global_2055_rescaled, y = xy_regional_ensemble_2055_rescaled, shape = "2041-2070\nGFDL ssp370"), 
     size = 2, stroke = 0.7, color = "black", fill = "wheat3"
     ) +
   # axes scaling
   scale_x_continuous(name = "'global' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) + 
   scale_y_continuous(name = "'regional_ensemble' model risk projection", limits = c(0, 1), breaks = breaks, labels = labels) +
   # quadrant labels
   # extreme risk, top right, quad4
   geom_label(aes(x = 0.75, y = 0.9, label = "extreme risk"), fill = "darkred", color = "azure", size = 5) +
   # high risk, top left, quad3
   geom_label(aes(x = 0.25, y = 0.9, label = "high risk"), fill = "darkorange", color = "azure", size = 5) +
   # moderate risk, bottom right, quad2
   geom_label(aes(x = 0.75, y = 0.1, label = "moderate risk"), fill = "gold", color = "azure", size = 5) +
   # low risk, bottom left, quad1
   geom_label(aes(x = 0.25, y = 0.1, label = "low risk"), fill = "azure4", color = "azure", size = 5) +
   # aesthetics
   scale_shape_manual(name = "Time period", values = c(21, 21)) +
   guides(shape = guide_legend(nrow = 1, override.aes = list(size = 2.5), reverse = TRUE)) +
   theme_bw() +
   theme(legend.position = "bottom", panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
   coord_fixed(ratio = 1)

  
  
  # append to output
  slf_plots_output[[i]] <- slf_plot
  
  
  # remove objects
  rm(slf_plot)
  rm(slf_intersects)
  rm(slf_populations_internal)
  rm(xy_joined_rescaled_internal)
  
}
  
```

## patchwork figure

```{r edit plots, eval = TRUE}

# USA
USA_risk_plot <- slf_plots_output[["USA"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold", size = 20)
    )

# edits
# remove geom_text
USA_risk_plot <- USA_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
USA_risk_plot$data[[4]]$size <- 1.5
USA_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#USA_risk_plot$data[[3]]$linewidth <- 0.1

USA_risk_plot <- ggplot_gtable(USA_risk_plot) %>%
  wrap_ggplot_grob()


# france
CHN_risk_plot <- slf_plots_output[["CHN"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "B") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold", size = 20)
    )

# edits
# remove geom_text
CHN_risk_plot <- CHN_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
CHN_risk_plot$data[[4]]$size <- 1.5
CHN_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#CHN_risk_plot$data[[3]]$linewidth <- 0.1

CHN_risk_plot <- ggplot_gtable(CHN_risk_plot) %>%
  wrap_ggplot_grob()

# JPN
JPN_risk_plot <- slf_plots_output[["JPN"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "C") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold", size = 20)
    )

JPN_risk_plot <- JPN_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
JPN_risk_plot$data[[4]]$size <- 1.5
JPN_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#JPN_risk_plot$data[[3]]$linewidth <- 0.1

JPN_risk_plot <- ggplot_gtable(JPN_risk_plot) %>%
  wrap_ggplot_grob()


# italy
KOR_risk_plot <- slf_plots_output[["KOR"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "D") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold", size = 20)
    )

KOR_risk_plot <- KOR_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
KOR_risk_plot$data[[4]]$size <- 1.5
KOR_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#KOR_risk_plot$data[[3]]$linewidth <- 0.1

KOR_risk_plot <- ggplot_gtable(KOR_risk_plot) %>%
  wrap_ggplot_grob()
  
  
```

```{r patchwork plots, eval = TRUE}

slf_plot_patchwork <- (
  USA_risk_plot + plot_spacer() + CHN_risk_plot +
  JPN_risk_plot + plot_spacer() + KOR_risk_plot 
  ) +
  # annotation
  plot_annotation(
    title = "SLF risk plots by country",
    subtitle = "USA | CHN | JPN | KOR"
    ) +
  plot_layout(ncol = 3, nrow = 2, widths = unit(c(5.5, 1, 5.5), "cm"), heights = unit(5.5, "cm"))


```

```{r save plot- SLF patchwork}

ggsave(
  slf_plot_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "slf_risk_plots_by_country.jpg"),
  height = 8, 
  width = 8,
  device = "jpeg",
  dpi = "retina"
  )

```

# viticultural regions map

```{r ggplot object for map style, eval = TRUE}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(legend.position = "bottom",
        panel.background = element_rect(fill = "lightblue2",
                                        colour = "lightblue2")
        ),
  scale_x_continuous(expand = c(0, 0)),
  scale_y_continuous(expand = c(0, 0)),
  viridis::scale_fill_viridis(option = "D"),
  coord_equal()
)

```

```{r load in files for plotting, eval = TRUE}

mypath <- file.path(here::here() %>% 
                     dirname(),
                   "maxent/historical_climate_rasters/chelsa2.1_30arcsec/v1_maxent_10km")

# background layer
global_bio2_df <- terra::rast(x = file.path(mypath, "bio2_1981-2010_global.asc")) %>%
  terra::as.data.frame(., xy = TRUE)


# load wineries
IVR_locations <- readr::read_rds(file.path(here::here(), "data", "wineries_tidied.rds"))

```

```{r plot points, eval = TRUE}

IVR_points_plot <- ggplot() +
    geom_raster(data = global_bio2_df, aes(x = x, y = y), fill = "azure1") +
    geom_point(data = IVR_locations, aes(x = x, y = y), fill = "purple3", shape = 21, size = 1, stroke = 0.3) +
    ggtitle("globally important viticultural regions") +
    map_style +
    theme(legend.position = "none") 

```

```{r save plot- IVR regions}

ggsave(
  IVR_points_plot, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "global_viticultural_regions.jpg"),
  height = 8, 
  width = 12,
  device = "jpeg",
  dpi = "retina"
  )

```

# Plot bioclim variables

```{r set wd, eval = TRUE}

mypath <- file.path(here::here() %>% 
                     dirname(),
                   "maxent/historical_climate_rasters/chelsa2.1_30arcsec/v1_maxent_10km")

```

```{r map style, eval = TRUE}

map_style <- list(
  xlab("longitude"),
  ylab("latitude"),
  theme_classic(),
  theme(
    # legend
    legend.position = "bottom",
    # background
    panel.background = element_rect(fill = "azure3",
                                        colour = "azure3")
  ),
  scale_x_continuous(expand = c(0, 0)),
  scale_y_continuous(expand = c(0, 0)),
  coord_equal()
)

```

```{r import datasets, eval = TRUE}

# bio 11
bio11_hist_df <- terra::rast(
  x = file.path(mypath, "bio11_1981-2010_global.asc")
) %>%
  terra::as.data.frame(xy = TRUE)

```

```{r bio 11 plot, eval = TRUE}

bio11_hist_plot <- ggplot() +
  geom_raster(data = bio11_hist_df, aes(x = x, y = y, fill = `CHELSA_bio11_1981-2010_V.2.1`)) +
  map_style +
  theme(legend.position = "none")

```


