---
title: "Retrieve host plant records from GBIF and tidy them for MaxEnt modeling"
author: "Sam Owens"
date: "2023-07-12"
contact: 'sam.owens@temple.edu'
output: html_document
---

This vignette will follow the same structure as part 1.1 of the SLF data retrieval vignette. As a result, this vignette will be less detailed about the choices that were made. For specific reasoning regarding workflow and functions used, please see the vignette `02_retrieve_SLF_records.Rmd`.

Occurrence records for both *Ailanthus altissima* (TOH) and *Vitis spp* will be retrieved from `GBIF`. I am interested in the potential effects of host plant distribution on both the current *Lycorma delicatula* invasion stage and in the future due to climate change. We will be using the distribution of these host plants to upweight the final models. Upweighted versions of these models will be considered in the final invasion stage analysis. 

# Setup

```{r load necesssary packages, echo = FALSE}

library(tidyverse)  #data manipulation

library(here) #making directory pathways easier on different instances
here()
# here() is "C:/Users/tun83449/OneDrive - Temple University/Shared drives/slfClimate/projects/slfSpread/slfSpread_pkg"

library(taxize) # get taxonomic ids
library(spocc) #query gbif and format as a dataframe

library(humboldt) # spatial thinning of points
library(tcltk) #humboldt progress bar
library(scrubr) #clean records for gbif data

library(patchwork) #nice plots

```

# 1. Retrieve and tidy TOH data

```{r get TOH gbif taxa ID}

# get species ID from gbif database
ids <- taxize::get_ids(sci_com = "Ailanthus altissima", db = "gbif")

```

The retrieved ID is `3190653`. This matches the ID for the [TOH repository](https://www.gbif.org/species/3190653) on GBIF.

### 1.1- Retrieve and save records- TOH

```{r retrieve toh records from GBIF}

if(FALSE) {

  toh_gbif <- spocc::occ(ids = ids[[1]], # search by ID, not species name
                         from = 'gbif', 
                         limit = 1e5, # gbif limits number of queries to this
                         has_coords = TRUE, # only those with attached coordinate data
                         throw_warnings = TRUE,
                         # list of commands to pass to rgbif::occ_data()
                         gbifopts = list(
                           occurrenceStatus = "PRESENT",
                           # all records except fossil records
                           basisOfRecord = c("HUMAN_OBSERVATION", "MATERIAL_CITATION", "MATERIAL_SAMPLE", "LIVING_SPECIMEN", "MACHINE_OBSERVATION", "OBSERVATION", "PRESERVED_SPECIMEN", "OCCURRENCE"),
                           hasGeospatialIssue = FALSE,
                           year = "1981, 2023"
                           ))
  
}

```

```{r save raw query and data}
  
if(FALSE) {

  # tibble raw queries
  toh_gbif_final <- as_tibble(toh_gbif$gbif$data$`3190653`) 
  
  # de-listify TOH
  toh_gbif_final$networkKeys <- NULL
  
  # save raw queries with date stamp as current date
  write_csv(x = toh_gbif_final, 
            file = file.path(here(), "data-raw", paste0( "toh_gbif_", format(Sys.Date(), "%Y-%m-%d"), ".csv")))
    
  # convert occ data to dataframe
  toh_gbif_coords1 <- spocc::occ2df(toh_gbif)
    
  # save coords as-is
  write_csv(x = toh_gbif_coords1, 
            file = file.path(
              here(), "data-raw", 
              paste0("toh_gbif_raw_coords_", format(Sys.Date(), "%Y-%m-%d"), ".csv"))
            )

}

```

```{r read in raw data}

toh_gbif_coords1 <- read_csv(file = file.path(here(), "data-raw", "toh_gbif_raw_coords_2023-07-14.csv"))

```

### 1.2- Coordinate veracity and spatial thinning- TOH

```{r coordinate veracity}

toh_gbif_coords1 <- toh_gbif_coords1 %>%
  coord_incomplete() %>%
  coord_impossible() %>%
  coord_unlikely()

```

```{r humboldt spatial thinning}

if(FALSE) {
  
  toh_gbif_coords2 <- humboldt::humboldt.occ.rarefy(in.pts = toh_gbif_coords1, 
                                             colxy = 2:3, # coordinate columns
                                             rarefy.dist = 1, 
                                             rarefy.units = "km", 
                                             run.silent.rar = F) # display progress bars
   
}

```

We use `humboldt::humboldt.occ.rarefy()` to perform the spatial thinning because it preserves the non-coordinate data. The spatial thinning left about 37,000 records out of 106,000.

```{r de-duplicate}

# de-duplicate
toh_gbif_coords3 <- toh_gbif_coords2 %>%
  scrubr::dedup(how = "one", tolerance = 0.99)

```

Unlike for SLF, TOH contains about 8 alternative species names. Since MaxEnt will treat other names as other species, we will change the name to the most common one, "Ailanthus altissima (Mill.) Swingle". This species name also matches the species name found on the GBIF repo (linked above). We will change the naming to be consistent and then save the output as the cleaned coordinates.

```{r tidy and save cleaned data}

# check species name
unique(toh_gbif_coords3$name)

# there are over 8- rename as the most common
toh_gbif_coords3$name <- "Ailanthus altissima (Mill.) Swingle"

# save
write_csv(toh_gbif_coords3, file = file.path(here(), "data", "toh_gbif_cleaned_coords_v1.csv"))

```

We will perform a final round of spatial thinning using `spThin::thin()`, just to be consistent with the SLF data.

```{r read in data}

toh_gbif_coords4 <- read_csv(file = file.path(here(), "data", "toh_gbif_cleaned_coords_v1.csv"))

```

```{r spThin spatial thinning}

if(TRUE) {

  toh_gbif_coords4 <- spThin::thin(loc.data = toh_gbif_coords4, 
                              lat.col = "latitude", long.col = "longitude",
                              spec.col = "name",
                              thin.par = 1,
                              reps = 10, # number of passes
                              # line below returns a list of data frames with the locations were preserved
                              locs.thinned.list.return = TRUE, 
                              write.files = TRUE, # returns list of thinned data as a separate .csv file
                              # next, create a file that logs the thinning run, write the path to and name of the log file
                              write.log.file = TRUE, 
                              log.file = file.path(here(), "vignette-outputs", "data-tables", paste0("toh_coords_cleaned_thinning_log_", format(Sys.Date(), "%Y-%m-%d"), ".txt")),
                              # finally, tell it where to write the new files and what to call the thinned dataset
                              out.dir = file.path(here(), "vignette-outputs", "data-tables"),
                              out.base = paste0("toh_coords_cleaned_", format(Sys.Date(), "%Y-%m-%d")),
                              verbose = TRUE # give details of run in the console
                              )

}

```

```{r export for maxent}

# reformat for MaxEnt
toh_gbif_coords4 <- toh_gbif_coords4 %>%
  dplyr::select(name, latitude, longitude, everything()) %>% # change order of columns to MaxEnt convention
  dplyr::rename(species = name, x = longitude, y = latitude) %>% # change column names to MaxEnt convention
  dplyr::select(species, x, y) # select only columns needed for MaxEnt

# save
write_csv(toh_gbif_coords4, file = file.path(here(), "data", "toh_gbif_final_coords_v1.csv"))

```

The occurrence data for TOH are now in a useable format for MaxEnt.

### 1.3- Visualize spatial thinning

First, we will visualize the results of the spatial thinning process.

```{r visualize raw and thinned GBIF records}

# raw data
toh_raw <- read_csv(file = file.path(here(), "data-raw", "toh_gbif_raw_coords_2023-07-14.csv"))

# thinned data
toh_thinned <- read_csv(file = file.path(here(), "data", "toh_gbif_cleaned_coords_v1.csv"))

# plot world map
map_toh_thinned <- ggplot() +
  # the second line is a basemap from ggplot2
  geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
  geom_point(data = toh_raw, aes(x = longitude, y = latitude), color = "blue", size = 2) +
  geom_point(data = toh_thinned, aes(x = longitude, y = latitude), color = "red", shape = 2) +
  coord_quickmap(xlim = c(-164.5, 163.5), ylim = c(-55, 85)) +
  ggtitle("TOH GBIF records- \n raw (blue) vs thinned (red)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  theme_bw()

# plot map of NAmerica
map_toh_thinned_NAmerica <- ggplot() +
  # the second line is a basemap from ggplot2
  geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
  geom_point(data = toh_raw, aes(x = longitude, y = latitude), color = "blue", size = 2) +
  geom_point(data = toh_thinned, aes(x = longitude, y = latitude), color = "red", shape = 2) +
  coord_quickmap(xlim = c(-133.593750, -52.294922), ylim = c(25.085599, 55.304138)) +
  ggtitle("TOH GBIF records- \n raw (blue) vs thinned (red)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  theme_bw()
  
# patchwork display of plots
map_toh_thinned + map_toh_thinned_NAmerica +
  plot_layout(ncol = 2)

```

From the mapping, it seems that the spatial thinning reduced the GBIF records, while preserving the same spatial extent. We also zoom in on the SLF invaded range, and indeed every blue point is paired with at least 1 red point. So, our goal was met!

# 2. Retrieve and tidy Vitis data

```{r get vitis gbif taxa ID}

# get species ID from gbif database
ids <- taxize::get_ids(sci_com = "Vitis", db = "gbif")

```

The retrieved ID is `7467468`. This matches the ID for the [Vitis repository](https://www.gbif.org/species/7467468) on GBIF.

I chose to retrieve records at the genus level, rather than the species level, because there isn't clear evidence in the literature that SLF prefers one species over another. These records may represent cultivated species, but the objective is to characterize the natural distribution of uncultivated species within this genus. While vineyards (monocultures) may be at higher risk, it is far more likely that these wild grape species will have a significant effect on SLF spread because of their global distribution. I will be referring to the entire genus of Vitis as "Vitis spp." moving forward.

### 2.1- Retrieve and save records- Vitis

```{r retrieve vitis records from GBIF}

if(FALSE) {

  vitis_gbif <- spocc::occ(ids = ids[[1]], # search by ID, not species name
                         from = 'gbif', 
                         limit = 1e5, # gbif limits number of queries to this
                         has_coords = TRUE, # only those with attached coordinate data
                         throw_warnings = TRUE,
                         # list of commands to pass to rgbif::occ_data()
                         gbifopts = list(
                           occurrenceStatus = "PRESENT",
                           # all records except fossil records
                           basisOfRecord = c("HUMAN_OBSERVATION", "MATERIAL_CITATION", "MATERIAL_SAMPLE", "LIVING_SPECIMEN", "MACHINE_OBSERVATION", "OBSERVATION", "PRESERVED_SPECIMEN", "OCCURRENCE"),
                           hasGeospatialIssue = FALSE,
                           year = "1981, 2023"
                           ))
  
}

```

```{r save raw query and data}

if(FALSE) {

  # tibble raw queries
  vitis_gbif_final <- as_tibble(vitis_gbif$gbif$data$`7467468`) 
  
  # save raw queries with date stamp as current date
  write_csv(x = vitis_gbif_final, 
            file = file.path(here(), "data-raw", paste0( "vitis_gbif_", format(Sys.Date(), "%Y-%m-%d"), ".csv")))
    
  # convert occ data to dataframe
  vitis_gbif_coords1 <- spocc::occ2df(vitis_gbif)
    
  # save coords as-is
  write_csv(x = vitis_gbif_coords1, 
            file = file.path(
              here(), "data-raw", 
              paste0("vitis_gbif_raw_coords_", format(Sys.Date(), "%Y-%m-%d"), ".csv"))
            )

}

```

```{r read in raw data}

vitis_gbif_coords1 <- read_csv(file = file.path(here(), "data-raw", "vitis_gbif_raw_coords_2023-12-07.csv"))

```

### 2.2- Coordinate veracity and spatial thinning- Vitis

```{r coordinate veracity}

vitis_gbif_coords1 <- vitis_gbif_coords1 %>%
  coord_incomplete() %>%
  coord_impossible() %>%
  coord_unlikely()

```

```{r spatial thinning}

if(FALSE) {
  
  vitis_gbif_coords2 <- humboldt::humboldt.occ.rarefy(in.pts = vitis_gbif_coords1, 
                                             colxy = 2:3, # coordinate columns
                                             rarefy.dist = 1, 
                                             rarefy.units = "km", 
                                             run.silent.rar = F) # display progress bars
  
}

```

### 2.3- Convert to format for MaxEnt and save- Vitis

```{r de-duplicate}

# de-duplicate
vitis_gbif_coords3 <- vitis_gbif_coords2 %>%
  scrubr::dedup(how = "one", tolerance = 0.99)

```

```{r save cleaned data and convert to maxent format}

# check species name
unique(vitis_coords3$name)

vitis_gbif_coords3$name <- "Vitis spp"

# tidy for maxent
vitis_gbif_coords3 <- vitis_gbif_coords3 %>%
  dplyr::select(name, latitude, longitude, everything()) %>% # change order of columns to MaxEnt convention
  dplyr::rename(species = name, x = longitude, y = latitude) %>% # change column names to MaxEnt convention
  dplyr::select(species, x, y) # select only columns needed for MaxEnt

write_csv(vitis_gbif_coords3, file = file.path(here(), "data", "vitis_gbif_cleaned_coords_v1.csv"))

```

The occurrence data for Vitis spp. are now in a useable format for MaxEnt.

# 3. Visualize Host Plant distributions

```{r visualize raw and thinned lydemapr records}

# raw data
lyde_raw <- read_csv(file = file.path(here(), "data-raw", "slf_lyde_raw_coords_2023-07-13.csv"))

# thinned data
lyde_thinned <- read_csv(file = file.path(here(), "data", "slf_lyde_cleaned_coords_v1.csv"))

# plot map of NAmerica
map_lyde_thinned <- ggplot() +
  # this plots a basemap of the USA because data here are exclusively from the USA
  geom_polygon(data = map_data('state'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
  geom_point(data = lyde_raw, aes(x = longitude, y = latitude), color = "blue", size = 2) +
  geom_point(data = lyde_thinned, aes(x = longitude, y = latitude), color = "red", shape = 2) +
  coord_quickmap(xlim = c(-133.593750, -52.294922), ylim = c(25.085599, 55.304138)) +
  ggtitle("SLF lydemapR records- \n raw (blue) vs thinned (red)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  theme_bw()
  
# include citizen science platforms
lyde_raw1 <- lyde_raw %>%
  filter(lyde_present == "TRUE", 
         lyde_established == "TRUE") %>% 
  mutate(species = "Lycorma delicatula")



# plot map including these records
map_lyde_allRecords <- ggplot() +
  # this plots a basemap of the USA because data here are exclusively from the USA
  geom_polygon(data = map_data('state'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
  geom_point(data = lyde_raw1, aes(x = longitude, y = latitude), color = "blue", size = 2) +
  geom_point(data = lyde_thinned, aes(x = longitude, y = latitude), color = "red", shape = 2) +
  coord_quickmap(xlim = c(-133.593750, -52.294922), ylim = c(25.085599, 55.304138)) +
  ggtitle("SLF lydemapR records- \n raw (blue) vs thinned (red)") +
  theme(panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank()) +
  theme_bw()

# patchwork display of plots
map_lyde_thinned + map_lyde_allRecords +
  plot_layout(ncol = 2)

```

```{r map all points}

 map_all <- ggplot() +
    geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
    geom_point(data = slf_all_coords3, aes(x = x, y = y), color = "red", size = 2) +
    coord_quickmap(xlim = c(-164.5, 163.5), ylim = c(-55, 85)) +
    ggtitle("All SLF presences") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank()) +
    theme_bw()

 map_all_NAmerica <- ggplot() +
    geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
    geom_point(data = slf_all_coords3, aes(x = x, y = y), color = "red", size = 2) +
    coord_quickmap(xlim = c(-133.593750, -52.294922), ylim = c(25.085599, 55.304138)) +
    ggtitle("SLF presences N AMerica") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank()) +
    theme_bw()
 
  map_all_Asia <- ggplot() +
    geom_polygon(data = map_data('world'), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
    geom_point(data = slf_all_coords3, aes(x = x, y = y), color = "red", size = 2) +
    coord_quickmap(xlim = c(68.906250, 152.534180), ylim = c(8.928487, 45.920587)) +
    ggtitle("SLF presences Asia") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank()) +
    theme_bw()

  # patchwork
  map_all_NAmerica + map_all_Asia +
  plot_layout(ncol = 2)
  
```


# Appendix

## Visualize changes in GBIF records between initial 2020 data pull and newest pull

First, we will visualize record changes in TOH

```{r visualize 2023 pulled gbif records vs 2020 pull- TOH}

if(FALSE) {
  
  # old data
  toh_2020 <- read_csv(file = file.path(here(), "data-raw", "data-old", "toh_gbif_cleaned_coords_2022.csv"))
  
  # new data
  toh_new <- read_csv(file = file.path(here(), "data", "toh_gbif_cleaned_v1.csv"))
  
  # plot toh
  map_toh <- ggplot() +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
    geom_point(data = toh_2020, aes(x = longitude, y = latitude), color = "red", size = 2) +
    geom_point(data = toh_new, aes(x = x, y = y), color = "blue", shape = 2) +
    coord_quickmap(xlim = c(-164.5, 163.5), ylim = c(-55, 85)) +
    ggtitle("TOH new points (red) vs old (blue)") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank()) +
    theme_bw()
  
  # new points are in red, old are in blue
}

```

Then for Vitis

```{r visualize 2023 pulled gbif records vs 2020 pull- Vitis}

if(FALSE) {
  
  # old data
  vitis_2020 <- read_csv(file = file.path(here(), "data-raw", "data-old", "NAME OF FILE"))
  
  # new data
  vitis_new <- read_csv(file = file.path(here(), "data", "vitis_gbif_cleaned_v1.csv"))
  
  # plot vitis
  map_vitis <- ggplot() +
    geom_polygon(data = map_data("world"), aes(x = long, y = lat, group = group), fill = NA, color = "black", lwd = 0.15) +
    geom_point(data = vitis_2020, aes(x = longitude, y = latitude), color = "red", size = 2) +
    geom_point(data = vitis_new, aes(x = x, y = y), color = "blue", shape = 2) +
    coord_quickmap(xlim = c(-164.5, 163.5), ylim = c(-55, 85)) +
    ggtitle("vitis new points (red) vs old (blue)") +
    theme(panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.background = element_blank()) +
    theme_bw()
  
  # new points are in red, old are in blue
}

```

# Citations

1. Ailanthus altissima (Mill.) Swingle in GBIF Secretariat (2022). GBIF Backbone Taxonomy. Checklist dataset https://doi.org/10.15468/39omei accessed via GBIF.org on 2023-07-.

2. Vitis L. in GBIF Secretariat (2022). GBIF Backbone Taxonomy. Checklist dataset https://doi.org/10.15468/39omei accessed via GBIF.org on 2023-07-.

