---
title: "test"
author: "Samuel M. Owens"
date: "2024-03-15"
output: html_document
---

# plot re-scaling

I need to test a custom function I wrote to assess whether this function performs correctly. It is supposed to center suitability values around a threshold as a median. I need to assess whether the same points are falling under the threshold using the function as without it.

```{r load necesssary packages, echo = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
here() # here() starts at the root folder of this package.
library(devtools)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# spatial data handling
library(raster) 
library(terra)
library(tidygeocoder)

# plot aesthetics
library(scales)
#install.packages("ggbreak")
library(ggbreak)
library(patchwork)

```

```{r set wd}

mypath <- file.path(here() %>% 
                       dirname(),
                     "maxent/models")

```

Load in summary files for thresholds.

```{r load in summary files}
# summary file to extract thresholds from

# global
summary_global <- read_csv(file = file.path(mypath, "slf_global_v2", "global_summary_all_iterations.csv"))


# regional_native
summary_regional_invaded <- read_csv(file = file.path(mypath, "slf_regional_invaded_v4", "regional_invaded_summary.csv"))


# regional_native
summary_regional_native <- read_csv(file = file.path(mypath, "slf_regional_native_v1", "regional_native_summary.csv"))


```


```{r load in xy suitability}

# global
slf_suit_global_hist <- read_csv(
  file = file.path(mypath, "slf_global_v2", "global_slf_all_coords_1981-2010_xy_pred_suit_clamped_cloglog_mean.csv")
  ) %>%
  dplyr::rename("slf_suit_global_hist" = "cloglog_suitability")

# regional_native
slf_suit_regional_native_hist <- read_csv(
  file = file.path(mypath, "slf_regional_native_v1", "regional_native_slf_all_coords_1981-2010_xy_pred_suit_clamped_cloglog.csv")
  ) %>%
  dplyr::rename("slf_suit_regional_native_hist" = "cloglog_suitability")

```


```{r re-scale based on MTSS thresh for visualization}


# global regular function
slf_suit_global_hist_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = slf_suit_global_hist,
  thresh = "MTSS",
  summary.file = summary_global,
  rescale.name = "slf_suit_global_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
slf_suit_global_hist_rescaled_thresholds <- slf_suit_global_hist_rescaled[[2]]
slf_suit_global_hist_rescaled <- slf_suit_global_hist_rescaled[[1]]


# regional_native
slf_suit_regional_native_hist_rescaled <- slfSpread::rescale_cloglog_suitability(
  xy.predicted = slf_suit_regional_native_hist,
  thresh = "MTSS",
  summary.file = summary_regional_native,
  rescale.name = "slf_suit_regional_native_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
slf_suit_regional_native_hist_rescaled_thresholds <- slf_suit_regional_native_hist_rescaled[[2]]
slf_suit_regional_native_hist_rescaled <- slf_suit_regional_native_hist_rescaled[[1]]


```

First, I will run some minor tests to make sure that the records being re-scaled relatively match the original records.

```{r}

# filter out above and below MTSS threshold for original dataset
test1 <- filter(slf_suit_global_hist, slf_suit_global_hist < 0.03403)
# filter same for new dataset, but based on 0.5 median threshold
test2 <- filter(slf_suit_global_hist_rescaled, slf_suit_global_hist_rescaled < 0.5)
# anit_join
anti_join(test1, test2, by = c("x", "y"))

```



```{r slf_global_invaded plot style}

scatter_style_global_native <- list(
  # aesthetics
   labs(title = "Current suitability for SLF establishment at locations with known SLF populations",
         subtitle = "'global' vs 'regional_native' models | 1981-2010",
         caption = "thresholds: dashed = MTSS, twodash = 10_percentile, dotted = MTP"
         ),
   xlab("'global' model suitability"),
   theme_bw()#,
  # MTP threshold
  #geom_vline(xintercept = as.numeric(summary_global[30, ncol(summary_global)]), linetype = "dotted", alpha = 0.5),
  #geom_hline(yintercept = as.numeric(summary_regional_invaded[30, 2]), linetype = "dotted", alpha = 0.5),
  # MTSS threshold
  #geom_vline(xintercept = as.numeric(summary_global[42, ncol(summary_global)]), linetype = "dashed", alpha = 0.5),
  #geom_hline(yintercept = as.numeric(summary_regional_invaded[42, 2]), linetype = "dashed", alpha = 0.5)
)

```

Now I will plot the data

```{r test plots}

# regular function
# join datasets for plotting
slf_invaded_global_rescaled_joined <- full_join(slf_suit_regional_native_hist_rescaled, slf_suit_global_hist_rescaled, by = c("Species", "x", "y"))

# plot
(ggplot() +
   # plot style
   scatter_style_global_native +
   # MTSS thresh
   geom_vline(xintercept = 0.5, linetype = "dashed") + 
   geom_hline(yintercept = 0.5, linetype = "dashed") +
   # MTP threshold
   geom_vline(xintercept = as.numeric(slf_suit_global_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(slf_suit_regional_native_hist_rescaled_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   # data
   geom_point(data = slf_invaded_global_rescaled_joined, 
               aes(x = slf_suit_global_hist_rescaled, y = slf_suit_regional_native_hist_rescaled), 
               color = "darkorange", size = 1
               ) #+
   #scale_x_continuous(name = "'global' model risk factor", limits = c(0, 1), breaks = breaks) + 
  # scale_y_continuous(name = "'regional_native' risk factor", limits = c(0, 1), breaks = breaks) 
)


```

```{r test variant of this function}

# global 1_1 function
slf_suit_global_hist_rescaled_1to1 <- slfSpread::rescale_cloglog_suitability_1to1(
  xy.predicted = slf_suit_global_hist,
  thresh = "MTSS",
  summary.file = summary_global,
  rescale.name = "slf_suit_global_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
slf_suit_global_hist_rescaled_1to1_thresholds <- slf_suit_global_hist_rescaled_1to1[[2]]
slf_suit_global_hist_rescaled_1to1 <- slf_suit_global_hist_rescaled_1to1[[1]]



# regional_native
slf_suit_regional_native_hist_rescaled_1to1 <- slfSpread::rescale_cloglog_suitability_1to1(
  xy.predicted = slf_suit_regional_native_hist,
  thresh = "MTSS",
  summary.file = summary_regional_native,
  rescale.name = "slf_suit_regional_native_hist",
  rescale.thresholds = TRUE
)
# extract rescaled values and thresholds
slf_suit_regional_native_hist_rescaled_1to1_thresholds <- slf_suit_regional_native_hist_rescaled_1to1[[2]]
slf_suit_regional_native_hist_rescaled_1to1 <- slf_suit_regional_native_hist_rescaled_1to1[[1]]

# 1to1 function
# join datasets for plotting
slf_invaded_global_rescaled_1to1_joined <- full_join(slf_suit_regional_native_hist_rescaled_1to1, slf_suit_global_hist_rescaled_1to1, by = c("Species", "x", "y"))

# plot
(ggplot() +
   # plot style
   scatter_style_global_native +
   # MTSS thresh
   geom_vline(xintercept = 0, linetype = "dashed") + 
   geom_hline(yintercept = 0, linetype = "dashed") +
   # MTP threshold
   geom_vline(xintercept = as.numeric(slf_suit_global_hist_rescaled_1to1_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   geom_hline(yintercept = as.numeric(slf_suit_regional_native_hist_rescaled_1to1_thresholds[1, 2]), linetype = "dotted", alpha = 0.5) +
   # data
   geom_point(data = slf_invaded_global_rescaled_1to1_joined, 
               aes(x = slf_suit_global_hist_rescaled, y = slf_suit_regional_native_hist_rescaled), 
               color = "darkorange", size = 1
               ) #+
   #scale_x_continuous(name = "'global' model risk factor", limits = c(0, 1), breaks = breaks) + 
  # scale_y_continuous(name = "'regional_native' risk factor", limits = c(0, 1), breaks = breaks) 
)

```



```{r testing other functions from scales package}
slf_suit_global_hist_rescaled <- scales::rescale_mid(
  x = slf_suit_global_hist$slf_suit_global_hist,
  to = c(-1, 1),
  from = range(slf_suit_global_hist$slf_suit_global_hist),
  mid = as.numeric(summary_global[42, ncol(summary_global)])
) %>%
  as.data.frame() %>%
  rename("slf_suit_global_hist_rescaled" = ".")

slf_suit_global_hist_rescaled <- cbind(slf_suit_global_hist, slf_suit_global_hist_rescaled) %>%
  dplyr::select(-slf_suit_global_hist)


slf_suit_regional_native_hist_rescaled <- scales::rescale_mid(
  x = slf_suit_regional_native_hist$slf_suit_regional_native_hist,
  to = c(-1, 1),
  from = range(slf_suit_regional_native_hist$slf_suit_regional_native_hist),
  mid = as.numeric(summary_regional_native[42, ncol(summary_regional_native)])
) %>%
  as.data.frame() %>%
  rename("slf_suit_regional_native_hist_rescaled" = ".")

slf_suit_regional_native_hist_rescaled <- cbind(slf_suit_regional_native_hist, slf_suit_regional_native_hist_rescaled) %>%
  dplyr::select(-slf_suit_regional_native_hist)

```

